---
title: "Baseline Variate Preparation for Analysis"
author: "Jesus Rodriguez, Sydney Shimizu, Julia Aptekar"
format: html
embed-resources: true
---

## Libraries/Loading
```{r setup}
library(tidyverse) # general tidying and manipulation
library(readxl) # reading xlsx files
library(kableExtra) # nice tables
library(lubridate) # easier working with dates
library(survival)
library(survminer)
library(table1)
library(cmprsk)
library(KMsurv)
# d <- readRDS("data/NEW_DATA.rds")
heart3 <- 
  read_csv("data/heart3.csv")
# analysis_data <-
#   readRDS("data/analysis_data.rds")
```

# Baseline Variates

```{r}
baseline_variates <- 
  heart3 |> 
  select(WL_ID_CODE,
         INIT_AGE, 
         age_at_status2, 
         GENDER_NEW,
         INIT_BMI_NEW,
         ETHCAT_NEW,
         ETHCAT2_NEW,
         BTYPE_REC,
         DIAB,     # THERE ARE FOUR `Diabetes Status Unknown` obs in `DIAB`
         DIAB_Y,   # DR. KAWAGUCHI seemed to set them as NOT HAVING DIABETES
         VENTILATOR_TCR,
         PRIOR_CARD_SURG_TCR,
         PRIOR_CARD_SURG_TCR_Y,
         ISCH_CARD_Y,
         # DIAL_AFTER_LIST,
         # DIAL_AFTER_LIST_Y,
         
         group, # stable / escalated variable
         
         FTIME_NEW, # time and event variables
         FSTATUS_NEW,         #
         END_DATE_NEW     #
  ) |> 
  mutate(
    across(
      c(INIT_AGE,
        age_at_status2,
        INIT_BMI_NEW, 
        FTIME_NEW),
      as.numeric
    ),
    across(
      c(WL_ID_CODE,
        GENDER_NEW,
        ETHCAT_NEW,
        ETHCAT2_NEW,
        BTYPE_REC,
        DIAB,
        DIAB_Y,
        VENTILATOR_TCR,
        PRIOR_CARD_SURG_TCR,
        PRIOR_CARD_SURG_TCR_Y,
        ISCH_CARD_Y,
        # DIAL_AFTER_LIST,
        # DIAL_AFTER_LIST_Y,
        FSTATUS_NEW),
      as.factor
    ),
    group = factor(group,
                   levels = c(0,1),
                   labels = c("Stable", "Escalated")),
    event_labels = factor(FSTATUS_NEW,
                   levels = c(1,2,3,4,5,6,0),
                   labels = c("Death/Deterioration",
                              "Transplantation",
                              "Recovery",
                              "Upgrade to Status 1",
                              "Downgrade to Status 3-6",
                              "Delisted",
                              "Censored"))
  )

write_rds(base)         
```

# Demographics

## Amount of Each Event Table

```{r}
baseline_variates |> 
  group_by(FSTATUS_NEW) |> 
  count()
```

## Table 1

### Table Prep

```{r}
label(baseline_variates$INIT_AGE) <- "Age at Initial Listing"
label(baseline_variates$age_at_status2) <- "Age when cand. reached Status 2"
label(baseline_variates$INIT_BMI_NEW) <- "BMI at Initial Listing"
label(baseline_variates$GENDER_NEW) <- "Gender"
label(baseline_variates$ETHCAT_NEW) <- "Race/Ethnicity (unsimplified)"
label(baseline_variates$ETHCAT2_NEW) <- "Race/Ethnicity"
label(baseline_variates$BTYPE_REC) <- "Blood Type"
label(baseline_variates$DIAB) <- "Diabetes Status"
label(baseline_variates$DIAB_Y) <- "Diabetes Status (0/1)"
label(baseline_variates$VENTILATOR_TCR) <- "Patient on Life Support (0/1)"
label(baseline_variates$PRIOR_CARD_SURG_TCR) <- "Prior Cardiace Sx at Listing (Non-Transplant)"
label(baseline_variates$PRIOR_CARD_SURG_TCR_Y) <- "Prior Cardiace Sx at Listing (Non-Transplant) (0/1)"
label(baseline_variates$ISCH_CARD_Y) <- "ISCH CARD (0/1)" # NOT SURE ABOUT THIS ONE
# label(baseline_variates$DIAL_AFTER_LIST) <- "Dialysis Occurring Between Listing and Transplant"
# label(baseline_variates$DIAL_AFTER_LIST) <- "Dialysis Occurring Between Listing and Transplant (0/1)"
label(baseline_variates$group) <- "Escalated Status"

```

### P-value shenanigans 

code from table1 CRAN documentation for pvalues

```{r}
pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- chisq.test(table(y, g))$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}
```



### The Table 1

```{r, eval = F}
# All variates

# Even the duplicates
# I have evaluate = FALSE right now for this one
# Don't think we'll need it

table1(
  ~ INIT_AGE +
    age_at_status2 +
    INIT_BMI_NEW +
    GENDER_NEW +
    ETHCAT_NEW +
    ETHCAT2_NEW +
    BTYPE_REC +
    DIAB +
    DIAB_Y +
    VENTILATOR_TCR +
    PRIOR_CARD_SURG_TCR +
    PRIOR_CARD_SURG_TCR_Y +
    ISCH_CARD_Y # +
    # DIAL_AFTER_LIST +
    # DIAL_AFTER_LIST_Y
  | group,
  data = baseline_variates,
  overall = "Overall"
)
```

#### Descriptive Statistics Table (That we Most Likely will Use)

```{r}
# Only the simple ones

# We will probably want to use this one

table1(
  ~ age_at_status2 +
    INIT_BMI_NEW +
    GENDER_NEW +
    ETHCAT2_NEW +
    BTYPE_REC +
    DIAB_Y +
    VENTILATOR_TCR +
    PRIOR_CARD_SURG_TCR_Y +
    ISCH_CARD_Y #+
    # DIAL_AFTER_LIST_Y
  | group,
  data = baseline_variates,
  caption = "Table 1.1 Descriptive Statistics by Stable/Escalated Groups"
  # overall = F
  # extra.col = list(`p-value`=pvalue)
)
```

#### Overall Statistics Table

```{r}
# Overall Statistics
table1(
  ~ group + 
    age_at_status2 +
    INIT_BMI_NEW +
    GENDER_NEW +
    ETHCAT2_NEW +
    BTYPE_REC +
    DIAB_Y +
    VENTILATOR_TCR +
    PRIOR_CARD_SURG_TCR_Y +
    ISCH_CARD_Y,
    # DIAL_AFTER_LIST_Y,
  data = baseline_variates,
)
```

#### Group and Event Tables

Very big table (21 columns), probably unncessary


```{r, eval = F}
# Table of covariates by group AND FSTATUS_NEW
table1(
  ~ age_at_status2 +
    INIT_BMI_NEW +
    GENDER_NEW +
    ETHCAT2_NEW +
    BTYPE_REC +
    DIAB_Y +
    VENTILATOR_TCR +
    PRIOR_CARD_SURG_TCR_Y +
    ISCH_CARD_Y
    # DIAL_AFTER_LIST_Y
  | group*event_labels,
  data = baseline_variates
  )
```
##### Descriptive Statistics by Outcome Table

```{r}

# Overall population
table1(
  ~ age_at_status2 +
    INIT_BMI_NEW +
    GENDER_NEW +
    ETHCAT2_NEW +
    BTYPE_REC +
    DIAB_Y +
    VENTILATOR_TCR +
    PRIOR_CARD_SURG_TCR_Y +
    ISCH_CARD_Y |
    event_labels,
  data = baseline_variates,
  overall = "Overall",
  caption = "Table 1.2: Descriptive Statistics by Outcome"
)
```




# Covariate Matrix Preparation

For the Fine-Gray models

*I commmented out the "duplicate" variables in case we will need them later*

```{r}
vars <- c(
  "FTIME_NEW",
  "FSTATUS_NEW",
  "group",
  # "INIT_AGE", ####
  "age_at_status2",
  "INIT_BMI_NEW",
  "GENDER_NEW",
  # "ETHCAT_NEW", ####
  "ETHCAT2_NEW",
  "BTYPE_REC",
  # "DIAB", ####
  "DIAB_Y",
  "VENTILATOR_TCR",
  # "PRIOR_CARD_SURG_TCR", ####
  "PRIOR_CARD_SURG_TCR_Y",
  "ISCH_CARD_Y"
  # "DIAL_AFTER_LIST", ####
  # "DIAL_AFTER_LIST_Y"
)  
```



```{r}
heart_matrix <- 
  baseline_variates[complete.cases(baseline_variates[, vars]), vars]
```



## Unadjusted Covariate Matrix

```{r}
covariate_matrix_unadjusted <- 
  model.matrix(
    ~ group,
    data = heart_matrix
  )[, -1]

# str(heart_matrix)
```

## Adjusted Covariate Matrix


```{r}
covariate_matrix_adjusted <- 
  model.matrix(
    ~ group +
      # INIT_AGE + 
      age_at_status2 +  # updated age adjusted to landmark
      GENDER_NEW +
      INIT_BMI_NEW +
      # ETHCAT_NEW + 
      ETHCAT2_NEW +     # simplified ethnicity var
      BTYPE_REC +
      # DIAB + 
      DIAB_Y +   # yes/no diabetes indicator 
      VENTILATOR_TCR +
      # PRIOR_CARD_SURG_TCR + 
      PRIOR_CARD_SURG_TCR_Y +  # binary indicator
      ISCH_CARD_Y,
      # DIAL_AFTER_LIST + 
      # DIAL_AFTER_LIST_Y,       # binary indicator
    data = heart_matrix
  )[, -1]

# str(heart_matrix)
```

# Unadjusted Models

## Cox Models

```{r}
# Cox Model 1 : Death / Deterioration
cs_model_death_unadjusted <- 
  baseline_variates %>% 
  coxph(
  Surv(FTIME_NEW, FSTATUS_NEW == 1) ~ 
    group, 
  data = .
)
summary(cs_model_death_unadjusted)

```

```{r}
# Cox Model 2 : Transplant
cs_model_transplant_unadjusted <- 
  baseline_variates %>% 
  coxph(
    Surv(FTIME_NEW, FSTATUS_NEW == 2) ~ group,
    data = .
  ) ; summary(cs_model_transplant_unadjusted)

```

```{r}
# Cox Model 3 : Recovery
cs_model_recovery_unadjusted <- 
  baseline_variates %>% 
  coxph(
    Surv(FTIME_NEW, FSTATUS_NEW == 3) ~ group,
    data = .
  ) ; summary(cs_model_recovery_unadjusted)
```

```{r}
# Cox Model 4: Upgrade to Status 1
cs_model_upgrade_unadjusted <- 
  baseline_variates %>% 
  coxph(
    Surv(FTIME_NEW, FSTATUS_NEW == 4) ~ group,
    data = .
  ) ; summary(cs_model_upgrade_unadjusted)
```

```{r}
# Cox Model 5: Downgrade to Status 3-6
cs_model_downgrade_unadjusted <- 
  baseline_variates %>% 
  coxph(
    Surv(FTIME_NEW, FSTATUS_NEW == 5) ~ group,
    data = .
  ) ; summary(cs_model_downgrade_unadjusted)
```

```{r}
# Cox Model 6: Delisted
cs_model_delisted_unadjusted <- 
  baseline_variates %>% 
  coxph(
    Surv(FTIME_NEW, FSTATUS_NEW == 6) ~ group,
    data = .
  ) ; summary(cs_model_delisted_unadjusted)
```

## Fine Gray Models 

### Death/Deterioration

```{r, error = T}
fg_death_unadjusted <- 
  crr(
    ftime = heart_matrix$FTIME_NEW,
    fstatus = heart_matrix$FSTATUS_NEW,
    cov1 = covariate_matrix_unadjusted, 
    failcode = 1
  ) ; summary(fg_death_unadjusted)
```

### Transplant

```{r}
fg_transplant_unadjusted <- 
  crr(
    ftime = heart_matrix$FTIME_NEW,
    fstatus = heart_matrix$FSTATUS_NEW,
    cov1 = covariate_matrix_unadjusted, 
    failcode = 2
  ) ; summary(fg_transplant_unadjusted)
```
### Recovery 

```{r, error = T}
fg_recovery_unadjusted <- 
  crr(
    ftime = heart_matrix$FTIME_NEW,
    fstatus = heart_matrix$FSTATUS_NEW,
    cov1 = covariate_matrix_unadjusted,
    failcode = 3
  ) ; summary(fg_recovery_unadjusted)
```

### Upgrade to Status 1

```{r}
fg_upgrade_unadjusted <- 
  crr(
    ftime = heart_matrix$FTIME_NEW,
    fstatus = heart_matrix$FSTATUS_NEW,
    cov1 = covariate_matrix_unadjusted,
    failcode = 4
  ) ; summary(fg_upgrade_unadjusted)
```

### Downgrade to Status 3-6

```{r}
fg_downgrade_unadjusted <- 
  crr(
    ftime = heart_matrix$FTIME_NEW,
    fstatus = heart_matrix$FSTATUS_NEW,
    cov1 = covariate_matrix_unadjusted,
    failcode = 5
  ) ; summary(fg_downgrade_unadjusted)
```

### Delisted

```{r}
fg_delisted_unadjusted <- 
  crr(
    ftime = heart_matrix$FTIME_NEW,
    fstatus = heart_matrix$FSTATUS_NEW,
    cov1 = covariate_matrix_unadjusted,
    failcode = 6
  ) ; summary(fg_delisted_unadjusted)
```

## Binomial Regression Models



```{r}
library(mets)
options(warn = 1)
```

### Death / Deterioration

```{r}
# Logistic Regression with IPCW Binomial Regression (Unadjusted) (Death)

baseline_variates$CAUSE_DEATH <- ifelse(baseline_variates$FSTATUS_NEW == 1, 1, 0)

binreg_death_unadjusted <- binreg(
  Event(FTIME_NEW, CAUSE_DEATH) ~ group,
  data = baseline_variates,
  time = 90
) ; summary(binreg_death_unadjusted)



```

### Transplant 

```{r}
# Logistic Regression with IPCW Binomial Regression (Unadjusted) (Transplant)

baseline_variates$CAUSE_T <- ifelse(baseline_variates$FSTATUS_NEW == 2, 1, 0)

binreg_transplant_unadjusted <- binreg(
  Event(FTIME_NEW, CAUSE_T) ~ group,
  data = baseline_variates,
  time = 90 
) ; summary(binreg_transplant_unadjusted)

```
### Recovery 

```{r}
# Logistic Regression with IPCW Binomial Regression (Unadjusted) (Recovery)

baseline_variates$CAUSE_R <- ifelse(baseline_variates$FSTATUS_NEW == 3, 1, 0)

binreg_recovery_unadjusted <- binreg(
  Event(FTIME_NEW, CAUSE_R) ~ group,
  data = baseline_variates,
  time = 90 
) ; summary(binreg_recovery_unadjusted)

```
### Upgrade

```{r}
# Logistic Regression with IPCW Binomial Regression (Unadjusted) ( upgrade)

baseline_variates$CAUSE_U <- ifelse(baseline_variates$FSTATUS_NEW == 4, 1, 0)

binreg_upgrade_unadjusted <- binreg(
  Event(FTIME_NEW, CAUSE_U) ~ group,
  data = baseline_variates,
  time = 90 
) ; summary(binreg_upgrade_unadjusted)

```
### Downgrade

```{r}
# Logistic Regression with IPCW Binomial Regression (Unadjusted) (downgrade)

baseline_variates$CAUSE_DG <- ifelse(baseline_variates$FSTATUS_NEW == 5, 1, 0)

binreg_downgrade_unadjusted <- binreg(
  Event(FTIME_NEW, CAUSE_DG) ~ group,
  data = baseline_variates,
  time = 90 
) ; summary(binreg_downgrade_unadjusted)

```

### Delisted

```{r}
# Logistic Regression with IPCW Binomial Regression (Unadjusted) (delisted)

baseline_variates$CAUSE_Dl <- ifelse(baseline_variates$FSTATUS_NEW == 6, 1, 0)

binreg_delisted_unadjusted <- binreg(
  Event(FTIME_NEW, CAUSE_Dl) ~ group,
  data = baseline_variates,
  time = 90 
) ; summary(binreg_delisted_unadjusted)

```

# Kable Table Prep

Assuming we want to make the exact same table from the poster, my code should work the same... hopefully

## Function to extract the info we need from the cox models

```{r}
# Extract from coxph model
extract_cox <- function(model, outcome) {
  s <- summary(model)
  data.frame(
    Outcome  = outcome,
    Estimate       = s$conf.int[1, "exp(coef)"],
    Lower    = s$conf.int[1, "lower .95"],
    Upper    = s$conf.int[1, "upper .95"],
    P        = s$coefficients[1, "Pr(>|z|)"]
  )
}
```

## Function to extract the info we need from the Fine-Gray models


```{r}
# Extract from crr model
extract_crr <- function(model, outcome) {
  s <- summary(model)
  data.frame(
    Outcome  = outcome,
    Estimate       = s$conf.int[1, "exp(coef)"],
    Lower    = s$conf.int[1, "2.5%"],
    Upper    = s$conf.int[1, "97.5%"],
    P        = s$coef[1, "p-value"]
  )
}
```

## Extract from binreg() model

```{r}

# Extract from binreg model
extract_binreg <- function(model, outcome) {
  s <- summary(model)
  data.frame(
    Outcome  = outcome,
    Estimate       = s$exp[2, "Estimate"],
    Lower    = s$exp[2, "2.5%"],
    Upper    = s$exp[2, "97.5%"],
    P        = s$coef[2, "P-value"]
  )
}
```


## Making the dataframe for the table 

```{r}
results_unadjusted <- 
  bind_rows(
    # Cox
    extract_cox(cs_model_death_unadjusted, "Death / Deterioration"),
    extract_cox(cs_model_transplant_unadjusted, "Transplantation"),
    # extract_cox(cs_model_recovery_unadjusted, "Recovery"),
    extract_cox(cs_model_upgrade_unadjusted, "Upgrade to Status 1"),
    extract_cox(cs_model_downgrade_unadjusted, "Downgrade to Status 3-6"),
    extract_cox(cs_model_delisted_unadjusted, "Delisted"),

    # Fine Gray
    extract_crr(fg_death_unadjusted, "Death / Deterioration"),
    extract_crr(fg_transplant_unadjusted, "Transplantation"),
    # extract_crr(fg_recovery_unadjusted, "Recovery"),
    extract_crr(fg_upgrade_unadjusted, "Upgrade to Status 1"),
    extract_crr(fg_downgrade_unadjusted, "Downgrade to Status 3-6"),
    extract_crr(fg_delisted_unadjusted, "Delisted"),

    # Direct Binomial Regression
    extract_binreg(binreg_death_unadjusted, "Death / Deterioration"),
    extract_binreg(binreg_transplant_unadjusted, "Transplantation"),
    # extract_binreg(binreg_recovery_unadjusted, "Recovery"),
    extract_binreg(binreg_upgrade_unadjusted, "Upgrade to Status 1"),
    extract_binreg(binreg_downgrade_unadjusted, "Downgrade to Status 3-6"),
    extract_binreg(binreg_delisted_unadjusted, "Delisted")
  ) |>
  mutate(
    Estimate = round(Estimate, 2),
    CI = paste0("(", round(Lower, 2), " , ", round(Upper, 2), ")"),
    P  = ifelse(P < 0.001, "P < 0.001", sprintf("%.3f", P))
  ) |>
  select(Outcome, Estimate, CI, P)

# results_adjusted$P[6] <- cell_spec(results_adjusted$P[6], color = "#ED907E")
```

# The Unadjusted Kable Table

```{r}
kable_table_unadjusted <- 
  results_unadjusted |>
   kable(
    col.names = c("Outcome", "Estimate**", "95% CI", "p-value"),
    align = c("l", "c", "c", "c"),
                  
    caption = "<span style='font-size:25px; text-align:center; font-weight:bold; display:block;'>Table 2. Unadjusted Effect Estimates for Waitlist Outcomes</span>",
  ) |> 
  kable_styling(
    full_width = T,
    # position = 'left',
    font_size = 19
  ) |> 
  column_spec(1, width = '10em', bold = T, color = "black") |> 
  column_spec(2:4, width = "7em", bold = F, color = "black") |> 
  
  pack_rows("Cause-Specific Cox Model (HR*)", 1, 5) |>
  
  pack_rows("Fine-Gray Subdistribution Model (SHR*)", 6, 10) |> 
  pack_rows("Direct Binomial Regression Model (OR*)", 11, 15)
```

```{r}
# Run this if you made changes to the table and want to save it as a png

# sometimes I have problems with `save_kable()` and this usually fixes it, but you need to know where your browser is installed in your device

# Sys.setenv(
#   CHROMOTE_CHROME =
#     "C:/Program Files (x86)/Microsoft/Edge/Application/msedge.exe"
# )


save_kable(kable_table_unadjusted, file = "tables/new_unadjusted_results_table.png")
```

```{r, error = T}
kable_table_unadjusted
```
















# Adjusted Models

## Cox Models

I commented out the "duplicate" variables in case we will need them later in the models

### Death / Deterioration 

```{r, error = T}
# Cox Model 1 : Death / Deterioration
cs_model_death_adjusted <- 
  baseline_variates %>% 
  coxph(
  Surv(FTIME_NEW, FSTATUS_NEW == 1) ~ 
    group + 
    # INIT_AGE + 
    age_at_status2 + 
    INIT_BMI_NEW + 
    GENDER_NEW +
    # ETHCAT_NEW +
    ETHCAT2_NEW + 
    BTYPE_REC +
    # DIAB + 
    DIAB_Y +
    VENTILATOR_TCR + 
    # PRIOR_CARD_SURG_TCR + 
    PRIOR_CARD_SURG_TCR_Y + 
    ISCH_CARD_Y, 
    # DIAL_AFTER_LIST + 
    # DIAL_AFTER_LIST_Y,
  data = .
) 
summary(cs_model_death_adjusted)
```

### Transplant

```{r}
# Cox Model 2 : Transplant
cs_model_transplant_adjusted <- 
  baseline_variates %>% 
  coxph(
  Surv(FTIME_NEW, FSTATUS_NEW == 2) ~ 
    group + 
    # INIT_AGE + 
    age_at_status2 + 
    INIT_BMI_NEW + 
    GENDER_NEW +
    # ETHCAT_NEW +
    ETHCAT2_NEW + 
    BTYPE_REC +
    # DIAB + 
    DIAB_Y +
    VENTILATOR_TCR + 
    # PRIOR_CARD_SURG_TCR + 
    PRIOR_CARD_SURG_TCR_Y + 
    ISCH_CARD_Y, 
    # DIAL_AFTER_LIST + 
    # DIAL_AFTER_LIST_Y,
  data = .
) ; summary(cs_model_transplant_adjusted)
```

### Recovery

!! ERROR

```{r, error = T}
# Cox Model 3 : Recovery
cs_model_recovery_adjusted <-
  baseline_variates %>%
  coxph(
    Surv(FTIME_NEW, FSTATUS_NEW == 3) ~
      group +
      # INIT_AGE +
      age_at_status2 +
      INIT_BMI_NEW +
      GENDER_NEW +
      # ETHCAT_NEW +
      
      ETHCAT2_NEW +
      BTYPE_REC +
      # DIAB + 
      
      DIAB_Y +
      VENTILATOR_TCR +
      # PRIOR_CARD_SURG_TCR + 
      
      PRIOR_CARD_SURG_TCR_Y +
      ISCH_CARD_Y,
      # DIAL_AFTER_LIST + 
      
      # DIAL_AFTER_LIST_Y,
    data = .
  )

summary(cs_model_recovery_adjusted)
```


### Upgrade 

```{r}
# Cox Model 4 : Upgrade
cs_model_upgrade_adjusted <-
  baseline_variates %>%
  coxph(
    Surv(FTIME_NEW, FSTATUS_NEW == 4) ~
      group +
      # INIT_AGE +
      age_at_status2 +
      INIT_BMI_NEW +
      GENDER_NEW +
      # ETHCAT_NEW +
      
      ETHCAT2_NEW +
      BTYPE_REC +
      # DIAB + 
      
      DIAB_Y +
      VENTILATOR_TCR +
      # PRIOR_CARD_SURG_TCR + 
      
      PRIOR_CARD_SURG_TCR_Y +
      ISCH_CARD_Y,
      # DIAL_AFTER_LIST + 
      
      # DIAL_AFTER_LIST_Y,
    data = .
  )

summary(cs_model_upgrade_adjusted)
```


### Downgrade 

```{r}
# Cox Model 5 : Downgrade
cs_model_downgrade_adjusted <-
  baseline_variates %>%
  coxph(
    Surv(FTIME_NEW, FSTATUS_NEW == 5) ~
      group +
      # INIT_AGE +
      age_at_status2 +
      INIT_BMI_NEW +
      GENDER_NEW +
      # ETHCAT_NEW +
      
      ETHCAT2_NEW +
      BTYPE_REC +
      # DIAB + 
      
      DIAB_Y +
      VENTILATOR_TCR +
      # PRIOR_CARD_SURG_TCR + 
      
      PRIOR_CARD_SURG_TCR_Y +
      ISCH_CARD_Y,
      # DIAL_AFTER_LIST + 
      
      # DIAL_AFTER_LIST_Y,
    data = .
  )

summary(cs_model_downgrade_adjusted)
```


### Delisted

```{r}
# Cox Model 6 : Delisted
cs_model_delisted_adjusted <-
  baseline_variates %>%
  coxph(
    Surv(FTIME_NEW, FSTATUS_NEW == 6) ~
      group +
      # INIT_AGE +
      age_at_status2 +
      INIT_BMI_NEW +
      GENDER_NEW +
      # ETHCAT_NEW +
      
      ETHCAT2_NEW +
      BTYPE_REC +
      # DIAB + 
      
      DIAB_Y +
      VENTILATOR_TCR +
      # PRIOR_CARD_SURG_TCR + 
      
      PRIOR_CARD_SURG_TCR_Y +
      ISCH_CARD_Y,
      # DIAL_AFTER_LIST + 
      
      # DIAL_AFTER_LIST_Y,
    data = .
  )

summary(cs_model_delisted_adjusted)
```


## Fine Gray Models

### Death/Deterioration Model


```{r, error = T}
fg_death_adjusted <- 
  crr(
    ftime = heart_matrix$FTIME_NEW,
    fstatus = heart_matrix$FSTATUS_NEW,
    cov1 = covariate_matrix_adjusted, 
    failcode = 1
  )
summary(fg_death_adjusted)
```

### Transplantation

```{r}
fg_transplant_adjusted <- 
crr(
  ftime = heart_matrix$FTIME_NEW,
  fstatus = heart_matrix$FSTATUS_NEW,
  cov1 = covariate_matrix_adjusted, 
  failcode = 2
) ; summary(fg_transplant_adjusted)
```

### Recovery 


```{r, error = T}
fg_recovery_adjusted <-
  crr(
    ftime = heart_matrix$FTIME_NEW,
    fstatus = heart_matrix$FSTATUS_NEW,
    cov1 = covariate_matrix_adjusted,
    failcode = 3
    )

summary(fg_recovery_adjusted)
```


### Upgrade

```{r}
fg_upgrade_adjusted <-
  crr(
    ftime = heart_matrix$FTIME_NEW,
    fstatus = heart_matrix$FSTATUS_NEW,
    cov1 = covariate_matrix_adjusted,
    failcode = 4
    )

summary(fg_upgrade_adjusted)
```


### Downgrade

```{r}
fg_downgrade_adjusted <-
  crr(
    ftime = heart_matrix$FTIME_NEW,
    fstatus = heart_matrix$FSTATUS_NEW,
    cov1 = covariate_matrix_adjusted,
    failcode = 5
  )

summary(fg_downgrade_adjusted)
```


### Delisted

```{r}
fg_delisted_adjusted <-
  crr(
    ftime = heart_matrix$FTIME_NEW,
    fstatus = heart_matrix$FSTATUS_NEW,
    cov1 = covariate_matrix_adjusted,
    failcode = 6
  )

summary(fg_delisted_adjusted)
```


## Direct Binomial Regression Models

```{r}
# Dropping na rows

adjusted_binreg_data <- baseline_variates |> 
  drop_na()

```


### Death 

```{r, error = T}
binreg_death_adjusted <-
  binreg(
    Event(FTIME_NEW, CAUSE_DEATH) ~
      group + 
      # INIT_AGE + 
      age_at_status2 + 
      INIT_BMI_NEW + 
      GENDER_NEW +
      # ETHCAT_NEW +
      ETHCAT2_NEW + 
      BTYPE_REC +
      # DIAB + 
      DIAB_Y +
      VENTILATOR_TCR + 
      # PRIOR_CARD_SURG_TCR + 
      PRIOR_CARD_SURG_TCR_Y + 
      ISCH_CARD_Y, 
    # DIAL_AFTER_LIST + 
    # DIAL_AFTER_LIST_Y,
    data = adjusted_binreg_data,
    time = 90
    # ,
    # cens.model = ~strata(group)
  )

summary(binreg_death_adjusted)
```

### Transplant

```{r}
binreg_transplant_adjusted <-
  binreg(
    Event(FTIME_NEW, CAUSE_T) ~
      group +
      # INIT_AGE +
      age_at_status2 +
      INIT_BMI_NEW +
      GENDER_NEW +
      # ETHCAT_NEW +
      
      ETHCAT2_NEW +
      BTYPE_REC +
      # DIAB + 
      
      DIAB_Y +
      VENTILATOR_TCR +
      # PRIOR_CARD_SURG_TCR + 
      
      PRIOR_CARD_SURG_TCR_Y +
      ISCH_CARD_Y,
      # DIAL_AFTER_LIST + 
      
      # DIAL_AFTER_LIST_Y,
    data = adjusted_binreg_data,
    time = 90
  )

summary(binreg_transplant_adjusted)
```


### Recovery 

```{r}
binreg_recovery_adjusted <-
  binreg(
    Event(FTIME_NEW, CAUSE_R) ~
      group +
      # INIT_AGE +
      age_at_status2 +
      INIT_BMI_NEW +
      GENDER_NEW +
      # ETHCAT_NEW +
      
      ETHCAT2_NEW +
      BTYPE_REC +
      # DIAB + 
      
      DIAB_Y +
      VENTILATOR_TCR +
      # PRIOR_CARD_SURG_TCR + 
      
      PRIOR_CARD_SURG_TCR_Y +
      ISCH_CARD_Y,
      # DIAL_AFTER_LIST + 
      
      # DIAL_AFTER_LIST_Y,
    data = adjusted_binreg_data,
    time = 90
  )

summary(binreg_recovery_adjusted)
```



### Upgrade

```{r}
binreg_upgrade_adjusted <-
  binreg(
    Event(FTIME_NEW, CAUSE_U) ~
      group +
      # INIT_AGE +
      age_at_status2 +
      INIT_BMI_NEW +
      GENDER_NEW +
      # ETHCAT_NEW +
      
      ETHCAT2_NEW +
      BTYPE_REC +
      # DIAB + 
      
      DIAB_Y +
      VENTILATOR_TCR +
      # PRIOR_CARD_SURG_TCR + 
      
      PRIOR_CARD_SURG_TCR_Y +
      ISCH_CARD_Y,
      # DIAL_AFTER_LIST + 
      
      # DIAL_AFTER_LIST_Y,
    data = adjusted_binreg_data,
    time = 90
  )

summary(binreg_upgrade_adjusted)
```


### Downgrade

```{r}
binreg_downgrade_adjusted <-
  binreg(
    Event(FTIME_NEW, CAUSE_DG) ~
      group +
      # INIT_AGE +
      age_at_status2 +
      INIT_BMI_NEW +
      GENDER_NEW +
      # ETHCAT_NEW +
      
      ETHCAT2_NEW +
      BTYPE_REC +
      # DIAB + 
      
      DIAB_Y +
      VENTILATOR_TCR +
      # PRIOR_CARD_SURG_TCR + 
      
      PRIOR_CARD_SURG_TCR_Y +
      ISCH_CARD_Y,
      # DIAL_AFTER_LIST + 
      
      # DIAL_AFTER_LIST_Y,
    data = adjusted_binreg_data,
    time = 90
  )

summary(binreg_downgrade_adjusted)
```

### Delisted

```{r}
binreg_delisted_adjusted <-
  binreg(
    Event(FTIME_NEW, CAUSE_Dl) ~
      group +
      # INIT_AGE +
      age_at_status2 +
      INIT_BMI_NEW +
      GENDER_NEW +
      # ETHCAT_NEW +
      
      ETHCAT2_NEW +
      BTYPE_REC +
      # DIAB + 
      
      DIAB_Y +
      VENTILATOR_TCR +
      # PRIOR_CARD_SURG_TCR + 
      
      PRIOR_CARD_SURG_TCR_Y +
      ISCH_CARD_Y,
      # DIAL_AFTER_LIST + 
      
      # DIAL_AFTER_LIST_Y,
    data = adjusted_binreg_data,
    time = 90
  )

summary(binreg_delisted_adjusted)
```

# Kable Table Prep

## Making the dataframe for the table 

```{r}
results_adjusted <- 
  bind_rows(
    # Cox
    extract_cox(cs_model_death_adjusted, "Death / Deterioration"),
    extract_cox(cs_model_transplant_adjusted, "Transplantation"),
    # extract_cox(cs_model_recovery_adjusted, "Recovery"),
    extract_cox(cs_model_upgrade_adjusted, "Upgrade to Status 1"),
    extract_cox(cs_model_downgrade_adjusted, "Downgrade to Status 3-6"),
    extract_cox(cs_model_delisted_adjusted, "Delisted"),

    # Fine Gray
    extract_crr(fg_death_adjusted, "Death / Deterioration"),
    extract_crr(fg_transplant_adjusted, "Transplantation"),
    # extract_crr(fg_recovery_adjusted, "Recovery"),
    extract_crr(fg_upgrade_adjusted, "Upgrade to Status 1"),
    extract_crr(fg_downgrade_adjusted, "Downgrade to Status 3-6"),
    extract_crr(fg_delisted_adjusted, "Delisted"),

    # Direct Binomial Regression
    extract_binreg(binreg_death_adjusted, "Death / Deterioration"),
    extract_binreg(binreg_transplant_adjusted, "Transplantation"),
    # extract_binreg(binreg_recovery_adjusted, "Recovery"),
    extract_binreg(binreg_upgrade_adjusted, "Upgrade Status 1"),
    extract_binreg(binreg_downgrade_adjusted, "Downgrade to Status 3-6"),
    extract_binreg(binreg_delisted_adjusted, "Delisted")
  ) |>
  mutate(
    Estimate = round(Estimate, 2),
    CI = paste0("(", round(Lower, 2), " , ", round(Upper, 2), ")"),
    P  = ifelse(P < 0.001, "P < 0.001", sprintf("%.3f", P))
  ) |>
  select(Outcome, Estimate, CI, P)

# results_adjusted$P[6] <- cell_spec(results_adjusted$P[6], color = "#ED907E")
```

# The Adjusted Kable Table

```{r}
kable_table_adjusted <- 
  results_adjusted |>
  kable(
    col.names = c("Outcome", "Estimate**", "95% CI", "p-value"),
    align = c("l", "c", "c", "c"),
                  
    caption = "<span style='font-size:25px; text-align:center; font-weight:bold; display:block;'>Table 3. Adjusted Effect Estimates for Waitlist Outcomes</span>",
  ) |>
  kable_styling(
    full_width = T,
    position = 'left',
    font_size = 19
  ) |> 
  column_spec(1, bold = T, color = "black") |> 
  column_spec(2:4, bold = F, color = "black") |> 
  pack_rows("Cause-Specific Cox Model (HR*)", 1, 5) |>
  
  pack_rows("Fine-Gray Subdistribution Model (SHR*)", 6, 10) |> 
  
  pack_rows("Direct Binomial Regression Model (OR*)", 11, 15) 
```

```{r}
# Run this if you made changes to the table and want to save it as a png

# sometimes I have problems with `save_kable()` and this usually fixes it, but you need to know where your browser is installed in your device

# Sys.setenv(
#   CHROMOTE_CHROME =
#     "C:/Program Files (x86)/Microsoft/Edge/Application/msedge.exe"
# )
save_kable(kable_table_adjusted, file = "tables/new_adjusted_results_table.png")
```

```{r, error = T}
kable_table_adjusted
```


