---
title: "Exception Status Exploration"
author: "Jesus Rodriguez" 
format: html
embed-resources: true
---

This .qmd file is to further check on group\*exception interaction effects, it's similar to functional_models.qmd, but it has the formulas changed to include `group*EXCEPTION` 


## Libraries/Loading

```{r setup, message = F, warning = F, output = F}
library(tidyverse) # general tidying and manipulation
library(readxl) # reading xlsx files
library(kableExtra) # nice tables
library(lubridate) # easier working with dates
library(survival)
library(survminer)
library(table1)
library(cmprsk)
library(KMsurv)
library(mets)
options(warn = 1)

# 1. Make sure the data you want to load in is loaded in the setup chunk.

# # Our cleaned data
# heart3 <-
#   read_csv("data/heart3.csv")

# Dr. Kawaguchi's cleaned data
analysis_data0 <-
  readRDS("data/analysis_data_081126.rds")
  
```

## Cleaning in `analysis_data`

This cleans the age_at_status2, group, FSTATUS codes, and 90 day censoring variables and rules in Dr. Kawaguchi's code

If using our `heart3.csv` data, this code will produce an error. 


```{r, error = T}
analysis_data <- 
  analysis_data0 |> 
  mutate(
    age_at_status2 = 
      INIT_AGE + (as.numeric(STAT2_INIT - INIT_DATE)) / 365.25,
    group = 
      case_when(
        INIT_STAT2 == 1 ~ 0, # stable
        INIT_STAT2 == 0 ~ 1  # escalated
      ),
    FSTATUS_NEW = 
      case_when(
        FTIME_NEW > 90 ~ 0,
        TRUE ~ FSTATUS_NEW
      ),
    FTIME_NEW = 
      case_when(
        FTIME_NEW > 90 ~ 90,
        TRUE ~ FTIME_NEW
      )
  ) 
```


# Baseline Variates

```{r}
baseline_variates <- 
  # 2. In the Baseline Variates section, make sure the data you loaded in is being called in the `baseline_variates` pipeline
  
  analysis_data |> # change this dataset to the one you'd like to use
                   # heart3 or analysis_data
  
  
  select(WL_ID_CODE,
         INIT_AGE, 
         age_at_status2, 
         GENDER_NEW,
         INIT_BMI_NEW,
         ETHCAT_NEW,
         ETHCAT2_NEW,
         BTYPE_REC,
         DIAB,     
         DIAB_Y,   
         VENTILATOR_TCR,
         PRIOR_CARD_SURG_TCR,
         PRIOR_CARD_SURG_TCR_Y,
         ISCH_CARD_Y,
         # DIAL_AFTER_LIST,
         # DIAL_AFTER_LIST_Y,
         
         group, # stable / escalated variable
         
         
         # 3. If you are using the `heart3.csv` data as of 08/09/2026, 
         # make sure to comment out the EXCEPTION variable in the 
         # same `baseline_variates` pipeline from before.
         
         EXCEPTION, # <- <- <- exception variable
         
         FTIME_NEW, # time and event variables
         FSTATUS_NEW,         #
         END_DATE_NEW     #
  ) |> 
  mutate(
    across(
      c(INIT_AGE,
        age_at_status2,
        INIT_BMI_NEW, 
        FTIME_NEW),
      as.numeric
    ),
    across(
      c(WL_ID_CODE,
        GENDER_NEW,
        ETHCAT_NEW,
        ETHCAT2_NEW,
        BTYPE_REC,
        DIAB,
        DIAB_Y,
        VENTILATOR_TCR,
        PRIOR_CARD_SURG_TCR,
        PRIOR_CARD_SURG_TCR_Y,
        ISCH_CARD_Y,
        # DIAL_AFTER_LIST,
        # DIAL_AFTER_LIST_Y,
        EXCEPTION,
        FSTATUS_NEW),
      as.factor
    ),
    group = factor(group,
                   levels = c(0,1),
                   labels = c("Stable", "Escalated")),
    EXCEPTION = factor(EXCEPTION,
                   levels = c(0,1),
                   labels = c("No", "Yes")),
    ETHCAT2_NEW = factor(ETHCAT2_NEW,
                   levels = c("White", "Black", "Hispanic", "Asian", 'Other')),
    event_labels = factor(FSTATUS_NEW,     
                   levels = c(1,2,3,4,5,6,0),
                   labels = c("Death/Deterioration",
                              "Transplantation",
                              "Recovery",
                              "Upgrade to Status 1",
                              "Downgrade to Status 3-6",
                              "Delisted",
                              "Censored"))
  )

str(baseline_variates)

```

# Baseline Variates with No NA values in any rows

For the sake of simplicity, I am gonna have all the models run on complete data

This is just so each model is using the same data as the others

```{r}
baseline_variates_no_NA <- 
  baseline_variates |> 
  drop_na()
```

## NAs in FSTATUS_NEW Exploration

```{r}
FSTATUS_NAs <- baseline_variates |> 
  count(FSTATUS_NEW, name = "n_before") |> 
  full_join(
    baseline_variates_no_NA |> 
      count(FSTATUS_NEW, name = "n_after"),
    by = "FSTATUS_NEW"
  ) |> 
  mutate(
    n_before = replace_na(n_before, 0),
    n_after = replace_na(n_after, 0),
    n_removed = n_before - n_after,
    percent_removed = 100 * n_removed / n_before
  ) |> 
  bind_rows(
    data.frame(
      FSTATUS_NEW = "Overall",
      n_before = nrow(baseline_variates),
      n_after = nrow(baseline_variates_no_NA),
      n_removed = nrow(baseline_variates) - nrow(baseline_variates_no_NA),
      percent_removed = 100 * (
        nrow(baseline_variates) - nrow(baseline_variates_no_NA)
      ) / nrow(baseline_variates)
    )
  )

kable(FSTATUS_NAs)
```

# Defining Variables for Models

The `adjusted_variables` are the variables that will be in the formula for the ADJUSTED Cox and Direct Binomial Regression models 

Their adjusted formulas look like this: 

```
"Surv(FTIME_NEW, cause) ~ group +",
          age_at_status2 + GENDER_NEW + ...
```

While their unadjusted formulas won't need these variables: 

```
Surv(FTIME_NEW, cause) ~ group,
```

!!!These `adjusted_variables` DO NOT HAVE a group * EXCEPTION interaction term!!!


```{r}
# adjusted_variables <- c(
#   # 'INIT_AGE' , 
#   'age_at_status2' ,  # updated age adjusted to landmark
#   'GENDER_NEW' ,
#   'INIT_BMI_NEW' ,
#   # 'ETHCAT_NEW' , 
#   'ETHCAT2_NEW' ,     # simplified ethnicity var
#   'BTYPE_REC' ,
#   # 'DIAB' , 
#   'DIAB_Y' ,   # yes/no diabetes indicator 
#   'VENTILATOR_TCR' ,
#   # 'PRIOR_CARD_SURG_TCR' , 
#   'PRIOR_CARD_SURG_TCR_Y' ,  # binary indicator
#   
#   'EXCEPTION', # <- exception variable!! comment out if not in baseline_variates!
#   
#   'ISCH_CARD_Y'
# )
```


These variables do have a group * EXCEPTION interaction term!

group*EXCEPTION = group:EXCEPTION + group + EXCEPTION

```{r}
adjusted_variables <- c(
   'EXCEPTION', # <- exception variable!! comment out if not in baseline_variates!
  'group:EXCEPTION',
  # 'INIT_AGE' , 
  'age_at_status2' ,  # updated age adjusted to landmark
  'GENDER_NEW' ,
  'INIT_BMI_NEW' ,
  # 'ETHCAT_NEW' , 
  'ETHCAT2_NEW' ,     # simplified ethnicity var
  'BTYPE_REC' ,
  # 'DIAB' , 
  'DIAB_Y' ,   # yes/no diabetes indicator 
  'VENTILATOR_TCR' ,
  # 'PRIOR_CARD_SURG_TCR' , 
  'PRIOR_CARD_SURG_TCR_Y' ,  # binary indicator
  'ISCH_CARD_Y'
)
```

The `competing_risks` variable is just a lookup table to loop each of the outcomes functionally

```{r}
competing_risks <- 
  data.frame(
    outcome = c(
      "Death / Deterioration",
      "Transplantation",
      "Upgrade to Status 1", 
      "Downgrade to Status 3-6",
      "Delisted"
    ), 
    code = c(1,2,4,5,6)
  )
```


In the `..._models` list variable, it'll look like: 

Unadjusted

-   [[1]]
-   ...
-   [[5]]

Adjusted

-   [[1]]
-   ...
-   [[5]]

The `[[]]` will contain the following outcomes in this order, just like how it is listed in the `competing_risks` variable:

-   [[1]] = Death/Deterioration
-   [[2]] = Transplantation
-   [[3]] = Upgrade to Status 1
-   [[4]] = Downgrade to Status 3-6
-   [[5]] = Delisted



# Cox Model Functions 

## Cox Formula 

```{r}
make_cox_formula <- 
  function(adjusted = F) { 
    
    if (adjusted == T) {
      as.formula(
        paste(
          "Surv(FTIME_NEW, cause) ~ group +",
          paste(adjusted_variables, collapse = "+")
        )
      )
    } else {
      Surv(FTIME_NEW, cause) ~ group * EXCEPTION
    }
  }
```

## Fit Cox Model

```{r}
fit_cox_model <- 
  function(data, event_code, adjusted = F) {
    
    outcome_name <- 
      competing_risks$outcome[
        competing_risks$code == event_code
      ]
    
    data <- data |> 
      mutate(
        cause = as.integer(FSTATUS_NEW == event_code)
      )
    
    formula <- make_cox_formula(adjusted)
    
    tryCatch(
      withCallingHandlers(
      
      
      coxph(
        formula, 
        data = data
      ),
      
      warning = function(w) {
        message(
          "Cox model ran into warning for outcome ",
          event_code,
          if (adjusted == T) " (adjusted)" else " (unadjusted)",
          ": ",
          conditionMessage(w)
        )
      }
      ),
      
      error = function(e) {
        message(
          "Cox model failed for outcome ",
          event_code,
          if (adjusted == T) " (adjusted)" else " (unadjusted)",
          ": ",
          conditionMessage(e)
          
        )
        NULL
      }
      
    )
  }

```

### Run Cox Models

```{r}
cox_models <- list(
  Unadjusted = 
    map(
      competing_risks$code,
      ~ fit_cox_model(
        data = baseline_variates_no_NA, 
        event_code = .x, 
        adjusted = F
      ) 
  ),
  Adjusted = map(
    competing_risks$code,
    ~ fit_cox_model(
      data = baseline_variates_no_NA,
      event_code = .x,
      adjusted = T
    )
  )
)
```

Death

```{r, eval = F}
summary(cox_models$Unadjusted[[1]]) 
summary(cox_models$Adjusted[[1]])
```
Transplantation 

```{r, eval = F}
summary(cox_models$Unadjusted[[2]])
summary(cox_models$Adjusted[[2]])
```


Upgraded to Status 1

```{r, eval = F}
summary(cox_models$Unadjusted[[3]])
summary(cox_models$Adjusted[[3]])
```

Downgraded to Status 3-6 

```{r, eval = F}
summary(cox_models$Unadjusted[[4]])
summary(cox_models$Adjusted[[4]])
```


Delisted 

```{r, eval = F}
summary(cox_models$Unadjusted[[5]])
summary(cox_models$Adjusted[[5]])
```

# Fine-Gray Model Functions

```{r}
fit_fg_model <- 
  function(data, event_code, adjusted = F) {
    
    tryCatch({
      
      if (adjusted) {
        
        covariate_matrix <- model.matrix(
          as.formula(
            paste(
              "~ group +",
              paste(adjusted_variables, collapse = " + ")
            )
          ),
          data = data
        )[, -1, drop = T]
        
      } else {
        
        covariate_matrix <- model.matrix(
          ~ group * EXCEPTION,
          data = data
        )[, -1, drop = F]
        
      }
      
      crr(
        ftime = data$FTIME_NEW,
        fstatus = data$FSTATUS_NEW,
        cov1 = covariate_matrix,
        failcode = event_code
      )
      
    }, error = function(e) {
      
      message(
        "Fine-Gray model failed for outcome",
        event_code,
        if (adjusted) " (adjusted)" else " (unadjusted)",
        ": ",
        e$message
      )
      
      NULL
    })
  }
```

## Run Fine-Gray Models

```{r}
fg_models <- 
  list(
    Unadjusted = 
      map(
        competing_risks$code,
        ~ fit_fg_model(
          data = baseline_variates_no_NA,
          event_code = .x,
          adjusted = F
        )
      ),
    Adjusted = 
      map(
        competing_risks$code,
        ~ fit_fg_model(
          data = baseline_variates_no_NA,
          event_code = .x,
          adjusted = T
        ) 
      )
  )
```

Summaries 

Death

```{r, eval = F}

summary(fg_models$Unadjusted[[1]])
summary(fg_models$Adjusted[[1]])

```

Transplantation

```{r, eval = F}
summary(fg_models$Unadjusted[[2]])
summary(fg_models$Adjusted[[2]])

```

Upgraded to Status 1

```{r, eval = F}

summary(fg_models$Unadjusted[[3]])
summary(fg_models$Adjusted[[3]])

```

Downgraded to Status 3-6

```{r, eval = F}
summary(fg_models$Unadjusted[[4]])
summary(fg_models$Adjusted[[4]])

```

Delisted

```{r, eval = F}
summary(fg_models$Unadjusted[[5]])
summary(fg_models$Adjusted[[5]])

```

# Direct Binomial Regression Model Functions

## DBR Formula

```{r}
make_binreg_formula <- 
  function(adjusted = F) { 
    
    if (adjusted == T) {
      as.formula(
        paste(
          "Event(FTIME_NEW, cause) ~ group +",
          paste(adjusted_variables, collapse = "+")
        )
      )
    } else {
      Event(FTIME_NEW, cause) ~ group * EXCEPTION
    }
  }
```

## Fit DBR Model

```{r}
fit_binreg_model <- 
  function(data, event_code, adjusted = F, time = 90) {
    
    outcome_name <- 
      competing_risks$outcome[
        competing_risks$code == event_code
      ]
    
    data <- data |> 
      mutate(
        cause = as.integer(FSTATUS_NEW == event_code)
      )
    
    formula <- make_binreg_formula(adjusted)
    
    tryCatch(
      withCallingHandlers(
        
      binreg(
        formula, 
        data = data,
        time = time
      ),
      
      warning = function(w) {
        message(
            "DBR model ran into warning for ",
            outcome_name,
            if (adjusted) " (adjusted)" else " (unadjusted)",
            ": ",
            conditionMessage(w)
        )
      }
      ),
      error = function(e) {
        message(
          "DBR model failed for outcome ",
          outcome_name,
          if (adjusted) "(adjusted)" else "(unadjusted)", 
          ": ",
          conditionMessage(e)
        )
        NULL
      }
    )
  }
```

### Rrun DBR Models

```{r}
binreg_models <- 
  list(
    Unadjusted = 
      map(
        competing_risks$code,
        ~ fit_binreg_model(
          data = baseline_variates_no_NA,
          event_code = .x,
          adjusted = F
        )
      ),
    Adjusted = 
      map(
        competing_risks$code,
        ~ fit_binreg_model(
          data = baseline_variates_no_NA,
          event_code = .x,
          adjusted = T
        )
      )
  )
```

Summaries 

```{r, eval = F}
summary(binreg_models$Unadjusted[[1]])
summary(binreg_models$Adjusted[[1]])
```

```{r, eval = F}
summary(binreg_models$Unadjusted[[2]])
summary(binreg_models$Adjusted[[2]])
```


```{r, eval = F}
summary(binreg_models$Unadjusted[[3]])
summary(binreg_models$Adjusted[[3]])
```

```{r, eval = F}
summary(binreg_models$Unadjusted[[4]])
summary(binreg_models$Adjusted[[4]])
```


```{r, eval = F}
summary(binreg_models$Unadjusted[[5]])
summary(binreg_models$Adjusted[[5]])
```

# Extracting information from models for Kable Tables



## Cox info extraction

```{r}
# Extract from coxph model

extract_cox <- function(model, outcome, adjustment) {
  
  # an error occurred with the model, have NA values
  # fill in the values
  if (is.null(model)) {
    return(
      data.frame(
        Model = "Cause-Specific Cox",
        Outcome = outcome,
        Adjustment = adjustment,
        Estimate = NA_real_,
        Lower = NA_real_,
        Upper = NA_real_,
        P = NA_real_
      )
    )
  }
  
  tryCatch({
    
    # if model exists extract its info
    
    s <- summary(model)
    
    data.frame(
      Model = "Cause-Specific Cox",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = s$conf.int[1, "exp(coef)"],
      Lower    = s$conf.int[1, "lower .95"],
      Upper    = s$conf.int[1, "upper .95"],
      P        = s$coefficients[1, "Pr(>|z|)"]
    )
    
    
     # if an error occurs while extract info from model,
     # make things NA
  }, error = function(e) {
    data.frame(
      Model = "Cause-Specific Cox",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = NA_real_,
      Lower = NA_real_,
      Upper = NA_real_,
      P = NA_real_
    )
  }) 
}
```


## Fine-Gray extraction

```{r}
extract_fg <- function(model, outcome, adjustment) {
  if (is.null(model)) {
    return(
      data.frame(
        Model = "Fine-Gray Subdistribution",
        Outcome = outcome,
        Adjustment = adjustment,
        Estimate = NA_real_,
        Lower = NA_real_,
        Upper = NA_real_,
        P = NA_real_
      )
    )
  }
  
  tryCatch({
    
    s <- summary(model)
    
    data.frame(
      Model = "Fine-Gray Subdistribution",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = s$conf.int[1, "exp(coef)"],
      Lower = s$conf.int[1, "2.5%"],
      Upper = s$conf.int[1, "97.5%"],
      P = s$coef[1, "p-value"]
    )
  }, error = function(e) {
    data.frame(
      Model = "Fine-Gray Subdistribution",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = NA_real_,
      Lower = NA_real_,
      Upper = NA_real_,
      P = NA_real_
    )
  })
}
```

## DBR extraction

```{r}
extract_binreg <- function(model, outcome, adjustment) {
  
  if (is.null(model)) {
    return(
      data.frame(
        Model = "Direct Binomial Regression",
        Outcome = outcome,
        Adjustment = adjustment,
        Estimate = NA_real_,
        Lower = NA_real_,
        Upper = NA_real_,
        P = NA_real_
      )
    )
  }
  
  tryCatch({
    
    s <- summary(model)
    
    data.frame(
      Model = "Direct Binomial Regression",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = s$exp[2, "Estimate"],
      Lower = s$exp[2, "2.5%"],
      Upper = s$exp[2, "97.5%"],
      P = s$coef[2, "P-value"]
    )
    
  }, error = function(e) {
    
    data.frame(
      Model = "Direct Binomial Regression",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = NA_real_,
      Lower = NA_real_,
      Upper = NA_real_,
      P = NA_real_
    )
  })
}
```

# Making the dataframes for the tables

`unadjusted_results` contains the unadjusted results ...

```{r}
unadjusted_results <- 
  bind_rows(
    
    map2_dfr(
      cox_models$Unadjusted,
      competing_risks$outcome,
      ~ extract_cox(.x, .y, "Unadjusted")
    ),
    
    map2_dfr(
      fg_models$Unadjusted,
      competing_risks$outcome,
      ~ extract_fg(.x, .y, "Unadjusted")
    ),
    
    map2_dfr(
      binreg_models$Unadjusted,
      competing_risks$outcome,
      ~extract_binreg(.x, .y, "Unadjusted")
    )
  )
```

`adjusted_results`

```{r}
adjusted_results <- 
  bind_rows(
    # Cox Model info
    map2_dfr(
      cox_models$Adjusted,
      competing_risks$outcome,
      ~ extract_cox(.x, .y, "Adjusted")
    ),
    # Fine-Gray Model info
    map2_dfr(
      fg_models$Adjusted,
      competing_risks$outcome,
      ~ extract_fg(.x, .y, "Adjusted")
    ),
    # DBR Model info
    map2_dfr(
      binreg_models$Adjusted,
      competing_risks$outcome,
      ~extract_binreg(.x, .y, "Adjusted")
    )
  )
```

# Kable Table Making

## Unadjusted Kable

```{r}
kable_table_unadjusted <- 
  unadjusted_results |>
  filter(Model != "Direct Binomial Regression") |> 
  mutate(
    Estimate = round(Estimate, 3),
    CI = paste0("(", round(Lower, 3), " , ", round(Upper, 3), ")"),
    P  = ifelse(P < 0.001, "p < 0.001", sprintf("%.3f", P))
  ) |>
  select(Outcome, Estimate, CI, P) |>
   kable(
    col.names = c("Outcome", "Estimate**", "95% CI", "p-value"),
    align = c("l", "c", "c", "c"),
                  
    caption = "<span style='font-size:25px; text-align:center; font-weight:bold; display:block;'>Table 4. Unadjusted Escalated Effect Estimates for Waitlist Outcomes at 90 Days</span>",
  ) |> 
  kable_styling(
    full_width = T,
    # position = 'left',
    font_size = 19
  ) |> 
  add_header_above(
    c("Models include Group*Exception interaction" = 4),
    bold = T,
    font_size = 15
  ) |> 
  column_spec(1, width = '10em', bold = T, color = "black") |> 
  column_spec(2:4, width = "7em", bold = F, color = "black") |> 
  
  pack_rows("Cause-Specific Cox Model (HR*)", 1, 5) |>
  
  pack_rows("Fine-Gray Subdistribution Model (SHR*)", 6, 10)
```

```{r, eval = F}
kable_table_unadjusted <- 
  unadjusted_results |>
  mutate(
    Estimate = round(Estimate, 3),
    CI = paste0("(", round(Lower, 3), " , ", round(Upper, 3), ")"),
    P  = ifelse(P < 0.001, "p < 0.001", sprintf("%.3f", P))
  ) |>
  select(Outcome, Estimate, CI, P) |>
   kable(
    col.names = c("Outcome", "Estimate**", "95% CI", "p-value"),
    align = c("l", "c", "c", "c"),
                  
    caption = "<span style='font-size:25px; text-align:center; font-weight:bold; display:block;'>Table 4. Unadjusted Escalated Effect Estimates for Waitlist Outcomes at 90 Days</span>",
  ) |> 
  kable_styling(
    full_width = T,
    # position = 'left',
    font_size = 19
  ) |> 
  add_header_above(
    c("Models include Group*Exception interaction" = 4),
    bold = T,
    font_size = 15
  ) |> 
  column_spec(1, width = '10em', bold = T, color = "black") |> 
  column_spec(2:4, width = "7em", bold = F, color = "black") |> 
  
  pack_rows("Cause-Specific Cox Model (HR*)", 1, 5) |>
  
  pack_rows("Fine-Gray Subdistribution Model (SHR*)", 6, 10) |> 
  pack_rows("Direct Binomial Regression Model (OR*)", 11, 15)
```

## Adjusted Kable

```{r}
kable_table_adjusted <- 
  adjusted_results |>
  filter(Model != "Direct Binomial Regression") |> 
  mutate(
    Estimate = round(Estimate, 3),
    CI = paste0("(", round(Lower, 3), " , ", round(Upper, 3), ")"),
    P  = ifelse(P < 0.001, "p < 0.001", sprintf("%.3f", P))
  ) |>
  select(Outcome, Estimate, CI, P) |>
  kable(
    col.names = c("Outcome", "Estimate**", "95% CI", "p-value"),
    align = c("l", "c", "c", "c"),
                  
    caption = "<span style='font-size:25px; text-align:center; font-weight:bold; display:block;'>Table 5. Adjusted Escalated Effect Estimates for Waitlist at Outcomes at 90 Days</span>",
  ) |>
  kable_styling(
    full_width = T,
    position = 'left',
    font_size = 19
  ) |> 
  add_header_above(
    c("Models include Group*Exception interaction" = 4),
    bold = T,
    font_size = 15
  ) |> 
  column_spec(1, bold = T, color = "black") |> 
  column_spec(2:4, bold = F, color = "black") |> 
  pack_rows("Cause-Specific Cox Model (HR*)", 1, 5) |>
  pack_rows("Fine-Gray Subdistribution Model (SHR*)", 6, 10)
```

```{r, eval = F}
kable_table_adjusted <- 
  adjusted_results |>
  mutate(
    Estimate = round(Estimate, 3),
    CI = paste0("(", round(Lower, 3), " , ", round(Upper, 3), ")"),
    P  = ifelse(P < 0.001, "p < 0.001", sprintf("%.3f", P))
  ) |>
  select(Outcome, Estimate, CI, P) |>
  kable(
    col.names = c("Outcome", "Estimate**", "95% CI", "p-value"),
    align = c("l", "c", "c", "c"),
                  
    caption = "<span style='font-size:25px; text-align:center; font-weight:bold; display:block;'>Table 5. Adjusted Escalated Effect Estimates for Waitlist at Outcomes at 90 Days</span>",
  ) |>
  kable_styling(
    full_width = T,
    position = 'left',
    font_size = 19
  ) |> 
  add_header_above(
    c("Models include Group*Exception interaction" = 4),
    bold = T,
    font_size = 15
  ) |> 
  column_spec(1, bold = T, color = "black") |> 
  column_spec(2:4, bold = F, color = "black") |> 
  pack_rows("Cause-Specific Cox Model (HR*)", 1, 5) |>
  
  pack_rows("Fine-Gray Subdistribution Model (SHR*)", 6, 10) |> 
  
  pack_rows("Direct Binomial Regression Model (OR*)", 11, 15) 
```

## Show and Save Kables

```{r}
kable_table_unadjusted
kable_table_adjusted
```


```{r}
Sys.setenv(
  CHROMOTE_CHROME =
    "C:/Program Files (x86)/Microsoft/Edge/Application/msedge.exe"
)
```



```{r}
save_kable(kable_table_unadjusted, file = "tables/unadjusted_results_table_incldues_interaction.png")
save_kable(kable_table_adjusted, file = "tables/adjusted_results_table_incldues_interaction.png")
```


# group*EXCEPTION interaction info extraction

# Unadjusted

```{r}
# Extract from coxph model

extract_cox <- function(model, outcome, adjustment) {
  
  # an error occurred with the model, have NA values
  # fill in the values
  if (is.null(model)) {
    return(
      data.frame(
        Model = "Cause-Specific Cox",
        Outcome = outcome,
        Adjustment = adjustment,
        Estimate = NA_real_,
        Lower = NA_real_,
        Upper = NA_real_,
        P = NA_real_
      )
    )
  }
  
  tryCatch({
    
    # if model exists extract its info
    
    s <- summary(model)
    
    data.frame(
      Model = "Cause-Specific Cox",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = s$conf.int[3, "exp(coef)"],
      Lower    = s$conf.int[3, "lower .95"],
      Upper    = s$conf.int[3, "upper .95"],
      P        = s$coefficients[3, "Pr(>|z|)"]
    )
    
    
     # if an error occurs while extract info from model,
     # make things NA
  }, error = function(e) {
    data.frame(
      Model = "Cause-Specific Cox",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = NA_real_,
      Lower = NA_real_,
      Upper = NA_real_,
      P = NA_real_
    )
  }) 
}
```


## Fine-Gray extraction

```{r}
extract_fg <- function(model, outcome, adjustment) {
  if (is.null(model)) {
    return(
      data.frame(
        Model = "Fine-Gray Subdistribution",
        Outcome = outcome,
        Adjustment = adjustment,
        Estimate = NA_real_,
        Lower = NA_real_,
        Upper = NA_real_,
        P = NA_real_
      )
    )
  }
  
  tryCatch({
    
    s <- summary(model)
    
    data.frame(
      Model = "Fine-Gray Subdistribution",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = s$conf.int[3, "exp(coef)"],
      Lower = s$conf.int[3, "2.5%"],
      Upper = s$conf.int[3, "97.5%"],
      P = s$coef[3, "p-value"]
    )
  }, error = function(e) {
    data.frame(
      Model = "Fine-Gray Subdistribution",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = NA_real_,
      Lower = NA_real_,
      Upper = NA_real_,
      P = NA_real_
    )
  })
}
```

## DBR extraction

```{r}
extract_binreg <- function(model, outcome, adjustment) {
  
  if (is.null(model)) {
    return(
      data.frame(
        Model = "Direct Binomial Regression",
        Outcome = outcome,
        Adjustment = adjustment,
        Estimate = NA_real_,
        Lower = NA_real_,
        Upper = NA_real_,
        P = NA_real_
      )
    )
  }
  
  tryCatch({
    
    s <- summary(model)
    
    data.frame(
      Model = "Direct Binomial Regression",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = s$exp[4, "Estimate"],
      Lower = s$exp[4, "2.5%"],
      Upper = s$exp[4, "97.5%"],
      P = s$coef[4, "P-value"]
    )
    
  }, error = function(e) {
    
    data.frame(
      Model = "Direct Binomial Regression",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = NA_real_,
      Lower = NA_real_,
      Upper = NA_real_,
      P = NA_real_
    )
  })
}
```

# Making the dataframes for the tables

`unadjusted_results` contains the unadjusted results ...

```{r}
unadjusted_results <- 
  bind_rows(
    
    map2_dfr(
      cox_models$Unadjusted,
      competing_risks$outcome,
      ~ extract_cox(.x, .y, "Unadjusted")
    ),
    
    map2_dfr(
      fg_models$Unadjusted,
      competing_risks$outcome,
      ~ extract_fg(.x, .y, "Unadjusted")
    ),
    
    map2_dfr(
      binreg_models$Unadjusted,
      competing_risks$outcome,
      ~extract_binreg(.x, .y, "Unadjusted")
    )
  )
```

# Kable Table Making

## Unadjusted Kable

```{r}
kable_table_unadjusted <- 
  unadjusted_results |>
  filter(Model != "Direct Binomial Regression") |> 
  mutate(
    Estimate = round(Estimate, 3),
    CI = paste0("(", round(Lower, 3), " , ", round(Upper, 3), ")"),
    P  = ifelse(P < 0.001, "p < 0.001", sprintf("%.3f", P))
  ) |>
  select(Outcome, Estimate, CI, P) |>
   kable(
    col.names = c("Outcome", "Estimate**", "95% CI", "p-value"),
    align = c("l", "c", "c", "c"),
                  
    caption = "<span style='font-size:25px; text-align:center; font-weight:bold; display:block;'>Table 6. Unadjusted Group*Exception Interaction Effect Estimates for Waitlist Outcomes at 90 Days</span>",
  ) |> 
  kable_styling(
    full_width = T,
    # position = 'left',
    font_size = 19
  ) |> 
  column_spec(1, width = '10em', bold = T, color = "black") |> 
  column_spec(2:4, width = "7em", bold = F, color = "black") |> 
  
  pack_rows("Cause-Specific Cox Model (HR*)", 1, 5) |>
  pack_rows("Fine-Gray Subdistribution Model (SHR*)", 6, 10) 
```

```{r, eval = F}
kable_table_unadjusted <- 
  unadjusted_results |>
  mutate(
    Estimate = round(Estimate, 3),
    CI = paste0("(", round(Lower, 3), " , ", round(Upper, 3), ")"),
    P  = ifelse(P < 0.001, "p < 0.001", sprintf("%.3f", P))
  ) |>
  select(Outcome, Estimate, CI, P) |>
   kable(
    col.names = c("Outcome", "Estimate**", "95% CI", "p-value"),
    align = c("l", "c", "c", "c"),
                  
    caption = "<span style='font-size:25px; text-align:center; font-weight:bold; display:block;'>Table 6. Unadjusted Group*Exception Interaction Effect Estimates for Waitlist Outcomes at 90 Days</span>",
  ) |> 
  kable_styling(
    full_width = T,
    # position = 'left',
    font_size = 19
  ) |> 
  column_spec(1, width = '10em', bold = T, color = "black") |> 
  column_spec(2:4, width = "7em", bold = F, color = "black") |> 
  
  pack_rows("Cause-Specific Cox Model (HR*)", 1, 5) |>
  
  pack_rows("Fine-Gray Subdistribution Model (SHR*)", 6, 10) |> 
  pack_rows("Direct Binomial Regression Model (OR*)", 11, 15)
```


# Adjusted

## Cox info extraction

```{r}
# Extract from coxph model

extract_cox <- function(model, outcome, adjustment) {
  
  # an error occurred with the model, have NA values
  # fill in the values
  if (is.null(model)) {
    return(
      data.frame(
        Model = "Cause-Specific Cox",
        Outcome = outcome,
        Adjustment = adjustment,
        Estimate = NA_real_,
        Lower = NA_real_,
        Upper = NA_real_,
        P = NA_real_
      )
    )
  }
  
  tryCatch({
    
    # if model exists extract its info
    
    s <- summary(model)
    
    data.frame(
      Model = "Cause-Specific Cox",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = s$conf.int[17, "exp(coef)"],
      Lower    = s$conf.int[17, "lower .95"],
      Upper    = s$conf.int[17, "upper .95"],
      P        = s$coefficients[17, "Pr(>|z|)"]
    )
    
    
     # if an error occurs while extract info from model,
     # make things NA
  }, error = function(e) {
    data.frame(
      Model = "Cause-Specific Cox",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = NA_real_,
      Lower = NA_real_,
      Upper = NA_real_,
      P = NA_real_
    )
  }) 
}
```


## Fine-Gray extraction

```{r}
extract_fg <- function(model, outcome, adjustment) {
  if (is.null(model)) {
    return(
      data.frame(
        Model = "Fine-Gray Subdistribution",
        Outcome = outcome,
        Adjustment = adjustment,
        Estimate = NA_real_,
        Lower = NA_real_,
        Upper = NA_real_,
        P = NA_real_
      )
    )
  }
  
  tryCatch({
    
    s <- summary(model)
    
    data.frame(
      Model = "Fine-Gray Subdistribution",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = s$conf.int[17, "exp(coef)"],
      Lower = s$conf.int[17, "2.5%"],
      Upper = s$conf.int[17, "97.5%"],
      P = s$coef[17, "p-value"]
    )
  }, error = function(e) {
    data.frame(
      Model = "Fine-Gray Subdistribution",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = NA_real_,
      Lower = NA_real_,
      Upper = NA_real_,
      P = NA_real_
    )
  })
}
```

## DBR extraction

```{r}
extract_binreg <- function(model, outcome, adjustment) {
  
  if (is.null(model)) {
    return(
      data.frame(
        Model = "Direct Binomial Regression",
        Outcome = outcome,
        Adjustment = adjustment,
        Estimate = NA_real_,
        Lower = NA_real_,
        Upper = NA_real_,
        P = NA_real_
      )
    )
  }
  
  tryCatch({
    
    s <- summary(model)
    
    data.frame(
      Model = "Direct Binomial Regression",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = s$exp[18, "Estimate"],
      Lower = s$exp[18, "2.5%"],
      Upper = s$exp[18, "97.5%"],
      P = s$coef[18, "P-value"]
    )
    
  }, error = function(e) {
    
    data.frame(
      Model = "Direct Binomial Regression",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = NA_real_,
      Lower = NA_real_,
      Upper = NA_real_,
      P = NA_real_
    )
  })
}
```

# Making the dataframes for the tables

`adjusted_results`

```{r}
adjusted_results <- 
  bind_rows(
    # Cox Model info
    map2_dfr(
      cox_models$Adjusted,
      competing_risks$outcome,
      ~ extract_cox(.x, .y, "Adjusted")
    ),
    # Fine-Gray Model info
    map2_dfr(
      fg_models$Adjusted,
      competing_risks$outcome,
      ~ extract_fg(.x, .y, "Adjusted")
    )
  )
```

```{r, eval = F}
adjusted_results <- 
  bind_rows(
    # Cox Model info
    map2_dfr(
      cox_models$Adjusted,
      competing_risks$outcome,
      ~ extract_cox(.x, .y, "Adjusted")
    ),
    # Fine-Gray Model info
    map2_dfr(
      fg_models$Adjusted,
      competing_risks$outcome,
      ~ extract_fg(.x, .y, "Adjusted")
    ),
    # DBR Model info
    map2_dfr(
      binreg_models$Adjusted,
      competing_risks$outcome,
      ~extract_binreg(.x, .y, "Adjusted")
    )
  )
```

# Kable Table Making

## Adjusted Kable


```{r}
kable_table_adjusted <- 
  adjusted_results |>
  filter(Model != "Direct Binomial Regression") |> 
  mutate(
    Estimate = round(Estimate, 3),
    CI = paste0("(", round(Lower, 3), " , ", round(Upper, 3), ")"),
    P  = ifelse(P < 0.001, "p < 0.001", sprintf("%.3f", P))
  ) |>
  select(Outcome, Estimate, CI, P) |>
  kable(
    col.names = c("Outcome", "Estimate**", "95% CI", "p-value"),
    align = c("l", "c", "c", "c"),
                  
    caption = "<span style='font-size:25px; text-align:center; font-weight:bold; display:block;'>Table 7. Adjusted Group*Exception Interaction Effect Estimates for Waitlist at Outcomes at 90 Days</span>",
  ) |>
  kable_styling(
    full_width = T,
    position = 'left',
    font_size = 19
  ) |> 
  column_spec(1, bold = T, color = "black") |> 
  column_spec(2:4, bold = F, color = "black") |> 
  pack_rows("Cause-Specific Cox Model (HR*)", 1, 5) |>
  pack_rows("Fine-Gray Subdistribution Model (SHR*)", 6, 10)  
```

```{r, eval = F}
kable_table_adjusted <- 
  adjusted_results |>
  mutate(
    Estimate = round(Estimate, 3),
    CI = paste0("(", round(Lower, 3), " , ", round(Upper, 3), ")"),
    P  = ifelse(P < 0.001, "p < 0.001", sprintf("%.3f", P))
  ) |>
  select(Outcome, Estimate, CI, P) |>
  kable(
    col.names = c("Outcome", "Estimate**", "95% CI", "p-value"),
    align = c("l", "c", "c", "c"),
                  
    caption = "<span style='font-size:25px; text-align:center; font-weight:bold; display:block;'>Table 7. Adjusted Group*Exception Interaction Effect Estimates for Waitlist at Outcomes at 90 Days</span>",
  ) |>
  kable_styling(
    full_width = T,
    position = 'left',
    font_size = 19
  ) |> 
  column_spec(1, bold = T, color = "black") |> 
  column_spec(2:4, bold = F, color = "black") |> 
  pack_rows("Cause-Specific Cox Model (HR*)", 1, 5) |>
  
  pack_rows("Fine-Gray Subdistribution Model (SHR*)", 6, 10) |>
  pack_rows("Direct Binomial Regression Model (OR*)", 11, 15) 
```

## Show and Save Kables

```{r}
kable_table_unadjusted
kable_table_adjusted
```

```{r}
save_kable(kable_table_unadjusted, file = "tables/unadjusted_group_exception_interaction_table.png")
save_kable(kable_table_adjusted, file = "tables/adjusted_group_exception_interaction_table.png")
```
