---
title: "Exception Status Exploration"
author: "Jesus Rodriguez" 
format: html
embed-resources: true
---

This .qmd file is to further check on INIT_STAT2\*exception interaction effects, it's similar to functional_models.qmd, but it has the formulas changed to include `INIT_STAT2*EXCEPTION_STAT2` 


## Libraries/Loading

```{r setup, message = F, warning = F, output = F}
library(tidyverse) # general tidying and manipulation
library(readxl) # reading xlsx files
library(kableExtra) # nice tables
library(lubridate) # easier working with dates
library(survival)
library(survminer)
library(table1)
library(cmprsk)
library(KMsurv)
library(mets)
library(writexl)

options(warn = 1)

# 1. Make sure the data you want to load in is loaded in the setup chunk.

# # Our cleaned data
# heart3 <-
#   read_csv("data/heart3.csv")

# Dr. Kawaguchi's cleaned data
analysis_data0 <-
  readRDS("data/analysis_data_082626.rds")
  
```

## Cleaning in `analysis_data`

This cleans the AGE_AT_STAT2, INIT_STAT2, FSTATUS codes, and 90 day censoring variables and rules in Dr. Kawaguchi's code

If using our `heart3.csv` data, this code will produce an error. 

```{r, error = T}
analysis_data <- 
  analysis_data0 |> 
  # We want to drop anyone that was STATUS 1 from our analytic cohort
  filter(INIT_STAT != 2110) |> 
  mutate(
    AGE_AT_STAT2 = 
      INIT_AGE + (as.numeric(STAT2_INIT - INIT_DATE)) / 365.25,
    
    INIT_STAT2 = 
      case_when(
        INIT_STAT == 2120 ~ 0, # Initial Status 2
        
        # Until Dr. K confirms with Dr. Wolfson,
        # we will keep candidates initally at STATUS 2999
        INIT_STAT %in% c(2130, 2140, 2150, 2160, 2999) ~ 1, # Escalated from 3-6
        TRUE ~ NA
      ),
    
    FSTATUS_NEW = 
      case_when(
        FTIME_NEW > 90 ~ 0,
        TRUE ~ FSTATUS_NEW
      ),
    FTIME_NEW = 
      case_when(
        FTIME_NEW > 90 ~ 90,
        TRUE ~ FTIME_NEW
      )
  )
```

Removing candidates that were initially STATUS 1 (2110) gives us 10174 observations

# Baseline Variates

```{r}
baseline_variates <- 
  # 2. In the Baseline Variates section, make sure the data you loaded in is being called in the `baseline_variates` pipeline
  
  analysis_data |> # change this dataset to the one you'd like to use
                   # heart3 or analysis_data
  
  
  select(WL_ID_CODE,
         INIT_AGE, 
         AGE_AT_STAT2, 
         GENDER_NEW,
         INIT_BMI_NEW,
         ETHCAT_NEW,
         ETHCAT2_NEW,
         BTYPE_REC,
         # DIAB,     
         DIAB_Y,   
         VENTILATOR_TCR,
         # PRIOR_CARD_SURG_TCR,
         PRIOR_CARD_SURG_TCR_Y,
         ISCH_CARD_Y,
         # DIAL_AFTER_LIST,
         # DIAL_AFTER_LIST_Y,
         
         INIT_STAT2, # stable / escalated variable
         
         EXCEPTION_STAT2, # <- <- <- exception variable
         
         FTIME_NEW, # time and event variables
         FSTATUS_NEW,         #
         END_DATE_NEW     #
  ) |> 
  mutate(
    across(
      c(INIT_AGE,
        AGE_AT_STAT2,
        INIT_BMI_NEW, 
        FTIME_NEW),
      as.numeric
    ),
    across(
      c(WL_ID_CODE,
        GENDER_NEW,
        ETHCAT_NEW,
        ETHCAT2_NEW,
        BTYPE_REC,
        # DIAB,
        DIAB_Y,
        VENTILATOR_TCR,
        # PRIOR_CARD_SURG_TCR,
        PRIOR_CARD_SURG_TCR_Y,
        ISCH_CARD_Y,
        # DIAL_AFTER_LIST,
        # DIAL_AFTER_LIST_Y,
        EXCEPTION_STAT2,
        FSTATUS_NEW),
      as.factor
    ),
    INIT_STAT2 = factor(INIT_STAT2,
                   levels = c(0,1),
                   labels = c("Initial Status 2", "Escalated from 3-6")),
    EXCEPTION_STAT2 = factor(EXCEPTION_STAT2,
                   levels = c(0,1),
                   labels = c("No Exception", "Exception")),
    ETHCAT2_NEW = factor(ETHCAT2_NEW,
                   levels = c("White", "Black", "Hispanic", "Asian", 'Other')),
    event_labels = factor(FSTATUS_NEW,     
                   levels = c(1,2,3,4,5,6,0),
                   labels = c("Death/Deterioration",
                              "Transplantation",
                              "Recovery",
                              "Upgraded to Status 1",
                              "Downgraded to Status 3-6",
                              "Delisted",
                              "Censored"))
  )
```

# Baseline Variates with No NA values in any rows

For the sake of simplicity, I am gonna have all the models run on complete data

This is just so each model is using the same data as the others

```{r}
baseline_variates_no_NA <- 
  baseline_variates |> 
  drop_na()
```

# Defining Variables for Models

The `adjusted_variables` are the variables that will be in the formula for the ADJUSTED Cox and Direct Binomial Regression models 

Their adjusted formulas look like this: 

```
"Surv(FTIME_NEW, cause) ~ INIT_STAT2 +",
          AGE_AT_STAT2 + GENDER_NEW + ...
```

While their unadjusted formulas won't need these variables: 

```
Surv(FTIME_NEW, cause) ~ INIT_STAT2,
```

```{r}
adjusted_variables <- c(
  # 'INIT_AGE' ,
  'EXCEPTION_STAT2', # <- exception variable!! comment out if not in baseline_variates!
  'INIT_STAT2:EXCEPTION_STAT2',
  
  
  'AGE_AT_STAT2' ,  # updated age adjusted to landmark
  'GENDER_NEW' ,
  'INIT_BMI_NEW' ,
  # 'ETHCAT_NEW' ,
  'ETHCAT2_NEW' ,     # simplified ethnicity var
  'BTYPE_REC' ,
  # 'DIAB' ,
  'DIAB_Y' ,   # yes/no diabetes indicator
  'VENTILATOR_TCR' ,
  # 'PRIOR_CARD_SURG_TCR' ,
  'PRIOR_CARD_SURG_TCR_Y' ,  # binary indicator 
  'ISCH_CARD_Y'
)
```

The `competing_risks` variable is just a lookup table to loop each of the outcomes functionally

```{r}
competing_risks <- 
  data.frame(
    outcome = c(
      "Death / Deterioration",
      "Transplantation",
      "Upgraded to Status 1", 
      "Downgraded to Status 3-6",
      "Delisted"
    ), 
    code = c(1,2,4,5,6)
  )
```


In the `..._models` list variable, it'll look like: 

Unadjusted

-   [[1]]
-   ...
-   [[5]]

Adjusted

-   [[1]]
-   ...
-   [[5]]

The `[[]]` will contain the following outcomes in this order, just like how it is listed in the `competing_risks` variable:

-   [[1]] = Death/Deterioration
-   [[2]] = Transplantation
-   [[3]] = Upgraded to Status 1
-   [[4]] = Downgraded to Status 3-6
-   [[5]] = Delisted

# Cox Model Functions 

## Cox Formula 

```{r}
make_cox_formula <- 
  function(adjusted = F) { 
    
    if (adjusted == T) {
      as.formula(
        paste(
          "Surv(FTIME_NEW, cause) ~ INIT_STAT2 +",
          paste(adjusted_variables, collapse = "+")
        )
      )
    } else {
      Surv(FTIME_NEW, cause) ~ INIT_STAT2 * EXCEPTION_STAT2
    }
  }
```

## Fit Cox Model

```{r}
fit_cox_model <- 
  function(data, event_code, adjusted = F) {
    
    outcome_name <- 
      competing_risks$outcome[
        competing_risks$code == event_code
      ]
    
    data <- data |> 
      mutate(
        cause = as.integer(FSTATUS_NEW == event_code)
      )
    
    formula <- make_cox_formula(adjusted)
    
    tryCatch(
      withCallingHandlers(
      
      
      coxph(
        formula, 
        data = data
      ),
      
      warning = function(w) {
        message(
          "Cox model ran into warning for outcome ",
          event_code,
          if (adjusted == T) " (adjusted)" else " (unadjusted)",
          ": ",
          conditionMessage(w)
        )
      }
      ),
      
      error = function(e) {
        message(
          "Cox model failed for outcome ",
          event_code,
          if (adjusted == T) " (adjusted)" else " (unadjusted)",
          ": ",
          conditionMessage(e)
          
        )
        NULL
      }
      
    )
  }

```

### Run Cox Models

```{r}
cox_models <- list(
  Unadjusted = 
    map(
      competing_risks$code,
      ~ fit_cox_model(
        data = baseline_variates, 
        event_code = .x, 
        adjusted = F
      ) 
  ),
  Adjusted = map(
    competing_risks$code,
    ~ fit_cox_model(
      data = baseline_variates,
      event_code = .x,
      adjusted = T
    )
  )
)
```

Death

```{r, eval = F}
summary(cox_models$Unadjusted[[1]]) 
summary(cox_models$Adjusted[[1]])
```
Transplantation 

```{r, eval = F}
summary(cox_models$Unadjusted[[2]])
summary(cox_models$Adjusted[[2]])
```


Upgraded to Status 1

```{r, eval = F}
summary(cox_models$Unadjusted[[3]])
summary(cox_models$Adjusted[[3]])
```

Downgraded to Status 3-6 

```{r, eval = F}
summary(cox_models$Unadjusted[[4]])
summary(cox_models$Adjusted[[4]])
```


Delisted 

```{r, eval = F}
summary(cox_models$Unadjusted[[5]])
summary(cox_models$Adjusted[[5]])
```

# Fine-Gray Model Functions

```{r}
fit_fg_model <- 
  function(data, event_code, adjusted = F) {
    
    tryCatch({
      
      if (adjusted) {
        
        covariate_matrix <- model.matrix(
          as.formula(
            paste(
              "~ INIT_STAT2 +",
              paste(adjusted_variables, collapse = " + ")
            )
          ),
          data = data
        )[, -1, drop = T]
        
      } else {
        
        covariate_matrix <- model.matrix(
          ~ INIT_STAT2 * EXCEPTION_STAT2,
          data = data
        )[, -1, drop = F]
        
      }
      
      crr(
        ftime = data$FTIME_NEW,
        fstatus = data$FSTATUS_NEW,
        cov1 = covariate_matrix,
        failcode = event_code
      )
      
    }, error = function(e) {
      
      message(
        "Fine-Gray model failed for outcome",
        event_code,
        if (adjusted) " (adjusted)" else " (unadjusted)",
        ": ",
        e$message
      )
      
      NULL
    })
  }
```

## Run Fine-Gray Models

```{r}
fg_models <- 
  list(
    Unadjusted = 
      map(
        competing_risks$code,
        ~ fit_fg_model(
          data = baseline_variates,
          event_code = .x,
          adjusted = F
        )
      ),
    Adjusted = 
      map(
        competing_risks$code,
        ~ fit_fg_model(
          data = baseline_variates_no_NA,
          event_code = .x,
          adjusted = T
        ) 
      )
  )
```

Summaries 

Death

```{r, eval = F}

summary(fg_models$Unadjusted[[1]])
summary(fg_models$Adjusted[[1]])

```

Transplantation

```{r, eval = F}
summary(fg_models$Unadjusted[[2]])
summary(fg_models$Adjusted[[2]])

```

Upgraded to Status 1

```{r, eval = F}

summary(fg_models$Unadjusted[[3]])
summary(fg_models$Adjusted[[3]])

```

Downgraded to Status 3-6

```{r, eval = F}
summary(fg_models$Unadjusted[[4]])
summary(fg_models$Adjusted[[4]])

```

Delisted

```{r, eval = F}
summary(fg_models$Unadjusted[[5]])
summary(fg_models$Adjusted[[5]])

```

# Direct Binomial Regression Model Functions

## DBR Formula

```{r,eval = F}
make_binreg_formula <- 
  function(adjusted = F) { 
    
    if (adjusted == T) {
      as.formula(
        paste(
          "Event(FTIME_NEW, cause) ~ INIT_STAT2 +",
          paste(adjusted_variables, collapse = "+")
        )
      )
    } else {
      Event(FTIME_NEW, cause) ~ INIT_STAT2 * EXCEPTION_STAT2
    }
  }
```

## Fit DBR Model

```{r, eval = F}
fit_binreg_model <- 
  function(data, event_code, adjusted = F, time = 90) {
    
    outcome_name <- 
      competing_risks$outcome[
        competing_risks$code == event_code
      ]
    
    data <- data |> 
      mutate(
        cause = as.integer(FSTATUS_NEW == event_code)
      )
    
    formula <- make_binreg_formula(adjusted)
    
    tryCatch(
      withCallingHandlers(
        
      binreg(
        formula, 
        data = data,
        time = time
      ),
      
      warning = function(w) {
        message(
            "DBR model ran into warning for ",
            outcome_name,
            if (adjusted) " (adjusted)" else " (unadjusted)",
            ": ",
            conditionMessage(w)
        )
      }
      ),
      error = function(e) {
        message(
          "DBR model failed for outcome ",
          outcome_name,
          if (adjusted) "(adjusted)" else "(unadjusted)", 
          ": ",
          conditionMessage(e)
        )
        NULL
      }
    )
  }
```

### Rrun DBR Models

```{r, eval = F}
binreg_models <- 
  list(
    Unadjusted = 
      map(
        competing_risks$code,
        ~ fit_binreg_model(
          data = baseline_variates_no_NA,
          event_code = .x,
          adjusted = F
        )
      ),
    Adjusted = 
      map(
        competing_risks$code,
        ~ fit_binreg_model(
          data = baseline_variates_no_NA,
          event_code = .x,
          adjusted = T
        )
      )
  )
```

Summaries 

```{r, eval = F}
summary(binreg_models$Unadjusted[[1]])
summary(binreg_models$Adjusted[[1]])
```

```{r, eval = F}
summary(binreg_models$Unadjusted[[2]])
summary(binreg_models$Adjusted[[2]])
```


```{r, eval = F}
summary(binreg_models$Unadjusted[[3]])
summary(binreg_models$Adjusted[[3]])
```

```{r, eval = F}
summary(binreg_models$Unadjusted[[4]])
summary(binreg_models$Adjusted[[4]])
```


```{r, eval = F}
summary(binreg_models$Unadjusted[[5]])
summary(binreg_models$Adjusted[[5]])
```



# Extracting information from models for Kable Tables

## Cox info extraction

```{r}
# Extract from coxph model

extract_cox <- function(model, outcome, adjustment) {
  
  # an error occurred with the model, have NA values
  # fill in the values
  if (is.null(model)) {
    return(
      data.frame(
        Model = "Cause-Specific Cox",
        Outcome = outcome,
        Adjustment = adjustment,
        Variable = NA,
        Estimate = NA,
        Lower = NA,
        Upper = NA,
        P = NA, 
        N = NA
      )
    )
  }
  
  tryCatch({
    
    # if model exists extract its info
    
    s <- summary(model)
    
    data.frame(
      Model = "Cause-Specific Cox",
      Outcome = outcome,
      Adjustment = adjustment,
      Variable = rownames(s$coefficients),
      Estimate = s$conf.int[, "exp(coef)"],
      Lower    = s$conf.int[, "lower .95"],
      Upper    = s$conf.int[, "upper .95"],
      P        = s$coefficients[, "Pr(>|z|)"],
      N        = model$n
    )
    
    
     # if an error occurs while extract info from model,
     # make things NA
  }, error = function(e) {
    data.frame(
      Model = "Cause-Specific Cox",
      Outcome = outcome,
      Adjustment = adjustment,
      Variable = NA,
      Estimate = NA,
      Lower = NA,
      Upper = NA,
      P = NA,
      N = NA
    )
  }) 
}
```


## Fine-Gray extraction

```{r}
extract_fg <- function(model, outcome, adjustment) {
  if (is.null(model)) {
    return(
      data.frame(
        Model = "Fine-Gray Subdistribution",
        Outcome = outcome,
        Adjustment = adjustment,
        Variable = NA,
        Estimate = NA,
        Lower = NA,
        Upper = NA,
        P = NA,
        N = NA
      )
    )
  }
  
  tryCatch({
    
    s <- summary(model)
    
    data.frame(
      Model = "Fine-Gray Subdistribution",
      Outcome = outcome,
      Adjustment = adjustment,
      Variable = rownames(s$coef),
      Estimate = s$conf.int[, "exp(coef)"],
      Lower = s$conf.int[, "2.5%"],
      Upper = s$conf.int[, "97.5%"],
      P = s$coef[, "p-value"],
      N = model$n
      
    )
  }, error = function(e) {
    data.frame(
      Model = "Fine-Gray Subdistribution",
      Outcome = outcome,
      Adjustment = adjustment,
      Variable = NA,
      Estimate = NA,
      Lower = NA,
      Upper = NA,
      P = NA,
      N = NA
    )
  })
}
```

## DBR extraction

```{r, eval = F}
extract_binreg <- function(model, outcome, adjustment) {
  
  if (is.null(model)) {
    return(
      data.frame(
        Model = "Direct Binomial Regression",
        Outcome = outcome,
        Adjustment = adjustment,
        Estimate = NA,
        Lower = NA,
        Upper = NA,
        P = NA
        N = NA
      )
    )
  }
  
  tryCatch({
    
    s <- summary(model)
    
    data.frame(
      Model = "Direct Binomial Regression",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = s$exp[2, "Estimate"],
      Lower = s$exp[2, "2.5%"],
      Upper = s$exp[2, "97.5%"],
      P = s$coef[2, "P-value"],
      N = model$n
    )
    
  }, error = function(e) {
    
    data.frame(
      Model = "Direct Binomial Regression",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = NA,
      Lower = NA,
      Upper = NA,
      P = NA,
      N = NA
    )
  })
}
```







# Making the dataframes for the tables

`unadjusted_results` contains the unadjusted results ...

```{r}
unadjusted_results <- 
  bind_rows(
    
    map2_dfr(
      cox_models$Unadjusted,
      competing_risks$outcome,
      ~ extract_cox(.x, .y, "Unadjusted")
    ),
    
    map2_dfr(
      fg_models$Unadjusted,
      competing_risks$outcome,
      ~ extract_fg(.x, .y, "Unadjusted")
    )
    
    # map2_dfr(
    #   binreg_models$Unadjusted,
    #   competing_risks$outcome,
    #   ~extract_binreg(.x, .y, "Unadjusted")
    # )
  )
```

`adjusted_results`

```{r}
adjusted_results <- 
  bind_rows(
    # Cox Model info
    map2_dfr(
      cox_models$Adjusted,
      competing_risks$outcome,
      ~ extract_cox(.x, .y, "Adjusted")
    ),
    # Fine-Gray Model info
    map2_dfr(
      fg_models$Adjusted,
      competing_risks$outcome,
      ~ extract_fg(.x, .y, "Adjusted")
    )
    # # DBR Model info
    # map2_dfr(
    #   binreg_models$Adjusted,
    #   competing_risks$outcome,
    #   ~extract_binreg(.x, .y, "Adjusted")
    # )
  )
```

INIT_STAT2*EXCEPTION_STAT2 interaction full results

```{r}
results_xlsx <-
  unadjusted_results |> 
  rbind(adjusted_results) |> 
  mutate(Estimate = round(Estimate, 3), 
         CI = paste0("(", round(Lower, 3), " , ", round(Upper, 3), ")"),
         P = ifelse(P < 0.001, "p < 0.001", round(P, 3))
  ) |> 
  select(-Lower, -Upper) |> 
  rename(HR = Estimate, `p-value` = P) |> 
  select(1:5, 8, 6:7)

rownames(results_xlsx) <- NULL

```

```{r}
write_xlsx(results_xlsx, "init_stat2xexception_results_082826.xlsx")
```

# Model Summary Extraction 

Not functionalized yet, very copy/paste heavy

You need to change the model list variable and number in summary(<cox_models>$<Adjusted[[<1>]]) to get the summary for other models

```{r}
model_summary <- summary(cox_models$Adjusted[[1]])

results <- as.data.frame(model_summary$coefficients)

results <- results |>
  rownames_to_column("Variable") |> 
  slice(
    1:2,       # keep first 2 rows
    n(),       # move last row to position 3
    3:(n()-1)  # then the remaining rows
  ) 

View(results)

# write_xlsx(results, "model_summary.xlsx")

```
