---
title: "Functionally Programming Fitting and Extracting Models"
author: "Jesus Rodriguez" 
format: html
embed-resources: true
---

The purpose of this quarto doc is to find a way to 'programmatically functionize' a way to run the models, know when there is an error with a model (like how we've had some outcomes with a variable where it is all NA), extract the information we need from the models and if an error occurred with the model then the info extract is just NA. 

All so that the kable tables created can be ran and if a model had issues, it will say NA on the table, so we know what outcome to take a deeper look into.

Hopefully this also makes the code more efficient because it takes a long time for my laptop to render the `baseline_variate_preparation.qmd` file

I have not yet updated any descriptive statistics tables here. 

## Libraries/Loading

```{r setup, message = F, warning = F, output = F}
library(tidyverse) # general tidying and manipulation
library(readxl) # reading xlsx files
library(kableExtra) # nice tables
library(lubridate) # easier working with dates
library(survival)
library(survminer)
library(table1)
library(cmprsk)
library(KMsurv)
library(mets)
library(writexl)

options(warn = 1)

# 1. Make sure the data you want to load in is loaded in the setup chunk.

# # Our cleaned data
# heart3 <-
#   read_csv("data/heart3.csv")

# Dr. Kawaguchi's cleaned data
analysis_data0 <-
  readRDS("data/analysis_data_082626.rds")
  
```

## Cleaning in `analysis_data`

This cleans the AGE_AT_STAT2, INIT_STAT2, FSTATUS codes, and 90 day censoring variables and rules in Dr. Kawaguchi's code

If using our `heart3.csv` data, this code will produce an error. 

```{r, error = T}
analysis_data <- 
  analysis_data0 |> 
  # We want to drop anyone that was STATUS 1 from our analytic cohort
  filter(INIT_STAT != 2110) |> 
  mutate(
    AGE_AT_STAT2 = 
      INIT_AGE + (as.numeric(STAT2_INIT - INIT_DATE)) / 365.25,
    
    INIT_STAT2 = 
      case_when(
        INIT_STAT == 2120 ~ 0, # Initial Status 2
        
        # Until Dr. K confirms with Dr. Wolfson,
        # we will keep candidates initally at STATUS 2999
        INIT_STAT %in% c(2130, 2140, 2150, 2160, 2999) ~ 1, # Escalated from 3-6
        TRUE ~ NA
      ),
    
    FSTATUS_NEW = 
      case_when(
        FTIME_NEW > 90 ~ 0,
        TRUE ~ FSTATUS_NEW
      ),
    FTIME_NEW = 
      case_when(
        FTIME_NEW > 90 ~ 90,
        TRUE ~ FTIME_NEW
      )
  )
```

Removing candidates that were initially STATUS 1 (2110) gives us 10174 observations

# Baseline Variates

```{r}
baseline_variates <- 
  # 2. In the Baseline Variates section, make sure the data you loaded in is being called in the `baseline_variates` pipeline
  
  analysis_data |> # change this dataset to the one you'd like to use
                   # heart3 or analysis_data
  
  
  select(WL_ID_CODE,
         INIT_AGE, 
         AGE_AT_STAT2, 
         GENDER_NEW,
         INIT_BMI_NEW,
         ETHCAT_NEW,
         ETHCAT2_NEW,
         BTYPE_REC,
         # DIAB,     
         DIAB_Y,   
         VENTILATOR_TCR,
         # PRIOR_CARD_SURG_TCR,
         PRIOR_CARD_SURG_TCR_Y,
         ISCH_CARD_Y,
         # DIAL_AFTER_LIST,
         # DIAL_AFTER_LIST_Y,
         
         INIT_STAT2, # stable / escalated variable
         
         EXCEPTION_STAT2, # <- <- <- exception variable
         
         FTIME_NEW, # time and event variables
         FSTATUS_NEW,         #
         END_DATE_NEW     #
  ) |> 
  mutate(
    across(
      c(INIT_AGE,
        AGE_AT_STAT2,
        INIT_BMI_NEW, 
        FTIME_NEW),
      as.numeric
    ),
    across(
      c(WL_ID_CODE,
        GENDER_NEW,
        ETHCAT_NEW,
        ETHCAT2_NEW,
        BTYPE_REC,
        # DIAB,
        DIAB_Y,
        VENTILATOR_TCR,
        # PRIOR_CARD_SURG_TCR,
        PRIOR_CARD_SURG_TCR_Y,
        ISCH_CARD_Y,
        # DIAL_AFTER_LIST,
        # DIAL_AFTER_LIST_Y,
        EXCEPTION_STAT2,
        FSTATUS_NEW),
      as.factor
    ),
    INIT_STAT2 = factor(INIT_STAT2,
                   levels = c(0,1),
                   labels = c("Initial Status 2", "Escalated from 3-6")),
    EXCEPTION_STAT2 = factor(EXCEPTION_STAT2,
                   levels = c(0,1),
                   labels = c("No Exception", "Exception")),
    ETHCAT2_NEW = factor(ETHCAT2_NEW,
                   levels = c("White", "Black", "Hispanic", "Asian", 'Other')),
    event_labels = factor(FSTATUS_NEW,     
                   levels = c(1,2,3,4,5,6,0),
                   labels = c("Death/Deterioration",
                              "Transplantation",
                              "Recovery",
                              "Upgraded to Status 1",
                              "Downgraded to Status 3-6",
                              "Delisted",
                              "Censored"))
  )
```

# Baseline Variates with No NA values in any rows

For the sake of simplicity, I am gonna have all the models run on complete data

This is just so each model is using the same data as the others

```{r}
baseline_variates_no_NA <- 
  baseline_variates |> 
  drop_na()
```

# Defining Variables for Models

The `adjusted_variables` are the variables that will be in the formula for the ADJUSTED Cox and Direct Binomial Regression models 

Their adjusted formulas look like this: 

```
"Surv(FTIME_NEW, cause) ~ INIT_STAT2 +",
          AGE_AT_STAT2 + GENDER_NEW + ...
```

While their unadjusted formulas won't need these variables: 

```
Surv(FTIME_NEW, cause) ~ INIT_STAT2,
```

```{r}
adjusted_variables <- c(
  # 'INIT_AGE' ,
  'EXCEPTION_STAT2', # <- exception variable!! comment out if not in baseline_variates!
  
  'AGE_AT_STAT2' ,  # updated age adjusted to landmark
  'GENDER_NEW' ,
  'INIT_BMI_NEW' ,
  # 'ETHCAT_NEW' ,
  'ETHCAT2_NEW' ,     # simplified ethnicity var
  'BTYPE_REC' ,
  # 'DIAB' ,
  'DIAB_Y' ,   # yes/no diabetes indicator
  'VENTILATOR_TCR' ,
  # 'PRIOR_CARD_SURG_TCR' ,
  'PRIOR_CARD_SURG_TCR_Y' ,  # binary indicator 
  'ISCH_CARD_Y'
)
```

The `competing_risks` variable is just a lookup table to loop each of the outcomes functionally

```{r}
competing_risks <- 
  data.frame(
    outcome = c(
      "Death / Deterioration",
      "Transplantation",
      "Upgraded to Status 1", 
      "Downgraded to Status 3-6",
      "Delisted"
    ), 
    code = c(1,2,4,5,6)
  )
```


In the `..._models` list variable, it'll look like: 

Unadjusted

-   [[1]]
-   ...
-   [[5]]

Adjusted

-   [[1]]
-   ...
-   [[5]]

The `[[]]` will contain the following outcomes in this order, just like how it is listed in the `competing_risks` variable:

-   [[1]] = Death/Deterioration
-   [[2]] = Transplantation
-   [[3]] = Upgraded to Status 1
-   [[4]] = Downgraded to Status 3-6
-   [[5]] = Delisted



# Cox Model Functions 

## Cox Formula 

```{r}
make_cox_formula <- 
  function(adjusted = F) { 
    
    if (adjusted == T) {
      as.formula(
        paste(
          "Surv(FTIME_NEW, cause) ~ INIT_STAT2 +",
          paste(adjusted_variables, collapse = "+")
        )
      )
    } else {
      Surv(FTIME_NEW, cause) ~ INIT_STAT2
    }
  }
```

## Fit Cox Model

```{r}
fit_cox_model <- 
  function(data, event_code, adjusted = F) {
    
    outcome_name <- 
      competing_risks$outcome[
        competing_risks$code == event_code
      ]
    
    data <- data |> 
      mutate(
        cause = as.integer(FSTATUS_NEW == event_code)
      )
    
    formula <- make_cox_formula(adjusted)
    
    tryCatch(
      withCallingHandlers(
      
      
      coxph(
        formula, 
        data = data
      ),
      
      warning = function(w) {
        message(
          "Cox model ran into warning for outcome ",
          event_code,
          if (adjusted == T) " (adjusted)" else " (unadjusted)",
          ": ",
          conditionMessage(w)
        )
      }
      ),
      
      error = function(e) {
        message(
          "Cox model failed for outcome ",
          event_code,
          if (adjusted == T) " (adjusted)" else " (unadjusted)",
          ": ",
          conditionMessage(e)
          
        )
        NULL
      }
      
    )
  }
```

### Run Cox Models

```{r}
cox_models <- list(
  Unadjusted = 
    map(
      competing_risks$code,
      ~ fit_cox_model(
        data = baseline_variates, 
        event_code = .x, 
        adjusted = F
      ) 
  ),
  Adjusted = map(
    competing_risks$code,
    ~ fit_cox_model(
      data = baseline_variates,
      event_code = .x,
      adjusted = T
    )
  )
)
```

Death

```{r, eval = F}
summary(cox_models$Unadjusted[[1]]) 
summary(cox_models$Adjusted[[1]])


```
Transplantation 

```{r, eval = F}
summary(cox_models$Unadjusted[[2]])
summary(cox_models$Adjusted[[2]])
```


Upgraded to Status 1

```{r, eval = F}
summary(cox_models$Unadjusted[[3]])
summary(cox_models$Adjusted[[3]])
```

Downgraded to Status 3-6 

```{r, eval = F}
summary(cox_models$Unadjusted[[4]])
summary(cox_models$Adjusted[[4]])
```


Delisted 

```{r, eval = F}
summary(cox_models$Unadjusted[[5]])
summary(cox_models$Adjusted[[5]])
```




# Fine-Gray Model Functions

```{r}
fit_fg_model <- 
  function(data, event_code, adjusted = F) {
    
    tryCatch({
      
      if (adjusted) {
        
        covariate_matrix <- model.matrix(
          as.formula(
            paste(
              "~ INIT_STAT2 +",
              paste(adjusted_variables, collapse = " + ")
            )
          ),
          data = data
        )[, -1, drop = T]
        
      } else {
        
        covariate_matrix <- model.matrix(
          ~ INIT_STAT2,
          data = data
        )[, -1, drop = F]
        
      }
      
      crr(
        ftime = data$FTIME_NEW,
        fstatus = data$FSTATUS_NEW,
        cov1 = covariate_matrix,
        failcode = event_code
      )
      
    }, error = function(e) {
      
      message(
        "Fine-Gray model failed for outcome",
        event_code,
        if (adjusted) " (adjusted)" else " (unadjusted)",
        ": ",
        e$message
      )
      
      NULL
    })
  }
```

## Run Fine-Gray Models

```{r}
fg_models <- 
  list(
    Unadjusted = 
      map(
        competing_risks$code,
        ~ fit_fg_model(
          data = baseline_variates,
          event_code = .x,
          adjusted = F
        )
      ),
    Adjusted = 
      map(
        competing_risks$code,
        ~ fit_fg_model(
          data = baseline_variates_no_NA,
          event_code = .x,
          adjusted = T
        ) 
      )
  )
```

Summaries 

Death

```{r, eval = F}
summary(fg_models$Unadjusted[[1]])
summary(fg_models$Adjusted[[1]])
```

Transplantation

```{r, eval = F}
summary(fg_models$Unadjusted[[2]])
summary(fg_models$Adjusted[[2]])
```

Upgraded to Status 1

```{r, eval = F}

summary(fg_models$Unadjusted[[3]])
summary(fg_models$Adjusted[[3]])

```

Downgraded to Status 3-6

```{r, eval = F}
summary(fg_models$Unadjusted[[4]])
summary(fg_models$Adjusted[[4]])

```

Delisted

```{r, eval = F}
summary(fg_models$Unadjusted[[5]])
summary(fg_models$Adjusted[[5]])

```

# Direct Binomial Regression Model Functions

## DBR Formula

```{r, eval = F }
make_binreg_formula <- 
  function(adjusted = F) { 
    
    if (adjusted == T) {
      as.formula(
        paste(
          "Event(FTIME_NEW, cause) ~ INIT_STAT2 +",
          paste(adjusted_variables, collapse = "+")
        )
      )
    } else {
      Event(FTIME_NEW, cause) ~ INIT_STAT2
    }
  }
```

## Fit DBR Model

```{r, eval = F}
fit_binreg_model <- 
  function(data, event_code, adjusted = F, time = 90) {
    
    outcome_name <- 
      competing_risks$outcome[
        competing_risks$code == event_code
      ]
    
    data <- data |> 
      mutate(
        cause = as.integer(FSTATUS_NEW == event_code)
      )
    
    formula <- make_binreg_formula(adjusted)
    
    tryCatch(
      withCallingHandlers(
        
      binreg(
        formula, 
        data = data,
        time = time
      ),
      
      warning = function(w) {
        message(
            "DBR model ran into warning for ",
            outcome_name,
            if (adjusted) " (adjusted)" else " (unadjusted)",
            ": ",
            conditionMessage(w)
        )
      }
      ),
      error = function(e) {
        message(
          "DBR model failed for outcome ",
          outcome_name,
          if (adjusted) "(adjusted)" else "(unadjusted)", 
          ": ",
          conditionMessage(e)
        )
        NULL
      }
    )
  }
```

### Rrun DBR Models

```{r, eval = F}
binreg_models <- 
  list(
    Unadjusted = 
      map(
        competing_risks$code,
        ~ fit_binreg_model(
          data = baseline_variates_no_NA,
          event_code = .x,
          adjusted = F
        )
      ),
    Adjusted = 
      map(
        competing_risks$code,
        ~ fit_binreg_model(
          data = baseline_variates_no_NA,
          event_code = .x,
          adjusted = T
        )
      )
  )
```

Summaries 

```{r, eval = F}
summary(binreg_models$Unadjusted[[1]])
summary(binreg_models$Adjusted[[1]])
```


```{r, eval = F}
summary(binreg_models$Unadjusted[[2]])
summary(binreg_models$Adjusted[[2]])
```


```{r, eval = F}
summary(binreg_models$Unadjusted[[3]])
summary(binreg_models$Adjusted[[3]])
```

```{r, eval = F}
summary(binreg_models$Unadjusted[[4]])
summary(binreg_models$Adjusted[[4]])
```


```{r, eval = F}
summary(binreg_models$Unadjusted[[5]])
summary(binreg_models$Adjusted[[5]])
```





# Extracting information from models for Kable Tables

## Cox info extraction

```{r}
# Extract from coxph model

extract_cox <- function(model, outcome, adjustment) {
  
  # an error occurred with the model, have NA values
  # fill in the values
  if (is.null(model)) {
    return(
      data.frame(
        Model = "Cause-Specific Cox",
        Outcome = outcome,
        Adjustment = adjustment,
        Variable = NA,
        Estimate = NA,
        Lower = NA,
        Upper = NA,
        P = NA, 
        N = NA
      )
    )
  }
  
  tryCatch({
    
    # if model exists extract its info
    
    s <- summary(model)
    
    data.frame(
      Model = "Cause-Specific Cox",
      Outcome = outcome,
      Adjustment = adjustment,
      Variable = rownames(s$coefficients),
      Estimate = s$conf.int[, "exp(coef)"],
      Lower    = s$conf.int[, "lower .95"],
      Upper    = s$conf.int[, "upper .95"],
      P        = s$coefficients[, "Pr(>|z|)"],
      N        = model$n
    )
    
    
     # if an error occurs while extract info from model,
     # make things NA
  }, error = function(e) {
    data.frame(
      Model = "Cause-Specific Cox",
      Outcome = outcome,
      Adjustment = adjustment,
      Variable = NA,
      Estimate = NA,
      Lower = NA,
      Upper = NA,
      P = NA,
      N = NA
    )
  }) 
}
```


## Fine-Gray extraction

```{r}
extract_fg <- function(model, outcome, adjustment) {
  if (is.null(model)) {
    return(
      data.frame(
        Model = "Fine-Gray Subdistribution",
        Outcome = outcome,
        Adjustment = adjustment,
        Variable = NA,
        Estimate = NA,
        Lower = NA,
        Upper = NA,
        P = NA,
        N = NA
      )
    )
  }
  
  tryCatch({
    
    s <- summary(model)
    
    data.frame(
      Model = "Fine-Gray Subdistribution",
      Outcome = outcome,
      Adjustment = adjustment,
      Variable = rownames(s$coef),
      Estimate = s$conf.int[, "exp(coef)"],
      Lower = s$conf.int[, "2.5%"],
      Upper = s$conf.int[, "97.5%"],
      P = s$coef[, "p-value"],
      N = model$n
      
    )
  }, error = function(e) {
    data.frame(
      Model = "Fine-Gray Subdistribution",
      Outcome = outcome,
      Adjustment = adjustment,
      Variable = NA,
      Estimate = NA,
      Lower = NA,
      Upper = NA,
      P = NA,
      N = NA
    )
  })
}
```

## DBR extraction

```{r, eval = F}
extract_binreg <- function(model, outcome, adjustment) {
  
  if (is.null(model)) {
    return(
      data.frame(
        Model = "Direct Binomial Regression",
        Outcome = outcome,
        Adjustment = adjustment,
        Estimate = NA,
        Lower = NA,
        Upper = NA,
        P = NA
        N = NA
      )
    )
  }
  
  tryCatch({
    
    s <- summary(model)
    
    data.frame(
      Model = "Direct Binomial Regression",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = s$exp[2, "Estimate"],
      Lower = s$exp[2, "2.5%"],
      Upper = s$exp[2, "97.5%"],
      P = s$coef[2, "P-value"],
      N = model$n
    )
    
  }, error = function(e) {
    
    data.frame(
      Model = "Direct Binomial Regression",
      Outcome = outcome,
      Adjustment = adjustment,
      Estimate = NA,
      Lower = NA,
      Upper = NA,
      P = NA,
      N = NA
    )
  })
}
```





# Making the dataframes for the tables

`unadjusted_results` contains the unadjusted results ...

```{r}
unadjusted_results <- 
  bind_rows(
    
    map2_dfr(
      cox_models$Unadjusted,
      competing_risks$outcome,
      ~ extract_cox(.x, .y, "Unadjusted")
    ),
    
    map2_dfr(
      fg_models$Unadjusted,
      competing_risks$outcome,
      ~ extract_fg(.x, .y, "Unadjusted")
    )
    
    # map2_dfr(
    #   binreg_models$Unadjusted,
    #   competing_risks$outcome,
    #   ~extract_binreg(.x, .y, "Unadjusted")
    # )
  )
```

`adjusted_results`

```{r}
adjusted_results <- 
  bind_rows(
    # Cox Model info
    map2_dfr(
      cox_models$Adjusted,
      competing_risks$outcome,
      ~ extract_cox(.x, .y, "Adjusted")
    ),
    # Fine-Gray Model info
    map2_dfr(
      fg_models$Adjusted,
      competing_risks$outcome,
      ~ extract_fg(.x, .y, "Adjusted")
    )
    # # DBR Model info
    # map2_dfr(
    #   binreg_models$Adjusted,
    #   competing_risks$outcome,
    #   ~extract_binreg(.x, .y, "Adjusted")
    # )
  )
```

# Kable Table Making

I have code here for four tables:

The first unadjusted table code and adjusted table code makes a table that does not include the Direct Binomial Regression results

The second set of codes do include the Direct Binomial Results, but its code chunk is set to `eval = F` so it shouldn't be ran when rendering

## Unadjusted Kable

```{r, eval = F}
kable_table_unadjusted <-
  unadjusted_results |>
  # filter(Model != "Direct Binomial Regression") |>
  mutate(
    Estimate = round(Estimate, 3),
    CI = paste0("(", round(Lower, 3), " , ", round(Upper, 3), ")"),
    P  = ifelse(P < 0.001, "p < 0.001", sprintf("%.3f", P))
  ) |>
  select(Outcome, Estimate, CI, P) |>
   kable(
    col.names = c("Outcome", "Estimate**", "95% CI", "p-value"),
    align = c("l", "c", "c", "c"),

    caption = "<span style='font-size:25px; text-align:center; font-weight:bold; display:block;'>Table 2. Unadjusted Escalated Effect Estimates for Waitlist Outcomes at 90 Days</span>",
  ) |>
  kable_styling(
    full_width = T,
    # position = 'left',
    font_size = 19
  ) |>
  #  add_header_above(
  #   c("Models do not include Group*Exception interaction" = 4),
  #   bold = T,
  #   font_size = 15
  # ) |>
  column_spec(1, width = '10em', bold = T, color = "black") |>
  column_spec(2:4, width = "7em", bold = F, color = "black") |>

  pack_rows("Cause-Specific Cox Model (HR*)", 1, 5) |>
  pack_rows("Fine-Gray Subdistribution Model (SHR*)", 6, 10)
```

```{r, eval = F}
kable_table_unadjusted <-
  unadjusted_results |>
  mutate(
    Estimate = round(Estimate, 3),
    CI = paste0("(", round(Lower, 3), " , ", round(Upper, 3), ")"),
    P  = ifelse(P < 0.001, "p < 0.001", sprintf("%.3f", P))
  ) |>
  select(Outcome, Estimate, CI, P) |>
   kable(
    col.names = c("Outcome", "Estimate**", "95% CI", "p-value"),
    align = c("l", "c", "c", "c"),

    caption = "<span style='font-size:25px; text-align:center; font-weight:bold; display:block;'>Table 2. Unadjusted Escalated Effect Estimates for Waitlist Outcomes at 90 Days</span>",
  ) |>
  kable_styling(
    full_width = T,
    # position = 'left',
    font_size = 19
  ) |>
   add_header_above(
    c("Models do not include Group*Exception interaction" = 4),
    bold = T,
    font_size = 15
  ) |>
  column_spec(1, width = '10em', bold = T, color = "black") |>
  column_spec(2:4, width = "7em", bold = F, color = "black") |>

  pack_rows("Cause-Specific Cox Model (HR*)", 1, 5) |>

  pack_rows("Fine-Gray Subdistribution Model (SHR*)", 6, 10) |>
  pack_rows("Direct Binomial Regression Model (OR*)", 11, 15)
```

## Adjusted Kable

```{r, eval = F}
kable_table_adjusted <-
  adjusted_results |>
  # filter(Model != "Direct Binomial Regression") |>
  mutate(
    Estimate = round(Estimate, 3),
    CI = paste0("(", round(Lower, 3), " , ", round(Upper, 3), ")"),
    P  = ifelse(P < 0.001, "p < 0.001", sprintf("%.3f", P))
  ) |>
  select(Outcome, Estimate, CI, P) |>
  kable(
    col.names = c("Outcome", "Estimate**", "95% CI", "p-value"),
    align = c("l", "c", "c", "c"),

    caption = "<span style='font-size:25px; text-align:center; font-weight:bold; display:block;'>Table 3. Adjusted Escalated Effect Estimates for Waitlist at Outcomes at 90 Days</span>",
  ) |>
  kable_styling(
    full_width = T,
    position = 'left',
    font_size = 19
  ) |>
  #  add_header_above(
  #   c("Models do not include Group*Exception interaction" = 4),
  #   bold = T,
  #   font_size = 15
  # ) |>
  column_spec(1, bold = T, color = "black") |>
  column_spec(2:4, bold = F, color = "black") |>
  pack_rows("Cause-Specific Cox Model (HR*)", 1, 5) |>
  pack_rows("Fine-Gray Subdistribution Model (SHR*)", 6, 10)
```

```{r, eval = F}
kable_table_adjusted <-
  adjusted_results |>
  mutate(
    Estimate = round(Estimate, 3),
    CI = paste0("(", round(Lower, 3), " , ", round(Upper, 3), ")"),
    P  = ifelse(P < 0.001, "p < 0.001", sprintf("%.3f", P))
  ) |>
  select(Outcome, Estimate, CI, P) |>
  kable(
    col.names = c("Outcome", "Estimate**", "95% CI", "p-value"),
    align = c("l", "c", "c", "c"),

    caption = "<span style='font-size:25px; text-align:center; font-weight:bold; display:block;'>Table 3. Adjusted Escalated Effect Estimates for Waitlist at Outcomes at 90 Days</span>",
  ) |>
  kable_styling(
    full_width = T,
    position = 'left',
    font_size = 19
  ) |>
   add_header_above(
    c("Models do not include Group*Exception interaction" = 4),
    bold = T,
    font_size = 15
  ) |>
  column_spec(1, bold = T, color = "black") |>
  column_spec(2:4, bold = F, color = "black") |>
  pack_rows("Cause-Specific Cox Model (HR*)", 1, 5) |>

  pack_rows("Fine-Gray Subdistribution Model (SHR*)", 6, 10) |>

  pack_rows("Direct Binomial Regression Model (OR*)", 11, 15)
```

## Show and Save Kables

```{r, eval = F}
kable_table_unadjusted
kable_table_adjusted
```


```{r, eval = F}
Sys.setenv(
  CHROMOTE_CHROME =
    "C:/Program Files (x86)/Microsoft/Edge/Application/msedge.exe"
)
```



```{r, eval = F}
save_kable(kable_table_unadjusted, file = "tables/new_unadjusted_results_table_DRAFT.png")
save_kable(kable_table_adjusted, file = "tables/new_adjusted_results_table_DRAFT.png")
```




# Table Making

```{r}
results_xlsx <-
  unadjusted_results |> 
  rbind(adjusted_results) |> 
  mutate(Estimate = round(Estimate, 3), 
         CI = paste0("(", round(Lower, 3), " , ", round(Upper, 3), ")"),
         P = ifelse(P < 0.001, "p < 0.001", round(P, 3))
  ) |> 
  select(-Lower, -Upper) |> 
  rename(HR = Estimate, `p-value` = P) |> 
  select(1:5, 8, 6:7)

rownames(results_xlsx) <- NULL
```

```{r}
write_xlsx(results_xlsx, "init_stat2_results_082826.xlsx")
```

# Model Summary Extraction

Not functionalized yet, very copy/paste heavy

You need to change the model list variable and number in summary(<cox_models>$<Adjusted[[<1>]]) to get the summary for other models

```{r}
model_summary <- summary(cox_models$Adjusted[[1]])

results <- as.data.frame(model_summary$coefficients)

results <- results |>
  rownames_to_column("Variable")

# write_xlsx(results, "model_summary.xlsx")

```




