<<<<<<< HEAD
---
title: "CIF Plot"
author: "Sydney Shimizu"
format: html
editor: visual
---

## CIF Plotting

```{r}

heart2<-read.csv("C:/Users/12063/Downloads/clean_heart_data_long.csv")

# install.packages("survminer") #

library(cmprsk)
library(survminer)
library(dplyr)
```

```{r}

# Compare Cumulative Incidence of Death/Deterioration and Cumulative Incidence of Heart Transplantation

heart2 <- heart2 %>%
  mutate(
    EVENT_TYPE = case_when(
      # Death / deterioration
      REM_CD_NEW %in% c(5,8,13) ~ 1,
      
      # Transplanted (deceased or living donor)
      REM_CD_NEW %in% c(2,3,4,14,21,22) ~ 2, 
      
      # Censor everyone else
      TRUE ~ 0
    ),
    
# Creating TIME_FROM_LANDMARK, TIME_30, and EVENT_30 Variables

TIME_FROM_LANDMARK = as.numeric(FTIME) - 7,
    
TIME_30 = pmin(TIME_FROM_LANDMARK, 30),

 EVENT_30 = case_when(
      TIME_FROM_LANDMARK <= 30 & EVENT_TYPE != 0 ~ EVENT_TYPE,
      TRUE ~ 0
    )
)
  
```

```{r}

# Create Labels for Status #

heart_clean <- heart2 %>%
  filter(!is.na(TIME_30), !is.na(EVENT_30), !is.na(ESCALATED)) %>%
  mutate(ESCALATED = factor(ESCALATED, levels = c(0, 1),
                            labels = c("Stable", "Escalated")))
```

```{r}

# Creating CIF Fit
cif_fit <- cuminc(
  ftime   = heart2$TIME_30,
  fstatus = heart2$EVENT_TYPE,
  group   = heart2$ESCALATED
)
```

```{r}

# P-Values for the Plot

pval_transplant    <- round(cif_fit$Tests["2", "pv"], 3)
pval_death         <- round(cif_fit$Tests["1", "pv"], 3)

pval_label <- paste0(
  "Transplanted: p-value = ", pval_transplant, "\n",
  "Death/Deterioration: p-value = ", pval_death
)

```

```{r}

# Plot the CIF

ggcompetingrisks(
  fit             = cif_fit,
  multiple_panels = FALSE,
  conf.int        = FALSE,
  palette         = c("#E69F00", "#56B4E9", "#E69F00", "#56B4E9"),
  linetype        = c("dotted", "solid", "dotted", "solid")
) +
  annotate(
    "text",
    x        = 1,
    y        = 0.95,
    label    = pval_label,
    hjust    = 0,
    vjust    = 1,
    fontface = "bold",
    size     = 4
  ) +
  scale_x_continuous(breaks = c(0, 7, 14, 21, 28), limits = c(0, 30)) +
  scale_y_continuous(breaks = seq(0, 0.9, by = 0.3), limits = c(0, 1)) +
  scale_color_manual(
    values = c("#E69F00", "#56B4E9"),
    labels = c("Death/Deterioration", "Transplanted")
  ) +
  scale_linetype_manual(
    values = c("dotted", "solid", "dotted", "solid"),
    labels = c("Stable", "Escalated", "Stable", "Escalated")
  ) +
  labs(
    x        = "Days since initial listing",
    y        = "Cumulative incidence function",
    color    = "Event",
    linetype = "Group"
  ) +
  theme_bw() +
  theme(
    legend.position  = "top",
    legend.text      = element_text(face = "bold", size = 11),
    axis.title       = element_text(face = "bold", size = 12),
    axis.text        = element_text(size = 11),
    panel.grid.minor = element_blank()
  )
```
=======
---
title: "CIF Plot"
author: "Sydney Shimizu"
format: html
editor: visual
embed-resources: true
---

## CIF Plotting

```{r}

heart2<-read.csv("C:/Users/12063/Downloads/heart2 (2).csv")

# install.packages("survminer") 

library(cmprsk)
library(survminer)
library(dplyr)
```

```{r}

# Compare Cumulative Incidence of Death/Deterioration and Cumulative Incidence of Heart Transplantation

heart2 <- heart2 %>%
  mutate(
    EVENT_TYPE = case_when(
      # Death / deterioration
      REM_CD_NEW %in% c(5,8,13) ~ 1,
      
      # Transplanted (deceased or living donor)
      REM_CD_NEW %in% c(2,3,4,14,21,22) ~ 2, 
      
      # Censor everyone else
      TRUE ~ 0
    ),
    
# Creating TIME_FROM_LANDMARK, TIME_30, and EVENT_30 Variables

TIME_FROM_LANDMARK = as.numeric(FTIME) - 7,
    
TIME_30 = pmin(TIME_FROM_LANDMARK, 30),

 EVENT_30 = case_when(
      TIME_FROM_LANDMARK <= 30 & EVENT_TYPE != 0 ~ EVENT_TYPE,
      TRUE ~ 0
    )
)

table(heart2$EVENT_30)
table(heart2$ESCALATED)
  
```

```{r}

# Create Labels for Status #

heart_clean <- heart2 %>%
  filter(!is.na(TIME_30), !is.na(EVENT_30), !is.na(ESCALATED)) %>%
  mutate(ESCALATED = factor(ESCALATED, levels = c(0, 1),
                            labels = c("Stable", "Escalated")))
```

```{r}

# Creating CIF Fit
cif_fit <- cuminc(
  ftime   = heart_clean$TIME_30,
  fstatus = heart_clean$EVENT_TYPE,
  group   = heart_clean$ESCALATED
)

cif_fit$Tests
table(heart_clean$ESCALATED, heart2$EVENT_30)
unique(heart_clean[,"ESCALATED"])


```

```{r}

# P-Values for the Plot

pval_transplant    <- round(cif_fit$Tests["2", "pv"], 3)
pval_death         <- round(cif_fit$Tests["1", "pv"], 3)

# Nicely formats small p-values as e.g. "< 0.001"
format_pval <- function(p) {
  ifelse(p < 0.001, "< 0.001", as.character(round(p, 3)))
}

pval_label <- paste0(
  "Transplanted: p-value = ", format_pval(pval_transplant), "\n",
  "Death/Deterioration: p-value = ", format_pval(pval_death)
)
```

```{r}

# Plot the CIF

ggcompetingrisks(
  fit             = cif_fit,
  multiple_panels = FALSE,
  conf.int        = FALSE,
  palette         = c("#E69F00", "#56B4E9", "#E69F00", "#56B4E9"),
  linetype        = c("dotted", "solid", "dotted", "solid"),
  size            = 1.5
) +
  annotate(
    "text",
    x        = 1,
    y        = 0.95,
    label    = pval_label,
    hjust    = 0,
    vjust    = 1,
    fontface = "bold",
    size     = 4
  ) +
  scale_x_continuous(breaks = c(0, 7, 14, 21, 28), limits = c(0, 30)) +
  scale_y_continuous(breaks = seq(0, 0.9, by = 0.3), limits = c(0, 1)) +
  scale_color_manual(
    values = c("#E69F00", "#56B4E9"),
    labels = c("Death/Deterioration", "Transplanted")
  ) +
  scale_linetype_manual(
    values = c("dotted", "solid", "dotted", "solid"),
    labels = c("Stable", "Escalated", "Stable", "Escalated")
  ) +
  labs(
    x        = "Days since landmark time",
    y        = "Cumulative incidence function",
    color    = "Event",
    linetype = "Group"
  ) +
  theme_bw() +
  theme(
    legend.position  = "top",
    legend.text      = element_text(face = "bold", size = 11),
    axis.title       = element_text(face = "bold", size = 12),
    axis.text        = element_text(size = 11),
    panel.grid.minor = element_blank()
  )

ggsave("C:/Users/12063/Documents/Programming in R/LAsBEST/cifplot.png", width = 8, height = 6, dpi = 300)
```

## Figures

```{r}
# BOXPLOTS FOR CONT VARS

library(ggplot2)
library(tidyr)
library(dplyr)

heart2 %>%
  select(INIT_AGE, INIT_BMI_NEW) %>%
  drop_na() %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
  mutate(variable = recode(variable,
                           "INIT_AGE"     = "Age",
                           "INIT_BMI_NEW" = "BMI")) %>%
  ggplot(aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(outlier.shape = 21, outlier.size = 1.5, alpha = 0.7) +
  facet_wrap(~variable, scales = "free") +
  scale_fill_manual(values = c("Age" = "#009E73", "BMI" = "#CC79A7")) +
  labs(
    x     = NULL,
    y     = "Value",
    title = "Distribution of Age and BMI"
  ) +
  theme_bw() +
  theme(
    legend.position  = "none",
    strip.text       = element_text(face = "bold", size = 11),
    axis.title       = element_text(face = "bold", size = 12),
    axis.text        = element_text(size = 11),
    panel.grid.minor = element_blank()
  )

ggsave("C:/Users/12063/Documents/Programming in R/LAsBEST/boxplot.png", width=8, height=6, dpi=300)
```

```{r}
mean(heart2$INIT_BMI_NEW,na.rm=TRUE)
mean(heart2$INIT_AGE,na.rm=TRUE)

```

```{r}
library(ggplot2)
library(patchwork)  # for combining plots side by side

p_age <- heart2 %>%
  select(INIT_AGE) %>%
  drop_na() %>%
  ggplot(aes(x = INIT_AGE)) +
  geom_histogram(fill = "#009E73", color = "white", bins = 30) +
  labs(
    x     = "Age (years)",
    y     = "Count",
    title = "Distribution of Age"
  ) +
  theme_bw() +
  theme(
    plot.title       = element_text(face = "bold", size = 12),
    axis.title       = element_text(face = "bold", size = 11),
    axis.text        = element_text(size = 10),
    panel.grid.minor = element_blank()
  )

p_bmi <- heart2 %>%
  select(INIT_BMI_NEW) %>%
  drop_na() %>%
  ggplot(aes(x = INIT_BMI_NEW)) +
  geom_histogram(fill = "#CC79A7", color = "white", bins = 30) +
  labs(
    x     = "BMI (kg/m²)",
    y     = "Count",
    title = "Distribution of BMI"
  ) +
  theme_bw() +
  theme(
    plot.title       = element_text(face = "bold", size = 12),
    axis.title       = element_text(face = "bold", size = 11),
    axis.text        = element_text(size = 10),
    panel.grid.minor = element_blank()
  )

p_age + p_bmi

ggsave("C:/Users/12063/Documents/Programming in R/LAsBEST/histograms.png", width=8, height=6, dpi=300)
```
>>>>>>> b4b68dd932f3b77685d79907564e2b4d870b7cbe
