---
title: "Side Analysis"
author: "Julia Aptekar"
format: html
embed-resources: true
---

## Goal
The goal of this side analysis is to study waitlist candidates from the time they reach status 2 until one of several possible outcomes occurs.

The landmark time, defined as time 0, is the first date on which a candidate is listed at status 2. For candidates initially listed at status 2, time 0 is their INIT_DATE. For candidates who reach status 2 later, time 0 is the first CHG_DATE on which their waitlist status becomes status 2.

After time 0, candidates are followed until one of the following outcomes occurs:

1. Death/deterioration
2. Heart transplantation
3. Recovery
4. Upgrade to Status 1
5. Downgrade to Status 3–6
6. Still on study or removed for another reason, treated as censoring
7. Delisting as Status 2999

## Idea
First, we identify the first date on which each candidate is listed at status 2. Candidates who are initially listed at status 2 have time 0 equal to their INIT_DATE. Candidates who are not initially listed at status 2 are included only if they later reach status 2; for these candidates, time 0 is the first CHG_DATE where their waitlist status becomes status 2.

Candidates who never reach status 2 are removed from the analysis. After time 0, we examine each candidate’s subsequent waitlist history and identify the first event that occurs among death/deterioration, heart transplantation, recovery, upgrade to status 1, downgrade to status 3–6, delisting as status 2999, or censoring.

The analysis time is calculated as the number of days from time 0 to the event or censoring date.

## Libraries/Loading
```{r, setup}
library(tidyverse)
library(lubridate)

heart <- read.csv("data/LAsBEST26_SUBSET.csv")
data <- readRDS("data/NEW_DATA.rds")
```


```{r setup,  output = F, message = F, warning= F}
library(tidyverse) # general tidying and manipulation
library(readxl) # reading xlsx files
library(kableExtra) # nice tables
library(lubridate) # easier working with dates
library(survival)
library(survminer)
library(table1)
library(cmprsk)
library(KMsurv)
# install.packages("fuzzyjoin")
library(fuzzyjoin)
```

# Loading in Data

```{r, output = F, message = F, warning= F}
heart_data <- read.csv("data/heart2.csv")
str(heart_data)

```

## Formatting columns

```{r}
heart_data <- heart_data %>% 
  mutate(
    DIAB = factor(
      if_else(
        DIAB == 1, 
        "No", 
        if_else(
          DIAB %in% c(2, 3, 4, 5), 
          "Yes", 
          NA_character_)
        ),
      levels = c("No", "Yes")
    ),
    ETHCAT = factor(
      case_when(
        ETHCAT == 1 ~ "White",
        ETHCAT == 2 ~ "Black",
        ETHCAT == 4 ~ "Hispanic",
        ETHCAT == 5 ~ "Asian",
        ETHCAT %in% c(6, 7, 9, 998) ~ "Other"
      ),
      levels = c("White", "Black", "Hispanic", "Asian", "Other")
    ),
    across(
      c(WL_ID_CODE, 
        INITIAL_STATUS, 
        STATUS_DAY7, 
        EXCEPTION, 
        REM_CD_NEW,
        FSTATUS, 
        GENDER, 
        BTYPE_REC, 
        ESCALATED), 
      as.factor
      ),
    across(
      c(FTIME, 
        INIT_BMI_NEW, 
        INIT_AGE), 
      as.numeric
      ),
    across(
      c(INIT_DATE, 
        DAY7_DATE, 
        LANDMARK_START, 
        HORIZON_DATE), 
      as_date)
  )
```

```{r}
data |> names()
names(heart_data)
```




```{r}

tibble(
  data_col = names(data),
  heart_data_col = names(heart_data)[best_match],
  edit_distance = apply(d, 1, min)
) |>
  arrange(edit_distance)

# columns that have stat in data

names(data)[grepl("stat", names(data), ignore.case = TRUE)]
```

Gender, Diab, initial age, initial date, ethcat, WL_ID_CODE, BMI, Blood Type, Rem_CD_NEW, FTIME, FSTATUS, EXCEPTION, 




```{r}
data$STATUS_DDR
```






## Cleaning
```{r}
data <- data %>% 
  select(WL_ID_CODE, INIT_STAT, END_STAT, INIT_DATE, END_DATE, CHG_DATE) %>%
  rename(INITIAL_STATUS = INIT_STAT,
    CHG_STATUS = END_STAT) %>%
  mutate(INIT_DATE = ymd(INIT_DATE),
    END_DATE = ymd(END_DATE),
    CHG_DATE = ymd(CHG_DATE),
    INITIAL_STATUS = case_when(
      INITIAL_STATUS == 2110 ~ 1,
      INITIAL_STATUS == 2120 ~ 2,
      INITIAL_STATUS == 2130 ~ 3,
      INITIAL_STATUS == 2140 ~ 4,
      INITIAL_STATUS == 2150 ~ 5,
      INITIAL_STATUS == 2160 ~ 6,
      TRUE ~ NA_real_),
    CHG_STATUS = case_when(
      CHG_STATUS == 2110 ~ 1,
      CHG_STATUS == 2120 ~ 2,
      CHG_STATUS == 2130 ~ 3,
      CHG_STATUS == 2140 ~ 4,
      CHG_STATUS == 2150 ~ 5,
      CHG_STATUS == 2160 ~ 6,
      CHG_STATUS == 2999 ~ 2999,
      TRUE ~ NA_real_))
```

## Code
```{r}
groups <- data %>%
  arrange(WL_ID_CODE, CHG_DATE) %>%
  group_by(WL_ID_CODE) %>%
  summarise(
    first_status = first(na.omit(INITIAL_STATUS)),
    init_date = first(na.omit(INIT_DATE)),
    time0 = {
      if (!is.na(first_status) && first_status == 2) {
        init_date
      } else {
        dates <- CHG_DATE[CHG_STATUS == 2 & !is.na(CHG_DATE)]
        if (length(dates) == 0) {
          as.Date(NA)
        } else {
          min(dates)
        }
      }
    },
    .groups = "drop") %>%
  filter(!is.na(time0)) %>%
  mutate(group = case_when(
    first_status == 2 ~ 0,
    TRUE ~ 1))
```

## Joins
```{r}
# This codes status-change events:
# 4 = upgrade to status 1
# 5 = downgrade to status 3-6
# 7 = delisted as status 2999

status_change_events <- data %>%
  left_join(groups %>% select(WL_ID_CODE, time0), by = "WL_ID_CODE") %>%
  filter(!is.na(time0)) %>%
  filter(CHG_DATE > time0) %>%
  mutate(event = case_when(
    INITIAL_STATUS == 2 & CHG_STATUS == 1 ~ 4, # upgrade from status 2 to status 1
    INITIAL_STATUS == 2 & CHG_STATUS %in% 3:6 ~ 5, # downgrade from status 2 to status 3-6
    INITIAL_STATUS == 2 & CHG_STATUS == 2999 ~ 7, # delisted from status 2 to status 2999
    TRUE ~ NA_real_)) %>%
  filter(!is.na(event)) %>%
  group_by(WL_ID_CODE) %>%
  slice_min(CHG_DATE, with_ties = FALSE) %>%
  ungroup() %>%
  transmute(WL_ID_CODE,
    event_date = CHG_DATE,
    event = event)

# joining some variables from original heart dataset
heart_sub <- heart %>%
  select(WL_ID_CODE, INIT_DATE, EXCEPTION, END_DATE_NEW, REM_CD_NEW, FSTATUS, GENDER, INIT_BMI_NEW, INIT_AGE, ETHCAT, DIAB, BTYPE_REC) %>%
  distinct(WL_ID_CODE, .keep_all = TRUE)

# This codes terminal/removal events:
# 1 = death/deterioration
# 2 = heart transplantation
# 6 = censoring/other reason
terminal_events <- heart_sub %>%
  left_join(groups %>% select(WL_ID_CODE, time0), by = "WL_ID_CODE") %>%
  mutate(END_DATE_NEW = ymd(END_DATE_NEW),
    event_date = END_DATE_NEW,
    event = case_when(REM_CD_NEW %in% c(5, 8, 13) ~ 1, # death/deterioration
      REM_CD_NEW %in% c(2, 3, 4, 14, 21, 22) ~ 2, # heart transplantation
      TRUE ~ 6)) %>% # censoring/other reason
  filter(!is.na(time0)) %>%
  filter(!is.na(event_date)) %>%
  filter(event_date > time0) %>%
  select(WL_ID_CODE, event_date, event)

# combine all event types and keep the first event after time0
all_events <- bind_rows(status_change_events, terminal_events) %>%
  group_by(WL_ID_CODE) %>%
  slice_min(event_date, with_ties = FALSE) %>%
  ungroup()

heart3 <- groups %>%
  left_join(all_events, by = "WL_ID_CODE") %>%
  left_join(heart_sub, by = "WL_ID_CODE") %>%
  mutate(time_to_event = as.numeric(event_date - time0))
```

## Writing CSV
```{r}
write.csv(heart3, "data/heart3.csv", row.names=FALSE)
```





















































## Cleaning
```{r}
data_og |> names()
```

```{r}
data <- 
  data %>% 
  select(WL_ID_CODE, INIT_STAT, END_STAT, INIT_DATE, END_DATE, CHG_DATE) %>%
  rename(
    INITIAL_STATUS = INIT_STAT,
    CHG_STATUS = END_STAT
  ) %>%
  mutate(
    INIT_DATE = ymd(INIT_DATE),
    END_DATE = ymd(END_DATE),
    CHG_DATE = ymd(CHG_DATE),
    INITIAL_STATUS = case_when(
      INITIAL_STATUS == 2110 ~ 1,
      INITIAL_STATUS == 2120 ~ 2,
      INITIAL_STATUS == 2130 ~ 3,
      INITIAL_STATUS == 2140 ~ 4,
      INITIAL_STATUS == 2150 ~ 5,
      INITIAL_STATUS == 2160 ~ 6,
      TRUE ~ NA_real_),
    CHG_STATUS = case_when(
      CHG_STATUS == 2110 ~ 1,
      CHG_STATUS == 2120 ~ 2,
      CHG_STATUS == 2130 ~ 3,
      CHG_STATUS == 2140 ~ 4,
      CHG_STATUS == 2150 ~ 5,
      CHG_STATUS == 2160 ~ 6,
      CHG_STATUS == 2999 ~ 2999,
      TRUE ~ NA_real_))
```

## Code
```{r}
groups <- data %>%
  # sorts all rows by candidate ID, then by change data 
  # This ensures chronological order within each candidate
  arrange(WL_ID_CODE, CHG_DATE) %>%
  
  # For each candidate...
  group_by(WL_ID_CODE) %>%
  summarise(
    # Grab the first nonNA value of INITIAL_STATUS AND INIT_DATE
    # these columns are constant per candidate (same listing info repeated across rows)
    # first() works fine here -- na.omit() is defensive
    first_status = first(na.omit(INITIAL_STATUS)),
    init_date = first(na.omit(INIT_DATE)),
    
    # if the candidate was initially listed at status 2, time0 = INIT_DATE
    
    time0 = {
      if (!is.na(first_status) && first_status == 2) {
        init_date
        
        # Otherwise, collects all CHG_DATES where CHG_STATUS became 2.
        # If none exist, time0 = NA (excluded later); if some exist, takes the earliest
      } else {
        dates <- CHG_DATE[CHG_STATUS == 2 & !is.na(CHG_DATE)]
        if (length(dates) == 0) {
          as.Date(NA)
        } else {
          min(dates)
        }
      }
    },
    .groups = "drop") %>%
  # drops candidates who never reached status 2
  filter(!is.na(time0)) %>%
  # Tags candidates: group 0 = initially listed at status 2
  # group 1 = reached it later
  mutate(group = case_when(
    first_status == 2 ~ 0,
    TRUE ~ 1))
```

## Joins
```{r}
# This codes status-change events:
## 3 = recovery 
# 4 = upgrade to status 1
# 5 = downgrade to status 3-6
# 7 = delisted as status 2999


# Brings time0 into the history table, then drops candidates not in the analysis
status_change_events <- data %>%
  left_join(groups %>% select(WL_ID_CODE, time0), by = "WL_ID_CODE") %>%
  filter(!is.na(time0)) %>%
  
  
  filter(CHG_DATE > time0) %>%
  mutate(event = case_when(
    INITIAL_STATUS == 2 & CHG_STATUS == 1 ~ 4, # upgrade from status 2 to status 1
    INITIAL_STATUS == 2 & CHG_STATUS %in% 3:6 ~ 5, # downgrade from status 2 to status 3-6
    INITIAL_STATUS == 2 & CHG_STATUS == 2999 ~ 7, # delisted from status 2 to status 2999
    TRUE ~ NA_real_)) %>%
  filter(!is.na(event)) %>%
  group_by(WL_ID_CODE) %>%
  slice_min(CHG_DATE, with_ties = FALSE) %>%
  ungroup() %>%
  transmute(WL_ID_CODE,
    event_date = CHG_DATE,
    event = event)

# joining some variables from original heart dataset
heart_sub <- heart %>%
  select(WL_ID_CODE, INIT_DATE, EXCEPTION, END_DATE_NEW, REM_CD_NEW, FSTATUS, GENDER, INIT_BMI_NEW, INIT_AGE, ETHCAT, DIAB, BTYPE_REC) %>%
  distinct(WL_ID_CODE, .keep_all = TRUE)

# This codes terminal/removal events:
# 1 = death/deterioration
# 2 = heart transplantation
# 6 = censoring/other reason
terminal_events <- heart_sub %>%
  left_join(groups %>% select(WL_ID_CODE, time0), by = "WL_ID_CODE") %>%
  mutate(END_DATE_NEW = ymd(END_DATE_NEW),
    event_date = END_DATE_NEW,
    event = case_when(REM_CD_NEW %in% c(5, 8, 13) ~ 1, # death/deterioration
      REM_CD_NEW %in% c(2, 3, 4, 14, 21, 22) ~ 2, # heart transplantation
      TRUE ~ 6)) %>% # censoring/other reason
  filter(!is.na(time0)) %>%
  filter(!is.na(event_date)) %>%
  filter(event_date > time0) %>%
  select(WL_ID_CODE, event_date, event)

# combine all event types and keep the first event after time0
all_events <- bind_rows(status_change_events, terminal_events) %>%
  group_by(WL_ID_CODE) %>%
  slice_min(event_date, with_ties = FALSE) %>%
  ungroup()

heart3 <- groups %>%
  left_join(all_events, by = "WL_ID_CODE") %>%
  left_join(heart_sub, by = "WL_ID_CODE") %>%
  mutate(time_to_event = as.numeric(event_date - time0))
```

## Writing CSV
```{r}
write.csv(heart3, "data/heart3.csv", row.names=FALSE)
```