---
title: "Heart Transplant Project - LA's BeST 2026"
author: "Jesus Rodriguez, Sydney Shimizu, Julia Aptekar"
format: html
---

```{r setup,  output = F, message = F, warning= F}
library(tidyverse) # general tidying and manipulation
library(readxl) # reading xlsx files
library(kableExtra) # nice tables
library(lubridate) # easier working with dates
```

# Loading in Data

```{r, output = F, message = F, warning= F}
heart_data <- read_csv("data/LAsBEST26_SUBSET.csv", col_select = -1)

# if needed here is the data dictionary
data_dictionary <- read_xlsx("data/STAR Files Data Dictionary.xlsx", skip = 1, sheet = 6)
data_dictionary <- read_xlsx("data/STAR Files Data Dictionary.xlsx", skip = 1, sheet = 1)

# note <- read_xlsx("data/STAR Files Data Dictionary.xlsx", n_max = 1, col_names = F)
```

# Data Cleaning

```{r}
# Checking structure
str(heart_data)

# Checking to see spread of certain variables to see if they are categorical or not
heart_data %>% 
  count(UNOS_CAND_STAT_CD)

heart_data %>% 
  count(REM_CD_NEW)

heart_data %>% 
  count(FTIME)

names(heart_data)
```

## Formatting columns

```{r}
heart_data <- 
  heart_data %>%
  mutate(
    across(
      c(WL_ID_CODE, #ID
        WLREG_AUDIT_ID_CODE, #ID 
        UNOS_CAND_STAT_CD_WAITLIST, 
        UNOS_CAND_STAT_CD, EXCEPTION, 
        REM_CD_NEW, 
        FSTATUS, 
        GENDER, 
        ETHCAT, 
        BTYPE_REC, 
        DIAB),
      as.factor
    ),
    across(
      c(FTIME, 
        INIT_BMI_NEW, 
        INIT_AGE),
      as.numeric
    ),
    across(
      c(INIT_DATE,
        CHG_DATE,
        END_DATE_NEW
      ),
      as_date
    )
  )
  
```

## Cleaning Priority Status

This just makes the data easier to work with.

Planning to clean `FINAL_STATUS` later, as at times in the data, it doesn't really reflect the actual "final" status a patient had when they were last seen

```{r}
#Recode UNOS_WAIT code to match up with STATUS#
heart_data <- heart_data %>%
  mutate(
    INITIAL_STATUS = dplyr::case_when(
      UNOS_CAND_STAT_CD_WAITLIST == 2110 ~ 1,
      UNOS_CAND_STAT_CD_WAITLIST == 2120 ~ 2,
      UNOS_CAND_STAT_CD_WAITLIST == 2130 ~ 3,
      UNOS_CAND_STAT_CD_WAITLIST == 2140 ~ 4,
      UNOS_CAND_STAT_CD_WAITLIST == 2150 ~ 5,
      UNOS_CAND_STAT_CD_WAITLIST == 2160 ~ 6,
      TRUE ~ NA_real_
    ),
     FINAL_STATUS = dplyr::case_when(
      UNOS_CAND_STAT_CD == 2110 ~ 1,
      UNOS_CAND_STAT_CD == 2120 ~ 2,
      UNOS_CAND_STAT_CD == 2130 ~ 3,
      UNOS_CAND_STAT_CD == 2140 ~ 4,
      UNOS_CAND_STAT_CD == 2150 ~ 5,
      UNOS_CAND_STAT_CD == 2160 ~ 6,
      TRUE ~ NA_real_
  ),
  across(
    c(INITIAL_STATUS, FINAL_STATUS),
    as.factor
  )
)
```

# Landmark and Horizon Dates

Making two variables: `landmark_date` and `horizon_date`

`landmark_date`: Adds 7 days (1 week) to the initial date an observation was seen `horizon_date`: Adds 37 days (1 week + 30 days) to the initial date an observation was seen

```{r}
heart_data <- 
  heart_data %>% 
  mutate(
    landmark_date = 
      INIT_DATE + weeks(1),
    horizon_date = 
      INIT_DATE + days(37)
  )
```

## Subsetting our cohort to not include anything past the horizon date (37 days)

If a date is greater than 37 days past the initial, remove it from the data

**We will still need to subset observations before the landmark date later.**

```{r}
heart_data <- 
  heart_data %>% 
  filter(CHG_DATE <= horizon_date) %>% 
  arrange(WL_ID_CODE, CHG_DATE)
```

# Refining our cohort

*BEWARE of people who 2->3->2 before landmark time, they do NOT count as stable* 

Individuals who were listed were status 2 at day 7

```{r}
heart_data2 <- 
  heart_data %>%
  group_by(WL_ID_CODE) %>%
  mutate(
    keep_in_study = 
      if_else(
        (FINAL_STATUS == 2 & CHG_DATE <= landmark_date),
        "YES",
        "NO"
      )
  ) %>% 
  filter(
    any(keep_in_study == "YES")
    ) %>% 
  arrange(WL_ID_CODE)
```

Making sure we are not losing any observations from IDs that have a final status of 2 before the landmark date

```{r}
counts2 <- heart_data2 %>% 
  count(WL_ID_CODE) %>% 
  arrange(desc(n))

counts <- heart_data %>% 
  count(WL_ID_CODE) %>% 
  arrange(desc(n))

counts_joined <- full_join(
  counts,
  counts2,
  by = "WL_ID_CODE",
  suffix = c("_heart_data", "_heart_data2")
) %>% 
  filter(!is.na(n_heart_data2)) %>% 
  mutate(equal = n_heart_data == n_heart_data2)
```

Checking whether we are missing data with counts_joined\$equal == FALSE, we can see that this argument is FALSE for all observations, meaning that we are not missing any data with this filter.

```{r}
heart_data <- heart_data2
```

## Removing individuals who were status 1

From my previous manipulation, I already made sure that everyone included so far has a priority status of 2 before the landmark date. So if anyone has an initial status of 1, then they should be removed from the analysis.

```{r}
heart_data <- 
heart_data %>%
  group_by(WL_ID_CODE) %>%
  mutate(
    keep_in_study = 
      if_else(
        (INITIAL_STATUS == 1),
        "NO",
        "YES"
      )
  ) %>% 
  filter(
    any(keep_in_study == "YES")
  ) %>% 
  arrange(WL_ID_CODE, CHG_DATE)
```

This is an ID of someone who was initally at status 1, but dropped to status 2 before the landmark date. We don't want them in our cohort, so they should be removed: 1430077


## Excluding observations before the landmark time

```{r}
heart_data_test_landmark <- 
  heart_data %>% 
  filter(CHG_DATE >= landmark_date) %>% 
  arrange(WL_ID_CODE, CHG_DATE)
```

# Double checking to see if we have the correct amount of unique IDs ~4347

Unique ID Codes

```{r}
heart_ids <- 
heart_data %>% 
  distinct(WL_ID_CODE)

heart_ids_after_landmark <- 
heart_data_test_landmark %>% 
  distinct(WL_ID_CODE)
```

FTIME AND FSTATUS

```{r}
dat <- dat %>%
  mutate(
    REM_CD_NEW = case_when(
      END_DATE > "2024-09-30" ~ 9999,
      .default = REM_CD
    ),
    END_DATE_NEW = case_when(
      END_DATE > "2024-09-30" ~ as.Date("2024-09-30"),
      .default = END_DATE
    ),
    # Create competing risks outcomes
    FTIME = END_DATE_NEW - INIT_DATE,
    FTIME = ifelse(FTIME == 0, 0.9, FTIME),
    FSTATUS = case_when(
      REM_CD_NEW %in% c(5, 8, 13) ~ 1, # Death/Deterioration
      REM_CD_NEW %in% c(2, 3, 4, 14, 21, 22) ~ 2, # Transplant 
      REM_CD_NEW == 12 ~ 3, # Recovered
      .default = 0
    )
  )
```

