---
title: "Heart Transplant Project - LA's BeST 2026"
author: "Jesus Rodriguez, Sydney Shimizu, Julia Aptekar"
format: html
---

```{r setup,  output = F, message = F, warning= F}
library(tidyverse) # general tidying and manipulation
library(readxl) # reading xlsx files
library(kableExtra) # nice tables
library(lubridate) # easier working with dates
```

# Loading in Data

```{r, output = F, message = F, warning= F}
heart_data <- read_csv("data/LAsBEST26_SUBSET.csv", col_select = -1)

# if needed here is the data dictionary
data_dictionary <- read_xlsx("data/STAR Files Data Dictionary.xlsx", skip = 1, sheet = 6)
data_dictionary <- read_xlsx("data/STAR Files Data Dictionary.xlsx", skip = 1, sheet = 1)

# note <- read_xlsx("data/STAR Files Data Dictionary.xlsx", n_max = 1, col_names = F)
```

# Data Cleaning

```{r}
# Checking structure
str(heart_data)

# Checking to see spread of certain variables to see if they are categorical or not
heart_data %>% 
  count(UNOS_CAND_STAT_CD)

heart_data %>% 
  count(REM_CD_NEW)

heart_data %>% 
  count(FTIME)

names(heart_data)
```

## Formatting columns

```{r}
heart_data <- 
  heart_data %>%
  mutate(
    across(
      c(WL_ID_CODE, #ID
        WLREG_AUDIT_ID_CODE, #ID 
        UNOS_CAND_STAT_CD_WAITLIST, 
        UNOS_CAND_STAT_CD, EXCEPTION, 
        REM_CD_NEW, 
        FSTATUS, 
        GENDER, 
        ETHCAT, 
        BTYPE_REC, 
        DIAB),
      as.factor
    ),
    across(
      c(FTIME, 
        INIT_BMI_NEW, 
        INIT_AGE),
      as.numeric
    ),
    across(
      c(INIT_DATE,
        CHG_DATE,
        END_DATE_NEW
      ),
      as_date
    )
  )
  
```

## Cleaning Priority Status

This just makes the data easier to work with.

Planning to clean `FINAL_STATUS` later, as at times in the data, it doesn't really reflect the actual "final" status a patient had when they were last seen

```{r}
#Recode UNOS_WAIT code to match up with STATUS#
heart_data <- heart_data %>%
  mutate(
    INITIAL_STATUS = dplyr::case_when(
      UNOS_CAND_STAT_CD_WAITLIST == 2110 ~ 1,
      UNOS_CAND_STAT_CD_WAITLIST == 2120 ~ 2,
      UNOS_CAND_STAT_CD_WAITLIST == 2130 ~ 3,
      UNOS_CAND_STAT_CD_WAITLIST == 2140 ~ 4,
      UNOS_CAND_STAT_CD_WAITLIST == 2150 ~ 5,
      UNOS_CAND_STAT_CD_WAITLIST == 2160 ~ 6,
      TRUE ~ NA_real_
    ),
     CHANGE_STATUS = dplyr::case_when(
      UNOS_CAND_STAT_CD == 2110 ~ 1,
      UNOS_CAND_STAT_CD == 2120 ~ 2,
      UNOS_CAND_STAT_CD == 2130 ~ 3,
      UNOS_CAND_STAT_CD == 2140 ~ 4,
      UNOS_CAND_STAT_CD == 2150 ~ 5,
      UNOS_CAND_STAT_CD == 2160 ~ 6,
      TRUE ~ NA_real_
  ),
  across(
    c(INITIAL_STATUS, CHANGE_STATUS),
    as.factor
  )
)
```

The true FINAL status

```{r}
heart_data <- 
  heart_data %>% 
  group_by(WL_ID_CODE) %>% 
  mutate(
    FINAL_STATUS = 
      last(CHANGE_STATUS)
  )
```


# Landmark and Horizon Dates

Making two variables: `landmark_date` and `horizon_date`

`landmark_date`: Adds 7 days (1 week) to the initial date an observation was seen `horizon_date`: Adds 37 days (1 week + 30 days) to the initial date an observation was seen

```{r}
heart_data <- 
  heart_data %>% 
  mutate(
    landmark_date = 
      INIT_DATE + weeks(1),
    horizon_date = 
      INIT_DATE + days(37)
  )
```

## Subsetting our cohort to not include anything past the horizon date (37 days)

If a date is greater than 37 days past the initial, remove it from the data

**We will still need to subset observations before the landmark date later.**

```{r}
heart_data <- 
  heart_data %>% 
  filter(CHG_DATE <= horizon_date) %>% 
  arrange(WL_ID_CODE, CHG_DATE)
```

# Refining our cohort

## A new approach

```{r}
heart_data_before_landmark <- 
  heart_data %>% 
  filter(CHG_DATE <= landmark_date)
```

## Remove people at initial status 1 

```{r}
heart_data2 <- 
  heart_data_before_landmark %>%
  group_by(WL_ID_CODE) %>%
  filter(!INITIAL_STATUS == 1) %>% 
  arrange(WL_ID_CODE, CHG_DATE)
```

## Classify Stable individuals, those whose CHANGE_STATUS was 2 for ALL before landmark date == YES rows

```{r}
heart_data2 <- 
  heart_data2 %>% 
  group_by(WL_ID_CODE) %>% 
  mutate(
    stable = 
      if_else(
        all(CHANGE_STATUS == 2),
        "YES",
        "NO"
      )
  ) %>% 
  filter(FTIME > 7)
```

```{r}
stable_ids <- 
  heart_data2 %>% 
  filter(stable == "YES")
```

## Classify Escalating individuals

```{r}
escalating_ids <- 
heart_data2 %>% 
  group_by(WL_ID_CODE) %>% 
  filter(stable == "NO") %>% 
  filter(last(CHANGE_STATUS == 2))
```


## Joining by IDs

```{r}
stable_ids <- 
  stable_ids %>% 
  select(WL_ID_CODE)
```

```{r}
escalating_ids <- 
  escalating_ids %>% 
  select(WL_ID_CODE)
```


```{r}
stable_ids %>% 
  distinct(WL_ID_CODE)

escalating_ids %>% 
  distinct(WL_ID_CODE)
```

## Joining

```{r}
left_join(
  
)
```


















































*BEWARE of people who 2->3->2 before landmark time, they do NOT count as stable*

## Removing individuals who were status 1

From my previous manipulation, I already made sure that everyone included so far has a priority status of 2 before the landmark date. So if anyone has an initial status of 1, then they should be removed from the analysis.

```{r}
heart_data <- 
  heart_data %>%
  group_by(WL_ID_CODE) %>%
  filter(!INITIAL_STATUS == 1) %>% 
  arrange(WL_ID_CODE, CHG_DATE)
```

This is an ID of someone who was initally at status 1, but dropped to status 2 before the landmark date. We don't want them in our cohort, so they should be removed: 1430077

## Classify Stable individuals, those whose CHANGE_STATUS was 2 for ALL before landmark date == YES rows

```{r}
heart_data <- 
  heart_data %>% 
  group_by(WL_ID_CODE) %>% 
  mutate(
    before_landmark_date = 
      if_else(
        (CHG_DATE <= landmark_date),
        "YES",
        "NO"
      )
  )
```


```{r}
heart_data <-
  heart_data %>%
  group_by(WL_ID_CODE) %>%
  mutate(
    stable =
      all(
          CHANGE_STATUS[before_landmark_date == "YES"] == 2
      )
  )

```

## Classify Escalated Patients

Those who were intially status 3-6, but changed to status 2 right before the landmark date

```{r}
heart_data <- 
  heart_data %>%
  group_by(WL_ID_CODE) %>%
  mutate(
    keep_in_study = 
      if_else(
        (CHANGE_STATUS == 2 & CHG_DATE <= landmark_date),
        "YES",
        "NO"
      )
  ) %>% 
  filter(
    any(keep_in_study == "YES")
    ) %>% 
  arrange(WL_ID_CODE)
```

```{r}
heart_data_escalated <- heart_data %>%
  group_by(WL_ID_CODE) %>%
  arrange(CHG_DATE) %>%
  filter(CHG_DATE <= landmark_date) %>%
  slice_tail(n = 1) %>%
  mutate(
    ESCALATED = if_else(
      as.numeric(as.character(INITIAL_STATUS)) %in% c(3,4,5,6) &
      as.numeric(as.character(CHANGE_STATUS)) == 2,
      1L,
      0L
    )
  ) %>%
  ungroup()
```

```{r}

```

```{r}

```



## Excluding observations before the landmark time

```{r}
heart_data_test_landmark <- 
  heart_data3 %>% 
  filter(CHG_DATE >= landmark_date) %>% 
  arrange(WL_ID_CODE, CHG_DATE)

heart_data_test_landmark <- 
  heart_data %>% 
  filter(CHG_DATE >= landmark_date) %>% 
  arrange(WL_ID_CODE, CHG_DATE)
```

# Double checking to see if we have the correct amount of unique IDs ~4347

Unique ID Codes

```{r}
heart_ids <- 
heart_data_escalated %>% 
  distinct(WL_ID_CODE)

heart_ids_after_landmark <- 
heart_data_test_landmark %>% 
  distinct(WL_ID_CODE)
```

FTIME AND FSTATUS

```{r}
dat <- dat %>%
  mutate(
    REM_CD_NEW = case_when(
      END_DATE > "2024-09-30" ~ 9999,
      .default = REM_CD
    ),
    END_DATE_NEW = case_when(
      END_DATE > "2024-09-30" ~ as.Date("2024-09-30"),
      .default = END_DATE
    ),
    # Create competing risks outcomes
    FTIME = END_DATE_NEW - INIT_DATE,
    FTIME = ifelse(FTIME == 0, 0.9, FTIME),
    FSTATUS = case_when(
      REM_CD_NEW %in% c(5, 8, 13) ~ 1, # Death/Deterioration
      REM_CD_NEW %in% c(2, 3, 4, 14, 21, 22) ~ 2, # Transplant 
      REM_CD_NEW == 12 ~ 3, # Recovered
      .default = 0
    )
  )
```

<<<<<<< HEAD

```{r}
# write.csv(heart2, file = "data/clean_heart_data.csv")
# write.csv(heart2_long, file = "data/clean_heart_data_long.csv")
```

# Descriptive Statistics

```{r}
library(tableone)
install.packages("table1")
library(table1)
```

```{r}
# Install and load package
install.packages("gtsummary")
library(gtsummary)

clean_heart_data <- 
  read_csv("data/clean_heart_data.csv")
clean_heart_data_long <- 
  read_csv("data/clean_heart_data_long.csv")
```







```{r}
clean_heart_data_long <- 
  clean_heart_data_long %>%
  mutate(
    across(
      c(WL_ID_CODE, #ID
        WLREG_AUDIT_ID_CODE, #ID 
        UNOS_CAND_STAT_CD_WAITLIST, 
        UNOS_CAND_STAT_CD, 
        EXCEPTION, 
        REM_CD_NEW, 
        FSTATUS, 
        GENDER, 
        ETHCAT, 
        BTYPE_REC, 
        DIAB,
        ESCALATED),
      as.factor
    ),
    across(
      c(FTIME, 
        INIT_BMI_NEW, 
        INIT_AGE),
      as.numeric
    ),
    across(
      c(INIT_DATE,
        CHG_DATE,
        END_DATE_NEW
      ),
      as_date
    )
  )
  
```

```{r}
clean_heart_data_long %>% 
  select(
    INIT_AGE, INIT_BMI_NEW, GENDER, ETHCAT, BTYPE_REC, EXCEPTION, DIAB, ESCALATED
  ) %>% 
tbl_summary(by = ESCALATED) 
```

```{r}
library(table1)

table1(
  ~ INIT_AGE +
    INIT_BMI_NEW +
    GENDER +
    ETHCAT +
    BTYPE_REC +
    EXCEPTION +
    DIAB
  | ESCALATED,
  data = clean_heart_data_long
)
```

# Making a table

```{r}
heart <- read_csv("data/clean_heart_data_long.csv")
```


```{r}
# DIAB statuses/types
heart$DIAB <- ifelse(is.na(heart$DIAB), 0, heart$DIAB) # making NA become 0s
heart$DIAB2 <- factor(heart$DIAB, levels = c(0, 1, 2, 3, 4, 5, 998),
                      labels = c("Not Reported", "No", "Type I", "Type II", "Type Other", "Type Unknown", "Status Unknown"), 
                      ordered=TRUE)

# Gender
heart$GENDER2 <- factor(heart$GENDER, levels = c("F", "M"),
                        labels = c("Female", "Male"),
                        ordered = TRUE)

# ETHCAT
heart$ETHCAT2 <- factor(heart$ETHCAT, levels = c(1, 2, 4, 5, 6, 7, 9, 998),
                        labels = c("White", "Black", "Hispanic", "Asian", "AI/AN", "Hawaiian/Pacific-Islander", "Multiracial", "Unknown"))

# EXCEPTION
heart$EXCEPTION2 <- factor(heart$EXCEPTION, levels = c(0, 1),
                           labels = c("No Exception", "Exception"))
```

```{r}
library(table1)

table1(
  ~ INIT_AGE +
    INIT_BMI_NEW +
    GENDER2 +
    ETHCAT2 +
    BTYPE_REC +
    EXCEPTION2 +
    DIAB2
  | ESCALATED,
  data = heart
)
```

=======
>>>>>>> 8da89ee13b7a3473da747eb282cef804a2fb0fba
