---
title: "Heart Transplant Project - LA's BeST 2026"
author: "Jesus Rodriguez, Sydney Shimizu, Julia Aptekar"
format: html
---

```{r setup,  output = F, message = F, warning= F}
library(tidyverse) # general tidying and manipulation
library(readxl) # reading xlsx files
library(kableExtra) # nice tables
library(lubridate) # easier working with dates
```

# Loading in Data

```{r, output = F, message = F, warning= F}
heart_data <- read_csv("data/LAsBEST26_SUBSET.csv", col_select = -1)

# if needed here is the data dictionary
data_dictionary <- read_xlsx("data/STAR Files Data Dictionary.xlsx", skip = 1, sheet = 6)
data_dictionary <- read_xlsx("data/STAR Files Data Dictionary.xlsx", skip = 1, sheet = 1)

# note <- read_xlsx("data/STAR Files Data Dictionary.xlsx", n_max = 1, col_names = F)
```

# Data Cleaning

```{r}
# Checking structure
str(heart_data)

# Checking to see spread of certain variables to see if they are categorical or not
heart_data %>% 
  count(UNOS_CAND_STAT_CD)

heart_data %>% 
  count(REM_CD_NEW)

heart_data %>% 
  count(FTIME)

names(heart_data)
```
## Formatting columns

```{r}
heart_data <- 
  heart_data %>%
  mutate(
    across(
      c(WL_ID_CODE, #ID
        WLREG_AUDIT_ID_CODE, #ID 
        UNOS_CAND_STAT_CD_WAITLIST, 
        UNOS_CAND_STAT_CD, EXCEPTION, 
        REM_CD_NEW, 
        FSTATUS, 
        GENDER, 
        ETHCAT, 
        BTYPE_REC, 
        DIAB),
      as.factor
    ),
    across(
      c(FTIME, 
        INIT_BMI_NEW, 
        INIT_AGE),
      as.numeric
    ),
    across(
      c(INIT_DATE,
        CHG_DATE,
        END_DATE_NEW
      ),
      as_date
    )
  )
  
```

## Cleaning Priority Status

```{r}
#Recode UNOS_WAIT code to match up with STATUS#
heart_data <- heart_data %>%
  mutate(
    INITIAL_STATUS = dplyr::case_when(
      UNOS_CAND_STAT_CD_WAITLIST == 2110 ~ 1,
      UNOS_CAND_STAT_CD_WAITLIST == 2120 ~ 2,
      UNOS_CAND_STAT_CD_WAITLIST == 2130 ~ 3,
      UNOS_CAND_STAT_CD_WAITLIST == 2140 ~ 4,
      UNOS_CAND_STAT_CD_WAITLIST == 2150 ~ 5,
      UNOS_CAND_STAT_CD_WAITLIST == 2160 ~ 6,
      TRUE ~ NA_real_
    ),
     FINAL_STATUS = dplyr::case_when(
      UNOS_CAND_STAT_CD == 2110 ~ 1,
      UNOS_CAND_STAT_CD == 2120 ~ 2,
      UNOS_CAND_STAT_CD == 2130 ~ 3,
      UNOS_CAND_STAT_CD == 2140 ~ 4,
      UNOS_CAND_STAT_CD == 2150 ~ 5,
      UNOS_CAND_STAT_CD == 2160 ~ 6,
      TRUE ~ NA_real_
  ),
  across(
    c(INITIAL_STATUS, FINAL_STATUS),
    as.factor
  )
)
```

# Using a landmark date of 7 days, define two groups

```{r}
heart_data <- 
  heart_data %>% 
  mutate(
    landmark_date = 
      INIT_DATE + weeks(1),
    horizon_date = 
      INIT_DATE + days(37)
  )
```

If a date is greater than 37 days past the initial, remove it from the data

```{r}
heart_data <- 
  heart_data %>% 
  filter(CHG_DATE <= horizon_date) %>% 
  arrange(WL_ID_CODE, CHG_DATE)
```

## Defining "Stable"


Individuals who were listed at status 2 and were status 2 at day 7

```{r}
heart_data2 <- 
heart_data %>%
  group_by(WL_ID_CODE) %>%
  mutate(
    keep_in_study = 
      if_else(
        (FINAL_STATUS == 2 & CHG_DATE < landmark_date),
        "YES",
        "NO"
      )
  ) %>% 
  filter(
    any(keep_in_study == "YES")
    ) %>% 
  arrange(WL_ID_CODE)
```

Making sure we are not losing any observations from IDs that have a final status of 2 before the landmark date

```{r}
counts2 <- heart_data2 %>% 
  count(WL_ID_CODE) %>% 
  arrange(desc(n))

counts <- heart_data %>% 
  count(WL_ID_CODE) %>% 
  arrange(desc(n))

counts_joined <- full_join(
  counts,
  counts2,
  by = "WL_ID_CODE",
  suffix = c("_heart_data", "_heart_data2")
) %>% 
  filter(!is.na(n_heart_data2)) %>% 
  mutate(equal = n_heart_data == n_heart_data2)
  
```

Unique ID codes

```{r}
heart_data %>% 
  distinct(WL_ID_CODE)

heart_data2 %>% 
  distinct(WL_ID_CODE)
```

```{r}
heart_data2 %>% 
  filter(WL_ID)
```

