---
title: "Heart Transplant Project - LA's BeST 2026"
author: "Jesus Rodriguez, Sydney Shimizu, Julia Aptekar"
format: html
embed-resources: true
---


```{r setup,  output = F, message = F, warning= F}
library(tidyverse) # general tidying and manipulation
library(readxl) # reading xlsx files
library(kableExtra) # nice tables
library(lubridate) # easier working with dates
library(survival)
library(survminer)
library(table1)
library(cmprsk)
library(KMsurv)
```

# Loading in Data

```{r, output = F, message = F, warning= F}
heart_data <- read.csv("data/heart2.csv")
str(heart_data)

```

## Formatting columns

```{r}
heart_data <- heart_data %>% 
  mutate(
    DIAB = factor(
      if_else(
        DIAB == 1, 
        "No", 
        if_else(
          DIAB %in% c(2, 3, 4, 5), 
          "Yes", 
          NA_character_)
        ),
      levels = c("No", "Yes")
    ),
    ETHCAT = factor(
      case_when(
        ETHCAT == 1 ~ "White",
        ETHCAT == 2 ~ "Black",
        ETHCAT == 4 ~ "Hispanic",
        ETHCAT == 5 ~ "Asian",
        ETHCAT %in% c(6, 7, 9, 998) ~ "Other"
      ),
      levels = c("White", "Black", "Hispanic", "Asian", "Other")
    ),
    across(
      c(WL_ID_CODE, 
        INITIAL_STATUS, 
        STATUS_DAY7, 
        EXCEPTION, 
        REM_CD_NEW,
        FSTATUS, 
        GENDER, 
        BTYPE_REC, 
        ESCALATED), 
      as.factor
      ),
    across(
      c(FTIME, 
        INIT_BMI_NEW, 
        INIT_AGE), 
      as.numeric
      ),
    across(
      c(INIT_DATE, 
        DAY7_DATE, 
        LANDMARK_START, 
        HORIZON_DATE), 
      as_date)
  )
```

## Cleaning

```{r}

# Compare Cumulative Incidence of Death/Deterioration and Cumulative Incidence of Heart Transplantation

heart_data <- heart_data %>%
  mutate(
    EVENT_TYPE = case_when(
      # Death / deterioration
      REM_CD_NEW %in% c(5,8,13) ~ 1,
      
      # Transplanted (deceased or living donor)
      REM_CD_NEW %in% c(2,3,4,14,21,22) ~ 2, 
      
      # Censor everyone else
      TRUE ~ 0
    ),
    
    # Creating TIME_FROM_LANDMARK, TIME_30, and EVENT_30 Variables
    
    TIME_FROM_LANDMARK = as.numeric(FTIME) - 7,
    
    TIME_30 = pmin(TIME_FROM_LANDMARK, 30),
    
    EVENT_30 = factor(case_when(
      TIME_FROM_LANDMARK <= 30 & EVENT_TYPE != 0 ~ EVENT_TYPE,
      TRUE ~ 0
    )
    )
  )

```

Create Labels for Status

```{r}
# Create Labels for Status #

heart_data <- heart_data %>%
  filter(!is.na(TIME_30), !is.na(EVENT_30), !is.na(ESCALATED)) %>%
  mutate(ESCALATED = factor(ESCALATED, levels = c(0, 1),
                            labels = c("Stable", "Escalated")))
```

# Survival Analysis with Competing Risks

Models

Subdistribution Hazard Models and Cause-Specific Hazard Models

# Covariance Matrix Setup

```{r}
vars <- c("TIME_30", "EVENT_30", "ESCALATED", "GENDER", "INIT_AGE",
          "INIT_BMI_NEW", "ETHCAT", "DIAB", "BTYPE_REC", "EXCEPTION")

heart_matrix <- 
  heart_data[complete.cases(heart_data[, vars]), vars]

covariance_matrix <- 
  model.matrix(
    ~ ESCALATED + GENDER + INIT_AGE + INIT_BMI_NEW + ETHCAT + DIAB + BTYPE_REC + EXCEPTION,
    data = heart_matrix
  )[, -1] 

# str(heart_matrix)
```

# Solo Variable Models (Unadjusted Models)

## Cause-Specific (Unadjusted)

```{r}
# Cox Model 1 : Death / Deterioration
cs_model_death <- 
  heart_data %>% 
  coxph(
  Surv(TIME_30, EVENT_30 == 1) ~ 
    ESCALATED,
  data = .
)
summary(cs_model_death)
```

```{r}

# Cox Model 2 : Transplantation
cs_model_transplant <- coxph(
  Surv(TIME_30, EVENT_30 == 2) ~ 
    ESCALATED,
  data = heart_data
)
summary(cs_model_transplant)
```

## Sub-Distribution Models (Unadjusted)

```{r}

# Fine-Gray subdistribution Model 1: Death / Deterioration
crr_death <- crr(
  ftime   = heart_matrix$TIME_30,
  fstatus = heart_matrix$EVENT_30,
  cov1    = covariance_matrix[,1],
  failcode = 1
  ) ; summary(crr_death)

# summary(crr_death)$coef[2]
# round(1-summary(crr_death)$coef[2], 2)

```

**Barebones Interpretation, needs work**

With a hazard ratio of `r summary(crr_death)$coef[2]` and a p-value of `r summary(crr_death)$coef[5]`. Escalated patients have a `r round(1-summary(crr_death)$coef[2], 2)*100`% lower subdistribution hazard of death/deterioration compared to stable patients

Which is good, if you escalated to status 2 for a heart transplant, it would be nice to hear that death is less likely than a transplant.




```{r}
# Fine-Gray subdistribution Model 2: Transplant
crr_transplant <- crr(
  ftime   = heart_matrix$TIME_30,
  fstatus = heart_matrix$EVENT_30,
  cov1    = covariance_matrix[,1],
  failcode = 2
  ) ; summary(crr_transplant)

```

**Barebones Interpretation, needs work**

Escalated patients have a 13% higher subdistribution hazard of transplant.

Patients who escalated to status 2, are a bit more likely to have transplantation as an outcome.

#  Adjusted Models

## Cause-Specific (Adjusted)

```{r}
# Cox Model 1 : Death / Deterioration
cs_model_death_adjusted <- 
  heart_data %>% 
  coxph(
  Surv(TIME_30, EVENT_30 == 1) ~ 
    ESCALATED + GENDER + ETHCAT + EXCEPTION + INIT_AGE + DIAB + INIT_BMI_NEW + BTYPE_REC,
  data = .
)
summary(cs_model_death_adjusted)
```

```{r}

# Cox Model 2 : Transplantation
cs_model_transplant_adjusted <- coxph(
  Surv(TIME_30, EVENT_30 == 2) ~ 
    ESCALATED + GENDER + ETHCAT + EXCEPTION + INIT_AGE + DIAB + INIT_BMI_NEW + BTYPE_REC,
  data = heart_data
)
summary(cs_model_transplant_adjusted)
```

## Sub-Distribution Models (Adjusted)

```{r}
# Fine-Gray subdistribution Model 1: Death / Deterioration
crr_death_adjusted <- crr(
  ftime   = heart_matrix$TIME_30,
  fstatus = heart_matrix$EVENT_30,
  cov1    = covariance_matrix,
  failcode = 1
  ) ; summary(crr_death_adjusted)
```

**Barebones Interpretation, needs work**





```{r}
# Fine-Gray subdistribution Model 2: Transplant
crr_transplant_adjusted <- crr(
  ftime   = heart_matrix$TIME_30,
  fstatus = heart_matrix$EVENT_30,
  cov1    = covariance_matrix,
  failcode = 2
  ) ; summary(crr_transplant_adjusted) 

summary(crr_transplant_adjusted)$coef
```

**Barebones Interpretation, needs work**

# Presentable Models

## Cox-Models

```{r}
summary(cs_model_transplant_adjusted)$coefficients[1, "Pr(>|z|)"]

summary(crr_death_adjusted)
```

## Results Table

Function to extrac the info we need from the cox models

```{r}
# Extract from coxph model
extract_cox <- function(model, outcome) {
  s <- summary(model)
  data.frame(
    Outcome  = outcome,
    HR       = s$conf.int[1, "exp(coef)"],
    Lower    = s$conf.int[1, "lower .95"],
    Upper    = s$conf.int[1, "upper .95"],
    P        = s$coefficients[1, "Pr(>|z|)"]
  )
}
```

Function to extract the info we need from the Fine-Gray models


```{r}
# Extract from crr model
extract_crr <- function(model, outcome) {
  s <- summary(model)
  data.frame(
    Outcome  = outcome,
    HR       = s$conf.int[1, "exp(coef)"],
    Lower    = s$conf.int[1, "2.5%"],
    Upper    = s$conf.int[1, "97.5%"],
    P        = s$coef[1, "p-value"]
  )
}
```

Making the dataframe for the kable

```{r}
results_adjusted <- bind_rows(
  extract_cox(cs_model_death_adjusted,      "Death / Deterioration"),
  extract_cox(cs_model_transplant_adjusted, "Transplant"),
  extract_crr(crr_death_adjusted,           "Death / Deterioration"),
  extract_crr(crr_transplant_adjusted,      "Transplant")
) |>
  mutate(
    HR = round(HR, 2),
    CI = paste0("(", round(Lower, 2), " , ", round(Upper, 2), ")"),
    P  = ifelse(P < 0.001, "<0.001", sprintf("%.3f", P))
  ) |>
  select(Outcome, HR, CI, P)
```

The Kable Table

```{r}
results_adjusted |>
  kable(
    col.names = c("Outcome", "Hazard Ratio*", "95% CI", "p-value"),
    align = "c",
    caption   = '<span style="font-size:22pt; font-weight:bold;">Association of Escalation Status with 30-Day Outcomes, Adjusted (Escalated vs. Stable)</span>'
  ) |>
  kable_styling(
    full_width = F,
    position = 'left',
    font_size = 18 # we can mess with this to change the title size
  ) |> 
  column_spec(1, width = '10em') |> 
  column_spec(2:4, width = "7em") |> 
  pack_rows("Cause-Specific Cox Model", 1, 2) |>
  pack_rows("Fine-Gray Subdistribution Model", 3, 4) |> 
  footnote(
    general = c("*Adjusted for gender, age, BMI, ethnicity, diabetes, blood type, and exception status.",
                "Reference group: Stable patients*")
  )

```

*\*Adjusted for gender, age, BMI, ethnicity, diabetes, blood type, and exception status. Reference group: Stable patients.*




```{r}
library(stargazer)
# install.packages("stargazer")
```





```{r}
# kable(summary(cs_model_death_adjusted))
```

# `modelsummary` package

```{r}
library(modelsummary)
```

```{r}
models_list <- 
  list("Death Cox-Model" = cs_model_death_adjusted, 
       "Transplant Cox-Model" = cs_model_transplant_adjusted,
       "Death Fine-Gray Model" = crr_death_adjusted,
       "Transplant Fine-Gray Model" = crr_transplant_adjusted)
```

```{r}
modelsummary(models_list, stars = T)
```

# Binomial Regression

```{r}

library(mets)
options(warn=1)

heart_data <- heart_data %>% drop_na(INIT_BMI_NEW)


#Logistic Regression with IPCW Binomial Regression (Unadjusted) (Death)

heart_data$CAUSE_DEATH <- ifelse(heart_data$EVENT_30 == 1, 1, 0)

out <- binreg(
  Event(heart_data$TIME_30, heart_data$CAUSE_DEATH) ~ ESCALATED,
  data = heart_data,
  time = 30
)

summary(out)

# Estimate: 0.67 w/ Stable as ref. Less death for escalated


#Logistic Regression with IPCW Binomial Regression (Adjusted) (Death)

out1 <- binreg(
  Event(heart_data$TIME_30, heart_data$CAUSE_DEATH) ~ ESCALATED+GENDER+INIT_BMI_NEW+INIT_AGE+ETHCAT+DIAB+BTYPE_REC+EXCEPTION,
  data = heart_data,
  time = 30
)

summary(out1)

# Estimate: 0.60 w/ Stable as ref. Less death for escalated after adjusting

```

```{r}


# Logistic Regression with IPCQ Binomial Regression (Unadjusted) (Transplant)

heart_data$CAUSE_T <- ifelse(heart_data$EVENT_30 == 2, 1, 0)


out2 <- binreg(
  Event(heart_data$TIME_30, heart_data$CAUSE_T) ~ ESCALATED,
  data = heart_data,
  time = 30
)

summary(out2)

# Estimate: 1.18 w/ Stab le as ref. Transplant higher in escalated


# Logistic Regression with IPCQ Binomial Regression (Adjusted) (Transplant)

out3 <- binreg(
  Event(heart_data$TIME_30, heart_data$CAUSE_T) ~ ESCALATED+GENDER+INIT_BMI_NEW+INIT_AGE+ETHCAT+DIAB+BTYPE_REC+EXCEPTION,
  data = heart_data,
  time = 30
)

summary(out3)

summary(out3)$coef

# Estimate: 1.19 w/ Stable as ref. Transplant is higher in escalated after adjusting
```
# Extracting info from Binomial Regression Models

Extract from binreg() model

```{r}

# Extract from crr model
extract_binreg <- function(model, outcome) {
  s <- summary(model)
  data.frame(
    Outcome  = outcome,
    HR       = s$exp[2, "Estimate"],
    Lower    = s$exp[2, "2.5%"],
    Upper    = s$exp[2, "97.5%"],
    P        = s$coef[2, "P-value"]
  )
}
```

Making the dataframe for the table 

```{r}
results_adjusted <- bind_rows(
  extract_cox(cs_model_death_adjusted,      "Death / Deterioration"),
  extract_cox(cs_model_transplant_adjusted, "Transplant"),
  extract_crr(crr_death_adjusted,           "Death / Deterioration"),
  extract_crr(crr_transplant_adjusted,      "Transplant"),
  extract_binreg(out1, "Death / Deterioration"),
  extract_binreg(out3, "Transplant")
) |>
  mutate(
    HR = round(HR, 2),
    CI = paste0("(", round(Lower, 2), " , ", round(Upper, 2), ")"),
    P  = ifelse(P < 0.001, "P < 0.001", sprintf("%.3f", P))
  ) |>
  select(Outcome, HR, CI, P)
```

The New Kable Table

```{r}
results_adjusted |>
  kable(
    col.names = c("Outcome", "Estimate*", "95% CI", "p-value"),
    align = "c",
    caption   = '<span style="font-size:22pt; font-weight:bold;">Association of Escalation Status with 30-Day Outcomes, Adjusted (Escalated vs. Stable)</span>'
  ) |>
  kable_styling(
    full_width = F,
    position = 'left',
    font_size = 18 # we can mess with this to change the title size
  ) |> 
  column_spec(1, width = '10em') |> 
  column_spec(2:4, width = "7em") |> 
  pack_rows("Cause-Specific Cox Model", 1, 2) |>
  pack_rows("Fine-Gray Subdistribution Model", 3, 4) |>
  pack_rows("Binomial Regression Model", 5, 6) 
```

*\*Adjusted for gender, age, BMI, ethnicity, diabetes, blood type, and exception status. Reference group: Stable patients.*
