---
title: "Side Analysis"
author: "Julia Aptekar"
format: html
embed-resources: true
---

## Goal
The goal of this side analysis is to study waitlist candidates from the time they reach status 2 until one of several possible outcomes occurs.

The landmark time, defined as time 0, is the first date on which a candidate is listed at status 2. For candidates initially listed at status 2, time 0 is their INIT_DATE. For candidates who reach status 2 later, time 0 is the first CHG_DATE on which their waitlist status becomes status 2.

After time 0, candidates are followed until one of the following outcomes occurs:

1. Death/deterioration
2. Heart transplantation
3. Recovery
4. Upgrade to Status 1
5. Downgrade to Status 3–6
6. Still on study or removed for another reason, treated as censoring
7. Delisting as Status 2999

## Idea
First, we identify the first date on which each candidate is listed at status 2. Candidates who are initially listed at status 2 have time 0 equal to their INIT_DATE. Candidates who are not initially listed at status 2 are included only if they later reach status 2; for these candidates, time 0 is the first CHG_DATE where their waitlist status becomes status 2.

Candidates who never reach status 2 are removed from the analysis. After time 0, we examine each candidate’s subsequent waitlist history and identify the first event that occurs among death/deterioration, heart transplantation, recovery, upgrade to status 1, downgrade to status 3–6, delisting as status 2999, or censoring.

The analysis time is calculated as the number of days from time 0 to the event or censoring date.

## Libraries/Loading
```{r}
library(dplyr)
library(lubridate)

heart <- read.csv("data/LAsBEST26_SUBSET.csv")
data <- readRDS("data/NEW_DATA.rds")
```

## Cleaning
```{r}
data <- data %>% 
  select(WL_ID_CODE, INIT_STAT, END_STAT, INIT_DATE, END_DATE, CHG_DATE, everything()) %>%
  rename(INITIAL_STATUS = INIT_STAT,
    CHG_STATUS = END_STAT) %>%
  mutate(INIT_DATE = ymd(INIT_DATE),
    END_DATE = ymd(END_DATE),
    CHG_DATE = ymd(CHG_DATE),
    INITIAL_STATUS = case_when(
      INITIAL_STATUS == 2110 ~ 1,
      INITIAL_STATUS == 2120 ~ 2,
      INITIAL_STATUS == 2130 ~ 3,
      INITIAL_STATUS == 2140 ~ 4,
      INITIAL_STATUS == 2150 ~ 5,
      INITIAL_STATUS == 2160 ~ 6,
      TRUE ~ NA_real_),
    CHG_STATUS = case_when(
      CHG_STATUS == 2110 ~ 1,
      CHG_STATUS == 2120 ~ 2,
      CHG_STATUS == 2130 ~ 3,
      CHG_STATUS == 2140 ~ 4,
      CHG_STATUS == 2150 ~ 5,
      CHG_STATUS == 2160 ~ 6,
      CHG_STATUS == 2999 ~ 2999,
      TRUE ~ NA_real_))
```

## Code
```{r}
groups <- data %>%
  arrange(WL_ID_CODE, CHG_DATE) %>%
  group_by(WL_ID_CODE) %>%
  summarise(
    first_status = first(INITIAL_STATUS),
    init_date = first(INIT_DATE),
    later_status2_date = {
      dates <- CHG_DATE[CHG_STATUS == 2 & !is.na(CHG_DATE)]
      if (length(dates) == 0) {
        as.Date(NA)
      } else {
        min(dates)
      }
    },
    .groups = "drop") %>%
  mutate(status2_chg_date = case_when(
      first_status == 2 ~ init_date,
      first_status != 2 & !is.na(later_status2_date) ~ later_status2_date,
      TRUE ~ as.Date(NA)),
    had2 = !is.na(status2_chg_date),
    group = case_when(
      first_status == 2 ~ 0,
      first_status != 2 & had2 ~ 1,
      TRUE ~ NA_real_),
    time0 = status2_chg_date) %>%
  filter(!is.na(group))
```

## Joins
```{r}
# This currently codes only status change events:
# 4 = upgrade to status 1
# 5 = downgrade to status 3-6
# 7 = delisted as status 2999
# Events 1, 2, 3, and 6 still need to be added using removal/outcome variables.

status_change_events <- data %>%
  left_join(groups %>% select(WL_ID_CODE, time0), by = "WL_ID_CODE") %>%
  filter(!is.na(time0)) %>%
  filter(CHG_DATE > time0) %>%
  mutate(event = case_when(
      INITIAL_STATUS == 2 & CHG_STATUS == 1 ~ 4,      # upgrade from status 2 to status 1
      INITIAL_STATUS == 2 & CHG_STATUS %in% 3:6 ~ 5,  # downgrade from status 2 to status 3-6
      INITIAL_STATUS == 2 & CHG_STATUS == 2999 ~ 7,   # delisted from status 2 to status 2999
      TRUE ~ NA_real_)) %>%
  filter(!is.na(event)) %>%
  group_by(WL_ID_CODE) %>%
  slice_min(CHG_DATE, with_ties = FALSE) %>%
  ungroup() %>%
  transmute(WL_ID_CODE,
    event_date = CHG_DATE,
    event = event)

# joining some variables from original heart dataset
heart_sub <- heart %>%
  select(WL_ID_CODE, INIT_DATE, EXCEPTION, END_DATE_NEW, REM_CD_NEW, FSTATUS, GENDER, INIT_BMI_NEW, INIT_AGE, ETHCAT, DIAB, BTYPE_REC) %>%
  distinct(WL_ID_CODE, .keep_all = TRUE)

heart3 <- groups %>%
  left_join(status_change_events, by = "WL_ID_CODE") %>%
  left_join(heart_sub, by = "WL_ID_CODE") %>%
  mutate(time_to_event = as.numeric(event_date - time0))
```

## Writing CSV
```{r}
write.csv(heart3, "data/heart3.csv", row.names=FALSE)
```