---
title: "Side Analysis"
author: "Julia Aptekar"
format: html
embed-resources: true
---

## Libraries/Loading
```{r}
library(dplyr)
library(lubridate)

d <- readRDS("data/NEW_DATA.rds")
```

## Cleaning
```{r}
data <- d %>% 
  select(WL_ID_CODE, WLREG_AUDIT_ID_CODE, INIT_STAT, INIT_DATE, CHG_DATE,
         UNOS_CAND_STAT_CD, END_DATE_NEW, FTIME, FSTATUS, REM_CD_NEW, 
         DAYSWAIT_CHRON, INIT_AGE) %>%
  mutate(INIT_DATE = ymd(INIT_DATE),
    END_DATE_NEW = ymd(END_DATE_NEW),
    CHG_DATE = ymd(CHG_DATE),
    INITIAL_STATUS = case_when(
      INIT_STAT == 2110 ~ 1,
      INIT_STAT == 2120 ~ 2,
      INIT_STAT == 2130 ~ 3,
      INIT_STAT == 2140 ~ 4,
      INIT_STAT == 2150 ~ 5,
      INIT_STAT == 2160 ~ 6,
      TRUE ~ NA_real_),
    STATUS = case_when(
      UNOS_CAND_STAT_CD == 2110 ~ 1,
      UNOS_CAND_STAT_CD == 2120 ~ 2,
      UNOS_CAND_STAT_CD == 2130 ~ 3,
      UNOS_CAND_STAT_CD == 2140 ~ 4,
      UNOS_CAND_STAT_CD == 2150 ~ 5,
      UNOS_CAND_STAT_CD == 2160 ~ 6,
      UNOS_CAND_STAT_CD == 2999 ~ 2999,
      TRUE ~ NA_real_))
```

## Code
```{r}
data1 <- data %>%
  group_by(WL_ID_CODE) %>%
  filter(any(STATUS == 2, na.rm = TRUE)) %>%
  ungroup()

# keeping rows starting at first status 2
data2 <- data1 %>%
  arrange(WL_ID_CODE, CHG_DATE, WLREG_AUDIT_ID_CODE) %>%
  group_by(WL_ID_CODE) %>%
  mutate(STAT2_INIT = min(CHG_DATE[STATUS == 2], na.rm = TRUE)) %>%
  filter(CHG_DATE >= STAT2_INIT) %>%
  ungroup()

groups <- data2 %>%
  group_by(WL_ID_CODE) %>%
  summarise(first_status = first(INITIAL_STATUS),
    init_date = first(INIT_DATE),
    time0 = first(CHG_DATE),
    .groups = "drop") %>%
  mutate(group = case_when(
      first_status == 2 ~ 0,
      TRUE ~ 1),
    n_days_to_status2 = as.numeric(time0 - init_date))
```

## Joins
```{r}
# choosing the first event after status 2
event_rows <- data2 %>%
  mutate(event_date = case_when(
      is.na(END_DATE_NEW) ~ CHG_DATE,
      is.na(CHG_DATE) ~ END_DATE_NEW,
      TRUE ~ pmin(CHG_DATE, END_DATE_NEW)),
    event_source = case_when(
      !is.na(END_DATE_NEW) & (is.na(CHG_DATE) | END_DATE_NEW <= CHG_DATE) ~ "removal",
      !is.na(CHG_DATE) & (is.na(END_DATE_NEW) | CHG_DATE < END_DATE_NEW) ~ "status_change",
      TRUE ~ NA_character_)) %>%
  filter(STATUS != 2 | END_DATE_NEW <= CHG_DATE) %>%
  group_by(WL_ID_CODE) %>%
  slice_min(event_date, with_ties = FALSE) %>%
  ungroup()

# coding the event type
heart_events <- event_rows %>%
  mutate(event = case_when(
      event_source == "removal" & REM_CD_NEW %in% c(5, 8, 13) ~ 1, # death
      event_source == "removal" & REM_CD_NEW %in% c(2, 3, 4, 14, 21, 22) ~ 2, # transplant
      event_source == "removal" & REM_CD_NEW == 12 ~ 3, # recovery
      event_source == "status_change" & STATUS == 1 ~ 4, # downgrade
      event_source == "status_change" & STATUS %in% 3:6 ~ 5, # upgrade
      event_source == "status_change" & STATUS == 2999 ~ 7, # delisted
      TRUE ~ 6)) %>% # censored
  select(WL_ID_CODE, event_date, event_source, event)

# merging back all original variables
sub <- d %>%
  distinct(WL_ID_CODE, .keep_all = TRUE)

heart3 <- groups %>%
  left_join(heart_events, by = "WL_ID_CODE") %>%
  left_join(sub, by = "WL_ID_CODE") %>%
  mutate(time_to_event = as.numeric(event_date - time0),
    age_at_status2 = INIT_AGE + n_days_to_status2 / 365.25) %>%
  filter(is.na(time_to_event) | time_to_event >= 0) %>%
  select(WL_ID_CODE, time0, event_date, event, event_source,
    time_to_event, group, n_days_to_status2, age_at_status2,
    first_status, init_date, INIT_AGE, everything())
```

## Checking counts
```{r}
# checking final sample size and date logic
heart3 %>%
  summarise(n_rows = n(),
    n_unique_ids = n_distinct(WL_ID_CODE),
    missing_time0 = sum(is.na(time0)),
    negative_time_to_event = sum(time_to_event < 0, na.rm = TRUE))

# checking event counts
heart3 %>% count(event, event_source, sort = TRUE)

# checking censoring codes
heart3 %>%
  filter(event == 6) %>%
  count(event_source, REM_CD_NEW, sort = TRUE)
```

## Writing CSV
```{r}
write.csv(heart3, "data/heart3.csv", row.names=FALSE)
```

## Variable Definitions

`time0` is the date when the candidate first reaches status 2. If the candidate started at status 2, `time0` is their `INIT_DATE`. If the candidate reached status 2 later, `time0` is their first `CHG_DATE` with status 2.

`group` shows how the candidate entered the status 2 group. Candidates with `group = 0` started at status 2. Candidates with `group = 1` reached status 2 later.

`event_date` is the date of the first event after `time0`.

`event_source` shows where the event came from. If `event_source = "removal"`, the event was based on `END_DATE_NEW` and `REM_CD_NEW`. If `event_source = "status_change"`, the event was based on `CHG_DATE` and `STATUS`.

`event` identifies the type of first event after `time0`:

1: Death/deterioration
2: Heart transplantation
3: Recovery
4: Upgrade to status 1
5: Downgrade to status 3–6
6: Censoring
7: Delisting as status 2999

`time_to_event` is the number of days from `time0` to `event_date`.

`n_days_to_status2` is the number of days from initial listing to `time0`.

`age_at_status2` is the candidate’s estimated age when they first reached status 2.
- For candidates who started at status 2, this is the same as their initial age. 
- For candidates who reached status 2 later, this is their initial age plus the time between initial listing and status 2, converted from days to years.