---
title: "Side Analysis"
author: "Julia Aptekar"
format: html
embed-resources: true
---

## Goal
The goal of this side analysis is to study waitlist candidates from the time they reach status 2 until one of several possible outcomes occurs.

The landmark time, defined as time0, is the first date on which a candidate is listed at status 2. For candidates initially listed at status 2, time 0 is their INIT_DATE. For candidates who reach status 2 later, time0 is the first CHG_DATE on which their waitlist status becomes status 2.

After time 0, candidates are followed until one of the following outcomes occurs:

1. Death/deterioration
2. Heart transplantation
3. Recovery
4. Upgrade to Status 1
5. Downgrade to Status 3–6
6. Still on study or removed for another reason, treated as censoring
7. Delisting as Status 2999

## Idea
First, we identify the first date on which each candidate is listed at status 2. Candidates who are initially listed at status 2 have time 0 equal to their INIT_DATE. Candidates who are not initially listed at status 2 are included only if they later reach status 2; for these candidates, time0 is the first CHG_DATE where their waitlist status becomes status 2.

Candidates who never reach status 2 are removed from the analysis. After time0, we examine each candidate’s waitlist history and identify the first event that occurs among death/deterioration, heart transplantation, recovery, upgrade to status 1, downgrade to status 3–6, delisting as status 2999, or censoring.

## Libraries/Loading
```{r}
library(dplyr)
library(lubridate)

d <- readRDS("data/NEW_DATA.rds")
```

## Cleaning
```{r}
data <- d %>% 
  select(WL_ID_CODE, WLREG_AUDIT_ID_CODE, INIT_STAT, INIT_DATE, CHG_DATE,
         UNOS_CAND_STAT_CD, END_DATE_NEW, FTIME, FSTATUS, REM_CD_NEW, DAYSWAIT_CHRON) %>%
  relocate(WL_ID_CODE, WLREG_AUDIT_ID_CODE, INIT_STAT, INIT_DATE, CHG_DATE,
           UNOS_CAND_STAT_CD, END_DATE_NEW, FSTATUS, FTIME, REM_CD_NEW) %>%
  mutate(INIT_DATE = ymd(INIT_DATE),
    END_DATE_NEW = ymd(END_DATE_NEW),
    CHG_DATE = ymd(CHG_DATE),
    INITIAL_STATUS = case_when(
      INIT_STAT == 2110 ~ 1,
      INIT_STAT == 2120 ~ 2,
      INIT_STAT == 2130 ~ 3,
      INIT_STAT == 2140 ~ 4,
      INIT_STAT == 2150 ~ 5,
      INIT_STAT == 2160 ~ 6,
      TRUE ~ NA_real_),
    STATUS = case_when(
      UNOS_CAND_STAT_CD == 2110 ~ 1,
      UNOS_CAND_STAT_CD == 2120 ~ 2,
      UNOS_CAND_STAT_CD == 2130 ~ 3,
      UNOS_CAND_STAT_CD == 2140 ~ 4,
      UNOS_CAND_STAT_CD == 2150 ~ 5,
      UNOS_CAND_STAT_CD == 2160 ~ 6,
      UNOS_CAND_STAT_CD == 2999 ~ 2999,
      TRUE ~ NA_real_))
```

## Code
```{r}
data <- data %>%
  group_by(WL_ID_CODE) %>%
  mutate(ANYSTAT2 = as.integer(any(STATUS == 2, na.rm = TRUE)),
    ANYSTAT1 = as.integer(any(STATUS == 1, na.rm = TRUE))) %>%
  ungroup()

dat1 <- data %>%
  filter(ANYSTAT2 == 1)

dat2 <- dat1 %>%
  arrange(WL_ID_CODE, CHG_DATE, WLREG_AUDIT_ID_CODE) %>%
  group_by(WL_ID_CODE) %>%
  mutate(FIRST_STAT2 = match(TRUE, STATUS == 2)) %>%
  filter(!is.na(FIRST_STAT2),
    row_number() >= FIRST_STAT2) %>%
  mutate(STAT2_INIT = first(CHG_DATE)) %>%
  ungroup() %>%
  select(-FIRST_STAT2)

groups <- dat2 %>%
  group_by(WL_ID_CODE) %>%
  summarise(first_status = first(INITIAL_STATUS),
    init_date = first(INIT_DATE),
    time0 = first(STAT2_INIT),
    .groups = "drop") %>%
  mutate(group = case_when(
    first_status == 2 ~ 0,
    TRUE ~ 1))
```

## Joins
```{r}
# event_date is the minimum of CHG_DATE and END_DATE_NEW
# event type comes from REM_CD_NEW if END_DATE_NEW is the minimum
# event type comes from STATUS if CHG_DATE is the minimum

event_rows <- dat2 %>%
  mutate(event_date = case_when(
      is.na(END_DATE_NEW) ~ CHG_DATE,
      is.na(CHG_DATE) ~ END_DATE_NEW,
      TRUE ~ pmin(CHG_DATE, END_DATE_NEW)),
    event_source = case_when(
      !is.na(END_DATE_NEW) & (is.na(CHG_DATE) | END_DATE_NEW <= CHG_DATE) ~ "removal",
      !is.na(CHG_DATE) & (is.na(END_DATE_NEW) | CHG_DATE < END_DATE_NEW) ~ "status_change",
      TRUE ~ NA_character_)) %>%
  filter(STATUS != 2 | END_DATE_NEW <= CHG_DATE) %>%
  group_by(WL_ID_CODE) %>%
  slice_min(event_date, with_ties = FALSE) %>%
  ungroup()

heart_events <- event_rows %>%
  mutate(event = case_when(
      event_source == "removal" & REM_CD_NEW %in% c(5, 8, 13) ~ 1, # death/deterioration
      event_source == "removal" & REM_CD_NEW %in% c(2, 3, 4, 14, 21, 22) ~ 2, # heart transplantation
      event_source == "removal" & REM_CD_NEW == 12 ~ 3, # recovery
      event_source == "status_change" & STATUS == 1 ~ 4, # upgrade to status 1
      event_source == "status_change" & STATUS %in% 3:6 ~ 5, # downgrade to status 3-6
      event_source == "status_change" & STATUS == 2999 ~ 7, # delisted as status 2999
      TRUE ~ 6)) %>% # censoring/other reason
  select(WL_ID_CODE, event_date, event_source, event)

# joining some variables from original dataset
sub <- d %>%
  select(WL_ID_CODE, EXCEPTION, GENDER, INIT_BMI_CALC, INIT_AGE, ETHNICITY, DIAB, ABO) %>%
  distinct(WL_ID_CODE, .keep_all = TRUE)

heart3 <- groups %>%
  left_join(heart_events, by = "WL_ID_CODE") %>%
  left_join(sub, by = "WL_ID_CODE") %>%
  mutate(time_to_event = as.numeric(event_date - time0))
```

## Writing CSV
```{r}
write.csv(heart3, "data/heart3.csv", row.names=FALSE)
```

## Variable Definitions
`time0` is the landmark date for each candidate. It represents the first date the candidate is listed at status 2. For candidates initially listed at status 2, `time0` is their `INIT_DATE`. For candidates who reach status 2 later, `time0` is the first `CHG_DATE` where their status becomes 2.

`group` identifies how the candidate entered the status 2 cohort. 
- Candidates with `group = 0` were initially listed at status 2. 
- Candidates with `group = 1` were not initially listed at status 2 but reached status 2 later.

`event_date` is the date of the first observed event after `time0`.

`event_source` identifies which part of the data was used to determine the first event. If `event_source = "removal"`, the event was based on `END_DATE_NEW` and `REM_CD_NEW`. If `event_source = "status_change"`, the event was based on `CHG_DATE` and `STATUS`.

`event` identifies the type of first event after `time0`. The event codes are:

1: Death/deterioration
2: Heart transplantation
3: Recovery
4: Upgrade to status 1
5: Downgrade to status 3–6
6: Still on study or removed for another reason, treated as censoring
7: Delisting as status 2999

`time_to_event` is the number of days from `time0` to `event_date`.

Note: Events based on removal outcomes are coded using `REM_CD_NEW`.

## Notes
- randomly sample some IDs from original datasets
- check censorship
- remove people with negative diffs in dates (chg date - time0)
- save all variables
- need 9870 obs
- next steps: create new age variable
  - we only have initial age at listing
  - want age at status2
  those who start at 2: age = init_age
  those who get into 2: age = init_age + n_day_at2
  - come up with idea on how to calculate that for those status 2 after waitlist
  - analysis: CIF plot (1 with all events, 1 focused on transplant vs death, 1 on downgrade vs upgrade), status2 at list vs. changed into status2, unadjusted Cox + FG models,